### 2024-08-20 ExPro.Rmd
init repo

---
title: "Examples of functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

To install the package the following line of code is needed:

```{r , include = FALSE}
devtools::load_all()
```

```{r load, eval=FALSE}
# require(devtools)
devtools::install_github("filippogambarota/criticalvalue")
library(criticalvalue)
```

## Examples on real data

For our examples on real data we used from the package 'psych' the dataset "holzinger.swineford" which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects. Once the package is retrieved with 'library', the dataset can be opened using 'data("name of the dataset")'. For simplicity we decided to rename it with a shorter name.

```{r open data}
library(psych)
library(psychTools)
```

```{r dataset}
data("holzinger.swineford")
Holz <- holzinger.swineford
```

#### Critical value for correlation

In the following chunk we want to get the critical value for a correlation of two cognitive sub-tests of 301 subjects. We can do so with the 'critical' function:

```{r correlation}
cc <- cor.test(Holz$t01_visperc, Holz$t02_cubes)
critical(cc)
```

First, the 'cor.test' function should be used on the two variables of interest and then the newly created object should be put into the 'critical' function. The output will add, underneath the usual output of the 'cor.test' function, the absolute critical value (rc), representing the smallest significant effect.

#### Critical value for t-test

Now we want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done with the same procedure using the 't.test' function:

```{r}
tt <- t.test(Holz$t01_visperc [ Holz$female == 1], Holz$t01_visperc [ Holz$female == 2])
critical(tt)
```

The output now gives a wider range of values: the cohen's d calculated on the data (d), the critical cohen's d (dc), the numerator of the formula for the critical cohen's d (bc), the cohen's d adjusted for small samples (g) and the critical cohen's d adjusted for small samples (gc).

#### Critical value for linear model

Now we want to calculate the critical coefficients for a linear model with visual perception predicted by age. We can do so by wrapping the 'critical' function into the 'summary' function:
```{r}
ll <- lm( t01_visperc ~ ageyr , data = Holz)
summary(critical(ll))
```

In the output, a column will appear next to the usual column of the estimates reporting critical values for both the intercept and the predictor(s).

## Examples with sample size

In the next examples we will show the use of the package's functions to calculate critical values in a prospective framework.

#### Critical t value from sample size

Let us say that we want to calculate the critical t value for a t-test for a sample of 1200 subjects for each group, the following chunk shows how to do so with the function 'critical_t2s':
```{r t from n}
n <- 1200
critical_t2s(n1 = n , n2 = n , hypothesis = "less")
```

The function requires to specify the hypothesis as "less", "greater" or "two.tailed". In output we can find the critical d and the critical g, which is the cohen's d adjusted for small samples.

#### Critical t value from sample size for a paired t-test.

Assuming that we want to calculate the critical value for a paired t-test (i.e. pre-post treatment), we can do so with the function 'critical_t2sp' in the following way:
```{r}
n <- 15
critical_t2sp(n = n , hypothesis = "less", conf.level = .95)
```

Specifying the direction of the hypothesis and the confidence interval, the function will give as output the standardized (dzc) and un-standardized (dc) critical cohen's d, and the critical cohen's d adjusted for small samples both standardized (gc) and un-standardized (gzc).

#### Critical value for a correlation from sample size

The same can be done for the critical value for a correlation on a defined sample size with the 'critical_cor' function:

```{r}
n <- 60
critical_cor(n = n, hypothesis = "two.sided", test = "z")
```

Once more, the direction of the hypothesis and the test to apply, either t-test or z-test, should be specified. The output will return the critical correlation value, the degrees of freedom and the type of test used.

#### Critical value of a beta from a model from saple size

Let us say that we want to know the critical beta of a model, we can use the 'critical_coef' function as shown in the following chunk:

```{r}
n <- 170
critical_coef(seb = .1, n = n, p = .01, conf.level = .95, hypothesis = "less")
```

It is important to specifying sample size, standard error of the estimate, p-value, confidence level and hypothesis. The output will return the critical beta and the test used.

### 2024-08-20 README.Rmd
init repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(scipen = 999, digits = 5)
```

# criticalvalue

<!-- badges: start -->
<!-- badges: end -->

The goal of criticalvalue is to ...

## Installation

You can install the development version of criticalvalue like so:

``` r
# require(devtools)
devtools::install_github("filippogambarota/criticalvalue")
```

## Installation (temporary)

The package is not production-ready thus to test the feature we suggest to clone or download the repository, open the R Project and load the functions with `devtools::load_all()`.

## Example from fitted objects

This is a basic example which shows you how to solve a common problem:

### T Test

```{r example}
# library(criticalvalue) # when the package will be ready
devtools::load_all() # loading all functions for testing

# t-test (welch)

x <- rnorm(30, 0.5, 1)
y <- rnorm(30, 0, 1)

ttest <- t.test(x, y)
critical(ttest)

# t-test (standard)

ttest <- t.test(x, y, var.equal = TRUE)
critical(ttest)

# t-test (standard) with monodirectional hyp

ttest <- t.test(x, y, var.equal = TRUE, alternative = "less")
critical(ttest)

# within the t-test object saved from critical we have all the new values

ttest <- critical(ttest)
str(ttest)
```

We can check the results using:

```{r}
ttest <- t.test(x, y)
ttest <- critical(ttest)

t <- ttest$bc/ttest$stderr # critical numerator / standard error

# should be 0.05 (or alpha)
(1 - pt(ttest$bc/ttest$stderr, ttest$parameter)) * 2
```

### Correlation Test

```{r}
# cor.test
ctest <- cor.test(x, y)
critical(ctest)
```

### Linear Model

```{r}
# linear model (unstandardized)
z <- rnorm(30)
q <- rnorm(30)

dat <- data.frame(x, y, z, q)
fit <- lm(y ~ x + q + z, data = dat)
fit <- critical(fit)
fit
summary(fit)

# linear model (standardized)
fit <- lm(y ~ x + q + z, data = dat)
fit <- critical(fit, standardize = TRUE)
fit
summary(fit)
```

### Meta-analysis

```{r}
require(metafor)
dat <- escalc(measure="RR", ai=tpos, bi=tneg, ci=cpos, di=cneg, data=dat.bcg)
fit <- rma(yi, vi, mods=cbind(ablat, year), data=dat)
critical(fit)
```

## Example from summary statistics

```{r}
x <- rnorm(30, 0.5, 1)
y <- rnorm(30, 0, 1)

ttest <- t.test(x, y)

m1 <- mean(x)
m2 <- mean(y)
sd1 <- sd(x)
sd2 <- sd(y)
n1 <- n2 <- 30

critical_t2s(m1, m2, sd1 = sd1, sd2 = sd2, n1 = n1, n2 = n2)

critical_t2s(t = ttest$statistic, se = ttest$stderr, n1 = n1, n2 = n2)

critical_t2s(t = ttest$statistic, n1 = n1, n2 = n2)
```

### 2024-08-20 paper/paper.Rmd
init repo

---
title             : "The benefits of reporting critical effect size values"
shorttitle        : "Reporting critical effect size values"
author: 
  - name          : "Ambra Perugini"
    affiliation   : "1"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Filippo Gambarota"
    affiliation   : "1"
    role:
      - "Writing – Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Methodology"
  - name          : "Enrico Toffalini"
    affiliation   : "2"
    role:
      - "Writing – Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Daniël Lakens"
    affiliation   : "3"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Massimiliano Pastore"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Livio Finos"
    affiliation   : "4"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Psicostat"
    affiliation   : "1"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Gianmarco Altoè"
    affiliation   : "1"
    corresponding: "yes"
    address       : "Via Venezia 8, 35131, Padova, Italy"
    email         : "gianmarco.altoe@unipd.it"
    role:
      - "Conceptualization"
      - "Writing - Review & Editing"
      - "Supervision"
affiliation:
  - id            : "1"
    institution   : "Department of Developmental and Social Psychology, University of Padova, Italy"
  - id            : "2"
    institution   : "Department of General Psychology, University of Padova, Italy"
  - id: "3"
    institution: "Eindhoven University of Technology, Netherlands"
  - id: "4"
    institution: "Department of Statistics, University of Padova, Italy"

abstract: |
  Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size. These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest. Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size. To assist researchers in calculating critical values, we have developed an R package that allows to report the critical effect size values of model parameters. The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis. What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations.

keywords          : critical effect size values, statistical significance, fucntions implementation 
wordcount         : "X"

bibliography      : "bibliography.bib"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : 
    papaja::apa6_pdf:
      latex_engine: xelatex
header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#>")
```

In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance. The "critical effect size values" refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample [@lakens2022sample]. To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's $r = -0.44$ and $r = 0.44$. As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference. This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.

Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null]. Over the past decades, numerous proposals have emerged to enhance inference based on this approach. These include conducting power analyses to determine sample size ($n$) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa]. Even for the most skeptical, there are suggestions that NHST may possess merits of its own. For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, that should subsequently be validated by "replication science".

In psychology, power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings. In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology. However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference [@altoe2020enhancing]. On the other hand, very large samples are occasionally available in psychological science, albeit less frequently. However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts [@wilson2020science]. Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation. In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.

There are two clear benefits of reporting the critical effect size for a corresponding test. First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations. If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test. An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect. Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant. This will focus the attention of researchers on the difference between statistical significance and practical significance. This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the 'crud factor' [@orben_crud_2020].

Let us consider the following two scenarios as examples. First, imagine researchers conducting a study involving a between-group comparison. Due to severe resource constraints, they are only able to collect a limited sample size of $n = 30$ ($n = 15$ per group). No statistically significant effect is detected with $p < 0.05$. Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent. Subsequently, they compute the critical values, revealing Cohen's $d = -0.75$ and $d = 0.75$. This indicates that any estimated |Cohen's $d| < 0.75$ will certainly fail to reach statistical significance. By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance. In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit. In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained on a small sample. Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.

In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore some bivariate correlations. With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes. However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions. In this context, the critical effect size values for significance are determined to be correlations of $r = \pm 0.03$. This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance. It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies. While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.

A similar scenario may arise in meta-analysis. Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance. While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted. Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.

To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research. The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community. The first case pertains to "Study 5" conducted by @mccrea2008self. This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure. One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward. With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups. The resulting effect was statistically significant, $t(26) = 1.87, p (one-tailed) < .05$. Based on the reported data, Cohen's $d = 0.736$, indicating a considerable effect size. However, the critical value for Cohen's $d$ in this context was $0.645$, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude. Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample ($n = 61$), $t(59)=2.325$. The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428). The second scenario involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences. With a notably large sample size of $n = 689'003$, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test). Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions. Despite the statistical significance, however, the critical values for Cohen's $d$ in this case are $-0.006$ and $+0.006$, emphasizing the importance of considering effect size beyond mere significance. This prompts reflection on whether an effect of such magnitude on such a large sample (actual $d = 0.02$) is truly meaningful in individuals' lives.

# How to Compute Critical Effect Size Values

In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models. These formulas have been incorporated into R functions of the package "criticalESvalues" accessible at: https://github.com/filippogambarota/criticalvalue, and are elaborated upon in the subsequent section.

## t-test

For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the $t$ statistics is computed as reported in Equation \@ref(eq:eq-gen-t).

\begin{equation}
    (\#eq:eq-gen-t)
    t = \frac{b}{\text{SE}_{b}}
\end{equation}

Where $b$ is the unstandardized effect size that depends on the type of test. For example, in the two sample t-test or the single mean for a one sample t-test. The denominator is the standard error of the numerator.

Similarly, Equation \@ref(eq:eq-gen-d) formalize a general form of the effect size.

\begin{equation}
    (\#eq:eq-gen-d)
    d = \frac{b}{s}
\end{equation}

Where $b$ is still the unstandardized effect size and $s$ is the standardization term. For example, in the two-sample case, $s$ is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.

### two-sample t-test

For the two sample t-test the critical value ($b_c$) is calculated using Equation \@ref(eq:eq-tw-sp). Where $t_c$ is the critical $t$ value calculated using a certain $\alpha$ value (e.g., $0.05$) and $n_1 + n_2 - 2$ degrees of freedom. Then the $b_c$ is simply divided by $s$ obtaining the standardized critical value. For the two-sample case ($s = s_p$ assuming equal variances) Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.

\begin{equation}
    (\#eq:eq-tw-sp)
    b_c = t_c \times \text{SE}_{b}
\end{equation}

\begin{equation}
    (\#eq:eq-s-p-twsmp)
    s_p = \sqrt{\frac{s_1^2 (n_1 - 1) + s_2^2 (n_2 - 1)}{n_1 + n_2 - 2}}
\end{equation}

When relaxing the assumption of equal variances (i.e., Welch's test) $s$ is simply the square root of the average between the two variances.

### one sample t-test

For the one sample case, the equations are the same. The only difference is that $b$ is the sample mean and $s$ is the sample standard deviation.

### paired sample t-test

For the paired sample case the situation is less straightforward. The reason is that the $t$ statistic is computed using $b$ being the average of the paired differences between the two (paired) samples and $s$ being the standard deviation/error of the differences. While for hypothesis testing and power analysis it is common to calculate the $d$ dividing by the standard deviation of the differences [@lakens2013calculating], this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation [@morris2002combining]. Practically, Equation \@ref(eq:eq-sp-sd) depicts the relationship between the pooled standard deviation and the standard deviation of the differences.

\begin{equation}
    (\#eq:eq-sp-sd)
    \begin{gathered}
        s_p =  \frac{s_D}{\sqrt{2(1 - \rho)}} \\
        s_D =  s_p \sqrt{2(1 - \rho)}
    \end{gathered}
\end{equation}

Where $\rho$ is the correlation between the two paired samples. Even if the hypothesis testing is computed using $s_D$, we reported the critical value both using the pooled standard deviation and the standard deviation of the differences. Beyond this difference, the equations are still the same as the previous example.

### Hedges's correction

The effect size calculated as in the previous step is known to be inflated thus we calculated also the corrected version usually known as Hedges's $g$ [@Hedges1981-za].

## Correlation Test

For the Pearson's correlation test, the critical value is computed using Equation \@ref(eq:eq-cor-c) where $t_c$ is still the critical value for the $t$ distribution and $n$ is the sample size.

\begin{equation}
    (\#eq:eq-cor-c)
    r_c = \frac{t_c}{\sqrt{n - 2 + t_c^2}}
\end{equation}

Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's $z$ transformation. Equation \@ref(eq:eq-f-rc) shows how to calculate the critical $F(r_c)$ value where $z_c$ is the critical value of the standard normal distribution with a chosen $\alpha$ level. Then the $F(r_c)$ can be transformed back into a correlation coefficient using Equation \@ref(eq:eq-bk-r).

\begin{equation}
    (\#eq:eq-f-rc)
    \begin{gathered}
        F(r_c) = \frac{z_c}{\sqrt{n - 3}}
    \end{gathered}
\end{equation}

\begin{equation}
    (\#eq:eq-bk-r)
    r_c = \frac{\exp(2z_c)-1}{\exp(2z_c)+1} = \operatorname{tanh}(z_c)
\end{equation}

<!-- %% see better this part -->

## Linear Regression

Hypothesis testing on single coefficients in linear regression (using the \texttt{lm} function in R) is performed using the Wald test. This test is basically a t-test where the $t$ value is calculated dividing the regression parameter $\beta_0, \dots, \beta_j$ by the standard error. Then a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept. The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical value as reported in Equation \@ref(eq:eq-critical-lm).

\begin{equation}
(\#eq:eq-wald-lm)
    t = \frac{\beta_j}{SE_{\beta_j}}
\end{equation}

\begin{equation}
(\#eq:eq-critical-lm)
    \beta_{jc} = t_c SE_{\beta_j}
\end{equation}

<!-- maybe worth adding more about standardization -->

In this case $\beta_{jc}$ is the unstandardized critical regression coefficient (as $b$ in the $t$-test section). We implemented also the possibility to use standardized regression coefficients. We can identify two types of standardization, namely full and partial. The full standardization involves both the response variable and the predictors. Regression coefficients are interpreted as the increase in standard deviations of the response variable for an increase of one standard deviation in the predictor, keeping fixed all other predictors. The partial standardization involves only the predictors and regression coefficients are interpreted as the increase in the (raw) response variable for a increase in one standard deviation of the predictor. By default, standardizing means dividing a quantity by the standard deviation and eventually centering (i.e., subtracting the mean). Gelman [@Gelman2008-zp] suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric. Equation \@ref(eq:eq-critical-lm) can be used but to calculate the critical value for the standardized regression coefficients the result need to be interpreted according to the chosen approach.

<!-- %% -->

### Meta-analysis

<!-- %% da rivedere -->

Meta-analysis allows to pool information from multiple studies related to a specific research question. The main advantage of meta-analysis is pooling multiple studies to obtain a more precise and powerful estimation of the effect. From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances. Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests. The calculation of critical values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm). The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.

### Other models

Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models. In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.

# Examples in R

In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalues" in R. The complete package can be accessed at: https://github.com/filippogambarota/criticalvalue. Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size. In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.

First the package should be downloaded and opened with the library function:

```{r, download-package}
# devtools::install_github("filippogambarota/criticalvalue")
library(criticalvalue)
```

For our examples on real data we used from the package 'psych' the dataset "holzinger.swineford" which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects. Once the package is retrieved with 'library', the dataset can be opened using 'data("name of the dataset")'. For simplicity we decided to rename it with a shorter name.

```{r load-dataset}
library(psych)
library(psychTools)
data("holzinger.swineford")
Holz <- holzinger.swineford
```

We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done with the same procedure using the 't.test' function:

```{r t-test}
tt <- t.test(Holz$t01_visperc[Holz$female == 1], 
             Holz$t01_visperc[Holz$female == 2])
critical(tt)
```

The output gives a wide range of values: the Cohen's $d$ calculated on the data ($d$), the critical Cohen's $d$ ($dc$), the numerator of the formula for the critical Cohen's $d$ ($bc$), the Cohen's $d$ adjusted for small samples ($g$) and the critical Cohen's $d$ adjusted for small samples ($gc$).

In the next example we will show the use of the package's function `critical_cor` to calculate the critical value for a correlation in a prospective framework.

```{r correlation}
n <- 60
critical_cor(n = n, hypothesis = "two.sided", test = "z")
```

The direction of the hypothesis and the test to apply, either $t$-test or $z$-test, should be specified. The output will return the critical correlation value, the degrees of freedom and the type of test used.

# Discussion

With the present article, we propose that researchers compute and report the "critical effect size value(s)" in their empirical articles. This is not intended to replace other strategies aimed at enhancing the NHST approach to inference. Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the a priori planning for statistical power are valuable in their own right. Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it). Interestingly, the critical value can be retrospectively applied even to already published studies. This possible application facilitates potential reframing of the original interpretations. Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.

An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations. The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings. For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects. To what extent can the reported results be interpreted, precisely? Knowing the critical value provides us with a clear benchmark. Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this. But is it of real theoretical relevance? If in comparing two groups, such as controls versus treatment, any Cohen's $d > 0.07$ would reach significance, is statistical significance enough to signal a "successful" treatment? Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical value certainly prompts some appropriate interpretive caution.

Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant. For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations. But as long as all correlations are based on the same sample size, researchers can simply remark 'the critical effect size is $r = 0.3$' and readers will know that all correlations larger than this value are statistically significant.

Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical value(s) can also serve an educational purpose. It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size. By highlighting a critical value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant. Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance. This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.

Real-case scenarios may not always be that simple. Hence, we chose to expand the application of computing critical significance values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients. This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes. A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance. Notably, however, this prerequisite aligns with the requirements of APA style guidelines concerning the reporting of effect sizes. For further illustration and application, additional examples are provided in the Supplementary online material.

We suggest that reporting critical effect size values is particularly valuable when sample size planning was not feasible or did not occur \textit{a priori}. In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward. However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical values for the obtained sample can help provide context for interpretation. Critical values can be computed and interpreted even retrospectively or for studies that have already been published.

In conclusion, the reporting of the critical value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings. While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2024-08-20 paper/paper.tex
init repo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The benefits of reporting critical effect size values},
  pdfauthor={Ambra Perugini1, Filippo Gambarota1, Enrico Toffalini2, Daniël Lakens3, Massimiliano Pastore1, Livio Finos4, Psicostat1, \& Gianmarco Altoè1},
  pdflang={en-EN},
  pdfkeywords={critical effect size values, statistical significance, fucntions implementation},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The benefits of reporting critical effect size values}
\author{Ambra Perugini\textsuperscript{1}, Filippo Gambarota\textsuperscript{1}, Enrico Toffalini\textsuperscript{2}, Daniël Lakens\textsuperscript{3}, Massimiliano Pastore\textsuperscript{1}, Livio Finos\textsuperscript{4}, Psicostat\textsuperscript{1}, \& Gianmarco Altoè\textsuperscript{1}}
\date{}

\shorttitle{Reporting critical effect size values}

\authornote{

The authors made the following contributions. Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Conceptualization, Writing - Review \& Editing, Supervision; Livio Finos: Conceptualization, Writing - Review \& Editing, Supervision; Psicostat: Conceptualization, Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.

Correspondence concerning this article should be addressed to Gianmarco Altoè, Via Venezia 8, 35131, Padova, Italy. E-mail: \href{mailto:gianmarco.altoe@unipd.it}{\nolinkurl{gianmarco.altoe@unipd.it}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Developmental and Social Psychology, University of Padova, Italy\\\textsuperscript{2} Department of General Psychology, University of Padova, Italy\\\textsuperscript{3} Eindhoven University of Technology, Netherlands\\\textsuperscript{4} Department of Statistics, University of Padova, Italy}

\abstract{%
Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size. These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest. Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size. To assist researchers in calculating critical values, we have developed an R package that allows to report the critical effect size values of model parameters. The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis. What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations.
}

\begin{document}
\maketitle

In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance. The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample (Lakens, 2022). To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's \(r = -0.44\) and \(r = 0.44\). As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference. This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.

Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001). Over the past decades, numerous proposals have emerged to enhance inference based on this approach. These include conducting power analyses to determine sample size (\(n\)) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019). Even for the most skeptical, there are suggestions that NHST may possess merits of its own. For instance, Wilson et al. (2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, that should subsequently be validated by ``replication science''.

In psychology, power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings. In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size (Lakens, 2022), and this scenario is arguably widespread in psychology. However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference (Altoè et al., 2020). On the other hand, very large samples are occasionally available in psychological science, albeit less frequently. However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts (Wilson et al., 2020). Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation. In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.

There are two clear benefits of reporting the critical effect size for a corresponding test. First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations. If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test. An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect. Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant. This will focus the attention of researchers on the difference between statistical significance and practical significance. This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the `crud factor' (Orben \& Lakens, 2020).

Let us consider the following two scenarios as examples. First, imagine researchers conducting a study involving a between-group comparison. Due to severe resource constraints, they are only able to collect a limited sample size of \(n = 30\) (\(n = 15\) per group). No statistically significant effect is detected with \(p < 0.05\). Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent. Subsequently, they compute the critical values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\). This indicates that any estimated \textbar Cohen's \(d| < 0.75\) will certainly fail to reach statistical significance. By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance. In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit. In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained on a small sample. Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.

In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore some bivariate correlations. With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes. However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions. In this context, the critical effect size values for significance are determined to be correlations of \(r = \pm 0.03\). This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance. It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies. While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.

A similar scenario may arise in meta-analysis. Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance. While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted. Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.

To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research. The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community. The first case pertains to ``Study 5'' conducted by McCrea (2008). This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure. One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward. With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups. The resulting effect was statistically significant, \(t(26) = 1.87, p (one-tailed) < .05\). Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size. However, the critical value for Cohen's \(d\) in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude. Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\). The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428). The second scenario involves the study by Kramer et al. (2014), which explored the impact of emotional content on Facebook users' experiences. With a notably large sample size of \(n = 689'003\), the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test). Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions. Despite the statistical significance, however, the critical values for Cohen's \(d\) in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance. This prompts reflection on whether an effect of such magnitude on such a large sample (actual \(d = 0.02\)) is truly meaningful in individuals' lives.

\hypertarget{how-to-compute-critical-effect-size-values}{%
\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}}

In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models. These formulas have been incorporated into R functions of the package ``criticalESvalues'' accessible at: \url{https://github.com/filippogambarota/criticalvalue}, and are elaborated upon in the subsequent section.

\hypertarget{t-test}{%
\subsection{t-test}\label{t-test}}

For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the \(t\) statistics is computed as reported in Equation \eqref{eq:eq-gen-t}.

\begin{equation}
    \label{eq:eq-gen-t}
    t = \frac{b}{\text{SE}_{b}}
\end{equation}

Where \(b\) is the unstandardized effect size that depends on the type of test. For example, in the two sample t-test or the single mean for a one sample t-test. The denominator is the standard error of the numerator.

Similarly, Equation \eqref{eq:eq-gen-d} formalize a general form of the effect size.

\begin{equation}
    \label{eq:eq-gen-d}
    d = \frac{b}{s}
\end{equation}

Where \(b\) is still the unstandardized effect size and \(s\) is the standardization term. For example, in the two-sample case, \(s\) is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.

\hypertarget{two-sample-t-test}{%
\subsubsection{two-sample t-test}\label{two-sample-t-test}}

For the two sample t-test the critical value (\(b_c\)) is calculated using Equation \eqref{eq:eq-tw-sp}. Where \(t_c\) is the critical \(t\) value calculated using a certain \(\alpha\) value (e.g., \(0.05\)) and \(n_1 + n_2 - 2\) degrees of freedom. Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical value. For the two-sample case (\(s = s_p\) assuming equal variances) Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.

\begin{equation}
    \label{eq:eq-tw-sp}
    b_c = t_c \times \text{SE}_{b}
\end{equation}

\begin{equation}
    \label{eq:eq-s-p-twsmp}
    s_p = \sqrt{\frac{s_1^2 (n_1 - 1) + s_2^2 (n_2 - 1)}{n_1 + n_2 - 2}}
\end{equation}

When relaxing the assumption of equal variances (i.e., Welch's test) \(s\) is simply the square root of the average between the two variances.

\hypertarget{one-sample-t-test}{%
\subsubsection{one sample t-test}\label{one-sample-t-test}}

For the one sample case, the equations are the same. The only difference is that \(b\) is the sample mean and \(s\) is the sample standard deviation.

\hypertarget{paired-sample-t-test}{%
\subsubsection{paired sample t-test}\label{paired-sample-t-test}}

For the paired sample case the situation is less straightforward. The reason is that the \(t\) statistic is computed using \(b\) being the average of the paired differences between the two (paired) samples and \(s\) being the standard deviation/error of the differences. While for hypothesis testing and power analysis it is common to calculate the \(d\) dividing by the standard deviation of the differences (Lakens, 2013), this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation (Morris \& DeShon, 2002). Practically, Equation \eqref{eq:eq-sp-sd} depicts the relationship between the pooled standard deviation and the standard deviation of the differences.

\begin{equation}
    \label{eq:eq-sp-sd}
    \begin{gathered}
        s_p =  \frac{s_D}{\sqrt{2(1 - \rho)}} \\
        s_D =  s_p \sqrt{2(1 - \rho)}
    \end{gathered}
\end{equation}

Where \(\rho\) is the correlation between the two paired samples. Even if the hypothesis testing is computed using \(s_D\), we reported the critical value both using the pooled standard deviation and the standard deviation of the differences. Beyond this difference, the equations are still the same as the previous example.

\hypertarget{hedgess-correction}{%
\subsubsection{Hedges's correction}\label{hedgess-correction}}

The effect size calculated as in the previous step is known to be inflated thus we calculated also the corrected version usually known as Hedges's \(g\) (Hedges, 1981).

\hypertarget{correlation-test}{%
\subsection{Correlation Test}\label{correlation-test}}

For the Pearson's correlation test, the critical value is computed using Equation \eqref{eq:eq-cor-c} where \(t_c\) is still the critical value for the \(t\) distribution and \(n\) is the sample size.

\begin{equation}
    \label{eq:eq-cor-c}
    r_c = \frac{t_c}{\sqrt{n - 2 + t_c^2}}
\end{equation}

Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's \(z\) transformation. Equation \eqref{eq:eq-f-rc} shows how to calculate the critical \(F(r_c)\) value where \(z_c\) is the critical value of the standard normal distribution with a chosen \(\alpha\) level. Then the \(F(r_c)\) can be transformed back into a correlation coefficient using Equation \eqref{eq:eq-bk-r}.

\begin{equation}
    \label{eq:eq-f-rc}
    \begin{gathered}
        F(r_c) = \frac{z_c}{\sqrt{n - 3}}
    \end{gathered}
\end{equation}

\begin{equation}
    \label{eq:eq-bk-r}
    r_c = \frac{\exp(2z_c)-1}{\exp(2z_c)+1} = \operatorname{tanh}(z_c)
\end{equation}

\hypertarget{linear-regression}{%
\subsection{Linear Regression}\label{linear-regression}}

Hypothesis testing on single coefficients in linear regression (using the \texttt{lm} function in R) is performed using the Wald test. This test is basically a t-test where the \(t\) value is calculated dividing the regression parameter \(\beta_0, \dots, \beta_j\) by the standard error. Then a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept. The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical value as reported in Equation \eqref{eq:eq-critical-lm}.

\begin{equation}
\label{eq:eq-wald-lm}
    t = \frac{\beta_j}{SE_{\beta_j}}
\end{equation}

\begin{equation}
\label{eq:eq-critical-lm}
    \beta_{jc} = t_c SE_{\beta_j}
\end{equation}

In this case \(\beta_{jc}\) is the unstandardized critical regression coefficient (as \(b\) in the \(t\)-test section). We implemented also the possibility to use standardized regression coefficients. We can identify two types of standardization, namely full and partial. The full standardization involves both the response variable and the predictors. Regression coefficients are interpreted as the increase in standard deviations of the response variable for an increase of one standard deviation in the predictor, keeping fixed all other predictors. The partial standardization involves only the predictors and regression coefficients are interpreted as the increase in the (raw) response variable for a increase in one standard deviation of the predictor. By default, standardizing means dividing a quantity by the standard deviation and eventually centering (i.e., subtracting the mean). Gelman (Gelman, 2008) suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric. Equation \eqref{eq:eq-critical-lm} can be used but to calculate the critical value for the standardized regression coefficients the result need to be interpreted according to the chosen approach.

\hypertarget{meta-analysis}{%
\subsubsection{Meta-analysis}\label{meta-analysis}}

Meta-analysis allows to pool information from multiple studies related to a specific research question. The main advantage of meta-analysis is pooling multiple studies to obtain a more precise and powerful estimation of the effect. From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances. Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests. The calculation of critical values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}. The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.

\hypertarget{other-models}{%
\subsubsection{Other models}\label{other-models}}

Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models. In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.

\hypertarget{examples-in-r}{%
\section{Examples in R}\label{examples-in-r}}

In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalues'' in R. The complete package can be accessed at: \url{https://github.com/filippogambarota/criticalvalue}. Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size. In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.

First the package should be downloaded and opened with the library function:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# devtools::install\_github("filippogambarota/criticalvalue")}
\FunctionTok{library}\NormalTok{(criticalvalue)}
\end{Highlighting}
\end{Shaded}

For our examples on real data we used from the package `psych' the dataset ``holzinger.swineford'' which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects. Once the package is retrieved with `library', the dataset can be opened using `data(``name of the dataset'')'. For simplicity we decided to rename it with a shorter name.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(psych)}
\FunctionTok{library}\NormalTok{(psychTools)}
\FunctionTok{data}\NormalTok{(}\StringTok{"holzinger.swineford"}\NormalTok{)}
\NormalTok{Holz }\OtherTok{\textless{}{-}}\NormalTok{ holzinger.swineford}
\end{Highlighting}
\end{Shaded}

We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done with the same procedure using the `t.test' function:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tt }\OtherTok{\textless{}{-}} \FunctionTok{t.test}\NormalTok{(Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{1}\NormalTok{], }
\NormalTok{             Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{2}\NormalTok{])}
\FunctionTok{critical}\NormalTok{(tt)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> 
#>  Welch Two Sample t-test
#> 
#> data:  Holz$t01_visperc[Holz$female == 1] and Holz$t01_visperc[Holz$female == 2]
#> t = 1.4095, df = 298.9, p-value = 0.1597
#> alternative hypothesis: true difference in means is not equal to 0
#> 95 percent confidence interval:
#>  -0.06419229  0.38823118
#> sample estimates:
#> mean of x mean of y 
#>  4.314090  4.152071 
#> 
#> |== Effect Size and Critical Value ==| 
#> d = 0.1623665 dc = ± 0.2269609 bc = ± 0.2262117 
#> g = 0.1619587 gc = ± 0.2263909
\end{verbatim}

The output gives a wide range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).

In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical value for a correlation in a prospective framework.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{60}
\FunctionTok{critical\_cor}\NormalTok{(}\AttributeTok{n =}\NormalTok{ n, }\AttributeTok{hypothesis =} \StringTok{"two.sided"}\NormalTok{, }\AttributeTok{test =} \StringTok{"z"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> $rc
#> [1] 0.2539247
#> 
#> $df
#> [1] 58
#> 
#> $test
#> [1] "z"
\end{verbatim}

The direction of the hypothesis and the test to apply, either \(t\)-test or \(z\)-test, should be specified. The output will return the critical correlation value, the degrees of freedom and the type of test used.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

With the present article, we propose that researchers compute and report the ``critical effect size value(s)'' in their empirical articles. This is not intended to replace other strategies aimed at enhancing the NHST approach to inference. Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the a priori planning for statistical power are valuable in their own right. Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it). Interestingly, the critical value can be retrospectively applied even to already published studies. This possible application facilitates potential reframing of the original interpretations. Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.

An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations. The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings. For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects. To what extent can the reported results be interpreted, precisely? Knowing the critical value provides us with a clear benchmark. Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this. But is it of real theoretical relevance? If in comparing two groups, such as controls versus treatment, any Cohen's \(d > 0.07\) would reach significance, is statistical significance enough to signal a ``successful'' treatment? Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical value certainly prompts some appropriate interpretive caution.

Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant. For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations. But as long as all correlations are based on the same sample size, researchers can simply remark `the critical effect size is \(r = 0.3\)' and readers will know that all correlations larger than this value are statistically significant.

Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical value(s) can also serve an educational purpose. It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size. By highlighting a critical value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant. Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance. This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.

Real-case scenarios may not always be that simple. Hence, we chose to expand the application of computing critical significance values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients. This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes. A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance. Notably, however, this prerequisite aligns with the requirements of APA style guidelines concerning the reporting of effect sizes. For further illustration and application, additional examples are provided in the Supplementary online material.

We suggest that reporting critical effect size values is particularly valuable when sample size planning was not feasible or did not occur \textit{a priori}. In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward. However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical values for the obtained sample can help provide context for interpretation. Critical values can be computed and interpreted even retrospectively or for studies that have already been published.

In conclusion, the reporting of the critical value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings. While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-altoe2020enhancing}{}}%
Altoè, G., Bertoldo, G., Zandonella Callegher, C., Toffalini, E., Calcagnı̀, A., Finos, L., \& Pastore, M. (2020). Enhancing statistical inference in psychological research via prospective and retrospective design analysis. \emph{Frontiers in Psychology}, \emph{10}, 499756.

\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.

\leavevmode\vadjust pre{\hypertarget{ref-open2015estimating}{}}%
Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.

\leavevmode\vadjust pre{\hypertarget{ref-funder2019evaluating}{}}%
Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: Sense and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008-zp}{}}%
Gelman, A. (2008). Scaling regression inputs by dividing by two standard deviations. \emph{Statistics in Medicine}, \emph{27}, 2865--2873. \url{https://doi.org/10.1002/sim.3107}

\leavevmode\vadjust pre{\hypertarget{ref-gigerenzer2004null}{}}%
Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.

\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
Hedges, L. V. (1981). Distribution theory for glass's estimator of effect size and related estimators. \emph{Journal of Educational and Behavioral Statistics: A Quarterly Publication Sponsored by the American Educational Research Association and the American Statistical Association}, \emph{6}, 107--128. \url{https://doi.org/10.3102/10769986006002107}

\leavevmode\vadjust pre{\hypertarget{ref-hedges1984estimation}{}}%
Hedges, L. V. (1984). Estimation of effect size under nonrandom sampling: The effects of censoring studies yielding statistically insignificant mean differences. \emph{Journal of Educational Statistics}, \emph{9}(1), 61--85.

\leavevmode\vadjust pre{\hypertarget{ref-kramer2014experimental}{}}%
Kramer, A. D., Guillory, J. E., \& Hancock, J. T. (2014). Experimental evidence of massive-scale emotional contagion through social networks. \emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{111}(24), 8788.

\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.

\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.

\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.

\leavevmode\vadjust pre{\hypertarget{ref-mccrea2008self}{}}%
McCrea, S. M. (2008). Self-handicapping, excuse making, and counterfactual thinking: Consequences for self-esteem and future motivation. \emph{Journal of Personality and Social Psychology}, \emph{95}(2), 274.

\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.

\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}

\leavevmode\vadjust pre{\hypertarget{ref-szucs2017empirical}{}}%
Szucs, D., \& Ioannidis, J. P. (2017). Empirical assessment of published effect sizes and power in the recent cognitive neuroscience and psychology literature. \emph{PLoS Biology}, \emph{15}(3), e2000797.

\leavevmode\vadjust pre{\hypertarget{ref-transue2019apa}{}}%
Transue, B. (2019). \emph{APA style 7th edition}.

\leavevmode\vadjust pre{\hypertarget{ref-wilson2020science}{}}%
Wilson, B. M., Harris, C. R., \& Wixted, J. T. (2020). Science is not a signal detection problem. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(11), 5559--5567.

\end{CSLReferences}

\end{document}

### 2024-08-20 R/criticalvalue-funs.R

@@ -1,0 +1,500 @@
+
+#' crit_from_t_t2s
+#' @description
+#' This function allows to calculate the cohen's d and the critical d given the t-value for a two samples t-test, sample size of the two groups and the standard error, specifying the confidence level of the interval, the direction of the hypothesis and the variance parameter.
+#' 
+#' @param t the t value.
+#' @param n1 a number corresponding to the sample size of group 1.
+#' @param n2 a number corresponding to the sample size of group 2.
+#' @param se a number corresponding to the standard error. 
+#' @param conf.level confidence level of the interval.
+#' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
+#' @param var.equal a logical variable indicating whether to treat the two variances as being equal.
+#'
+#' @export
+#' @return the output returns a "d" which is the cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the "bc" is the numerator of the formula from which the d is calculated and "df" are the degrees of freedom.
+#'
+crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
+                            conf.level, 
+                            hypothesis,
+                            var.equal = FALSE){
+  if(!var.equal){
+    warning("When var.equal = FALSE the critical value calculated from t assume sd1 = sd2!")
+  }
+  alpha <- .get_alpha(conf.level, hypothesis)
+  df <- n1 + n2 - 2
+  tc <- abs(qt(alpha, df))
+  if(is.null(t)){
+    warning("When t is NULL, d cannot be computed, returning NA")
+    d <- NA
+  }else{
+    d <- t * sqrt(1/n1 + 1/n2)
+  }
+  dc <- tc * sqrt(1/n1 + 1/n2)
+  if(is.null(se)){
+    warning("When se = NULL bc cannot be computed, returning NA!")
+    bc <- NA
+  }else{
+    bc <- (tc * se)
+  }
+  out <- list(d = d, dc = dc, bc = bc, df = df)
+  return(out)
+}
+
+#' crit_from_data_t2s
+#' @description
+#' This function allows to calculate the cohen's d and the critical d given the mean of the two groups, the standard deviation of the means and sample size of the two groups, specifying the confidence level of the interval, the direction of the hypothesis, the standard error, degrees of freedom and the variance parameter.
+#' 
[…]

### 2024-08-20 R/criticalvalue-methods.R

@@ -1,0 +1,233 @@
+
+#' Compute critical effect size values for a range of different objects
+#' @description Compute critical effect size values for objects of classes htest (Student's t-test, correlation test), lm, or rma
+#' @param x an object of class htest (available for t.test, cor.test), lm (a linear regression model), or rma (a meta-analytic linear model fitted with the "metafor" package)
+#' @param ... other arguments passed for the specific method, depending on the class of the object x
+#' @return an object of class critvalue
+#' @export
+critical <- function(x, ...){
+  UseMethod("critical")
+}
+
+
+#' Compute critical effect size values for t-test and correlation tests
+#' @description Compute critical effect size values for objects of classes htest computed with t.test or cor.test
+#' @param x an object of class htest (available for t.test, cor.test)
+#' @param type a value among "Two Sample", "One Sample", "correlation", "Paired"
+#' @return an object of class critvalue
+#' @export
+critical.htest <- function(x, type = NULL){
+  # all implemented subtype of htest objects
+  mtds <- c("Two Sample", "One Sample", "correlation", "Paired")
+  
+  if(all(!grepl(paste0(mtds, collapse = "|"), x$method))){
+    stop(
+      sprintf("method %s of class %s not implemented yet!",
+              x$method, class(x))
+    )
+  } else {
+    D <- insight::get_data(x)
+    conf.level <- attributes(x$conf.int)$conf.level
+    
+    # hypothesis <- ifelse(x$alternative == "two.sided", "2t", "1t")
+    hypothesis <- x$alternative
+    
+    alpha <- .get_alpha(conf.level, hypothesis)
+    df <- x$parameter
+    tc <- qt(alpha, df)
+    t <- x$statistic
+    se <- x$stderr
+    b <- x$estimate
+    method <- x$method
+  }
+  
+  if(grepl("correlation", method)){
+    n <- df + 2
+    cc <- critical_cor(n = n, conf.level = conf.level, hypothesis = hypothesis)
+    dc <- bc <- cc$rc
+    d <- b
+    class(x) <- c("critvalue", "ctest", "htest")
+  }else{ # if t-test
+    if(grepl("Two Sample", method)){
[…]

### 2024-08-20 R/utils.R

@@ -1,0 +1,50 @@
+.p_value <- function(q,
+                 df = NULL,
+                 test = c("t", "z"),
+                 conf.level = 0.95,
+                 hypothesis = c("1t", "2t")){
+  hypothesis <- match.arg(hypothesis)
+  test <- match.arg(test)
+  # two tails
+  if(test == "t"){
+    p <- 2 * pt(-abs(q), df)
+  }else{
+    p <- 2 * pnorm(-abs(q))
+  }
+  if(hypothesis == "1t"){
+    p <- p / 2
+  }
+  return(p)
+}
+
+.get_alpha <- function(conf.level = 0.95,
+                      hypothesis = c("2t", "1t")){
+  hypothesis <- if(hypothesis == "two.sided") "2t" else "1t"
+  alpha <- 1 - conf.level
+  if(hypothesis == "2t"){
+    alpha <- alpha / 2
+  }
+  return(alpha)
+}
+
+.get_J <- function(df){
+  # taken from https://github.com/easystats/effectsize/blob/a579a5093549be2edc264e2d1da95726ff4f4369/R/cohens_d.R#L315
+  exp(lgamma(df/2) - log(sqrt(df/2)) - lgamma((df - 1)/2))
+}
+
+.round_list <- function(x, digits = getOption("digits")){
+  out <- lapply(x, function(el){
+    if(is.numeric(el)){
+      round(el, digits)
+    } else{
+      el
+    }
+  })
+  class(out) <- class(x)
+  out
+}
+
+.hyp_sign <- function(x, hypothesis){
+  x <- abs(x)
+  ifelse(hypothesis %in% c("two.sided", "greater"), x, -x) 
+}

### 2024-08-20 example.R

@@ -1,0 +1,38 @@
+
+devtools::load_all()
+library(criticalvalue)
+
+# two sample t.test (var.equal = TRUE)
+
+x <- rnorm(500, 0.5, 1)
+y <- rnorm(500, 0, 1)
+
+tt <- t.test(x, y)
+tt
+cc = cor.test(x,y)
+fit <- lm(y ~ x)
+
+critical(fit)
+critical(fit, standardize = TRUE)
+critical(tt)
+critical(cc)
+
+n = 20
+yi = rnorm(n, .5, .2)
+vi = runif(n,.05,.10)
+xi = rnorm(n, 2, 1)
+library(metafor)
+fitrma = rma(yi=yi,vi=vi,mods=~xi)
+summary(fitrma)
+critical(fitrma)
+
+
+
+niter = 1e4
+n = 500
+ps = rep(NA,niter)
+
+pwr::pwr.t.test(n=500,power=0.8)
+metafor:::.cmicalc()
+
+

### 2024-08-21 R/criticalvalue-funs.R
update

@@ -1,4 +1,3 @@
-
 #' crit_from_t_t2s
 #' @description
 #' This function allows to calculate the cohen's d and the critical d given the t-value for a two samples t-test, sample size of the two groups and the standard error, specifying the confidence level of the interval, the direction of the hypothesis and the variance parameter.
@@ -37,7 +36,7 @@ crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
   }else{
     bc <- (tc * se)
   }
-  out <- list(d = d, dc = dc, bc = bc, df = df)
+  out <- list(d = d, dc = dc, bc = bc, se = se, df = df)
   return(out)
 }
 
@@ -81,14 +80,14 @@ crit_from_data_t2s <- function(m1, m2,
   }else{ # standard
     # pooled sd
     s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1)) / (n1 + n2 - 2))
-    if(is.null(se)) se <- s * (1/n1 + 1/n2)
+    if(is.null(se)) se <- s * sqrt(1/n1 + 1/n2)
     if(is.null(df)) df <- n1 + n2 - 2
   }
   tc <- abs(qt(alpha, df))
   d <- b / s
   bc <- tc * se
   dc <- tc * sqrt(1/n1 + 1/n2)
-  out <- list(d = d, dc = dc, bc = bc, df = df)
+  out <- list(d = d, dc = dc, bc = bc, se = se, df = df)
   return(out)
 }
 
@@ -126,7 +125,7 @@ crit_from_t_t1s <- function(t = NULL, n, se = NULL,
   }else{
     bc <- tc * se
   }
-  out <- list(d = d, dc = dc, bc = bc, df = df)
+  out <- list(d = d, dc = dc, bc = bc, se = se, df = df)
   return(out)
 }
 
@@ -155,7 +154,7 @@ crit_from_data_t1s <- function(m, s, n, se = NULL, df = NULL,
   d <- m / s
   dc <- tc * sqrt(1/n)
   bc <- tc * se
-  out <- list(d = d, dc = dc, bc = bc, df = df)
+  out <- list(d = d, dc = dc, bc = bc, se = se, df = df)
   return(out)
 }
 
@@ -201,7 +200,7 @@ crit_from_t_t2sp <- function(t = NULL, n, se = NULL, r12 = NULL, hypothesis, con
     bc <- tc * se
   }
   
-  out <- list(dz = dz, dzc = dzc, d = d, dc = dc, bc = bc, df = df)
[…]

### 2024-08-21 R/criticalvalue-methods.R

@@ -27,6 +27,11 @@ critical.htest <- function(x, type = NULL){
     )
   } else {
     D <- insight::get_data(x)
+    
+    if(is.null(D)){
+      stop("insight::get_data(x) returning NULL. Are you using the formula syntax (y ~ x)? This syntax is not supported yet")
+    }
+    
     conf.level <- attributes(x$conf.int)$conf.level
     
     # hypothesis <- ifelse(x$alternative == "two.sided", "2t", "1t")

### 2024-08-21 R/deprecated/criticalvalue-funs.R

@@ -1,174 +1,0 @@
-critical_ttest <- function(m1 = NULL, m2 = NULL,
-                           sd1 = NULL, sd2 = NULL,
-                           n1, n2,
-                           r12 = NULL,
-                           se = NULL,
-                           sed = NULL,
-                           t = NULL,
-                           type = "ts",
-                           var.equal = FALSE,
-                           denominator = "pooled",
-                           conf.level = 0.95,
-                           hypothesis = c("><", "<", ">")){
-  
-  # check arguments with multiple options
-  type <- match.arg(type, c("ts", "os", "ps"))
-  denominator <- match.arg(denominator, c("pooled", "differences"))
-  hypothesis <- match.arg(hypothesis)
-  
-  # get the alpha value for the critical value
-  alpha <- .get_alpha(conf.level, hypothesis)
-  
-  if(type == "ts"){ # two sample t test
-    # t, n1, n2 -> works but assume sd1 = sd2
-    df <- n1 + n2 - 2
-    if(is.null(t)){
-      if(is.null(m1) | is.null(m2) | is.null(sd1) | is.null(sd2)){
-        stop("When t is NULL m1, m2, sd1 and sd2 need to be provided!")
-      }
-      b <- m1 - m2
-      if(!var.equal){ # welch
-        se1 <- sd1 / sqrt(n1)
-        se2 <- sd2 / sqrt(n2)
-        se <- sqrt((se1^2 + se2^2)/2)
-        s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1))/(n1 + n2 - 2))
-        df <- se^4/(se1^4/(n1-1) + se2^4/(n2-1))
-      }else{ # standard
-        s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1)) / (n1 + n2 - 2))
-        se <- s * (1/n1 + 1/n2)
-      }
-    }else{
-      if(is.null(se)){
-        warning("bc cannot be calculated without se, returning NA")
-        s <- NA
-        se <- NA
-      }else{
-        s <- se / (1/n1 + 1/n2)
-      }
-    }
-  } else if(type == "os"){ # one sample
-    if(is.null(t)){
-      if(is.null(m1) | is.null(sd1)){
[…]

### 2024-08-21 R/deprecated/criticalvalue-methods.R

@@ -1,142 +1,0 @@
-critical <- function(x, ...){
-  UseMethod("critical")
-}
-
-#' @export
-critical.htest <- function(x, denominator = "pooled"){
-  denominator <- match.arg(denominator, c("pooled", "differences"))
-  # all implemented subtype of htest objects
-  mtds <- c("Two Sample", "One Sample", "correlation", "Paired")
-  
-  if(all(!grepl(paste0(mtds, collapse = "|"), x$method))){
-    stop(
-      sprintf("method %s of class %s not implemented yet!",
-              x$method, class(x))
-    )
-  } else{
-    D <- insight::get_data(x)
-    conf.level <- attributes(x$conf.int)$conf.level
-    
-    hypothesis <- switch(x$alternative,
-      "two.sided" = "><",
-      "greater" = ">",
-      "less" = "<"
-    )
-    
-    alpha <- .get_alpha(conf.level, hypothesis)
-    df <- x$parameter
-    tc <- qt(alpha, df)
-    t <- x$statistic
-    se <- x$stderr
-    b <- x$estimate
-    method <- x$method
-  }
-  
-  if(grepl("correlation", method)){
-    d <- b
-    dc <- bc <- tc / sqrt((df + tc^2))
-    class(x) <- c("critvalue", "ctest", "htest")
-  }else{ # if t-test
-    if(grepl("Two Sample", method)){
-      b <- b[1] - b[2] # x - y
-      n <- tapply(D$x, D$y, length)
-      if(grepl("Welch", method)){
-        ss <- tapply(D$x, D$y, sd)
-        s <- sqrt((ss[1]^2 + ss[2]^2)/2) # average sd
-      }else{
-        se_df <- sqrt(1/n[1] + 1/n[2]) # standard error without sd
-        s <- se / se_df
-      }
-    } else if(grepl("One Sample", method)){
-      n <- length(D$x)
-      se_df <- sqrt(1/n) # standard error without sd
-      s <- se / se_df
-    } else if(grepl("Paired", method)){
-      n <- tapply(D$x, D$y, length)
-      se_df <- sqrt(1/n[1]) # standard error without sd
-      ss <- tapply(D$x, D$y, sd)
-      if(denominator == "pooled"){
-        s <- sqrt((ss[1]^2 + ss[2]^2)/2)
-      } else if(denominator == "differences"){
-        s <- se / se_df
[…]

### 2024-08-21 R/deprecated/extra.R

@@ -1,72 +1,0 @@
-zscore <- function(x, s = 1, center = TRUE, scale = TRUE, na.rm = FALSE){
-  m <- mean(x, na.rm = na.rm)
-  sd <- sd(x, na.rm = na.rm)
-  if(center){
-    x <- x - m
-  }
-  if(scale){
-    x <- x / (sd*s)
-  }
-  return(x)
-}
-
-.weighted.sd <- function(x, w){
-  n <- length(x)
-  sqrt(sum(((x - weighted.mean(x, w))*w)^2)/(n - 1))
-}
-
-standardize <- function(x, stand.y = TRUE, stand.x = TRUE){
-  
-  Y <- model.frame(x)[, 1, drop = FALSE]
-  X <- model.frame(x)[, -1, drop = FALSE]
-  
-  numIDX <- sapply(X, is.numeric)
-  
-  if(stand.y){
-    Y <- scale(Y)
-  }
-  if(stand.x){
-    for(i in 1:ncol(X)){
-      if(is.numeric(X[, i])){
-        X[, i] <- scale(X[, i])
-      }
-    }
-  }
-  D <- cbind(Y, X)
-  update(x, data = D)
-  
-}
-
-old_standardize <- function(x, 
-                        stand.x = TRUE, 
-                        stand.y = FALSE,
-                        nsd = 1){
-  Y <- model.frame(x)[, 1]
-  X <- model.matrix(x)
-  if(exists("weights", x)){
-    w <- x$weights
-  }else{
-    w <- rep(1, nrow(x$model))
-  }
-  
-  sY <- .weighted.sd(Y, w)
-  sX <- apply(X, 2, .weighted.sd, w)
-  
-  sY <- sY * nsd
-  sX <- sX * nsd
-  
-  B <- coef(x)
-  SE <- summary(x)$coefficients[, 2]
-  
-  if(stand.x){
-    B <- B * sX
-    SE <- SE * sX 
-  }
-  if(stand.y){
-    B <- B/sY
-    SE <- SE/sY
-  }
-  x$standardized.coefficients <- B
-  x$standardized.se <- SE
-  return(x)
-}

### 2024-08-21 R/utils.R

@@ -48,3 +48,29 @@
   x <- abs(x)
   ifelse(hypothesis %in% c("two.sided", "greater"), x, -x) 
 }
+
+sim_t_data <- function(m1, m2 = NULL, 
+                       sd1, sd2 = NULL, 
+                       n1, n2 = NULL,
+                       type = c("1s", "2s", "2sp"),
+                       r12 = 0,
+                       empirical = TRUE){
+  type <- match.arg(type)
+  
+  if(type == "1s"){
+    y <- MASS::mvrnorm(n1, m1, Sigma = sd1^2, empirical = empirical)
+    data <- data.frame(y = y[, 1])
+  } else if(type == "2s"){
+    y1 <- MASS::mvrnorm(n1, m1, Sigma = sd1^2, empirical = empirical)
+    y2 <- MASS::mvrnorm(n2, m2, Sigma = sd2^2, empirical = empirical)
+    data <- data.frame(y = c(y1, y2), x = rep(0:1, c(n1, n2)))
+  } else{
+    R <- matrix(c(1, r12, r12, 1), nrow = 2)
+    S <- diag(c(sd1, sd2)) %*% R %*% diag(c(sd1, sd2))
+    Y <- MASS::mvrnorm(n1, c(m1, m2), Sigma = S, empirical = empirical)
+    data <- data.frame(y = c(Y[, 1], Y[, 2]), x = rep(0:1, each = n1))
+  }
+  
+  return(data)
+  
+}

### 2024-08-21 tests/testthat/test-utils2.R

@@ -2,28 +2,28 @@
 
 test_that("critical value for correlation works (t, two sided)", {
   n <- 40
-  rr <- critical_cortest(n, test = "t")
+  rr <- critical_cor(n = n, test = "t")
   t <- rr$rc / rr$se_rc
   # should be 0.05
-  pval <- p_value(t, rr$df, test = "t", hypothesis = "><")
+  pval <- .p_value(q = t, df = rr$df, test = "t", hypothesis = "2t")
   expect_equal(pval, 0.05)
 })
 
 test_that("critical value for correlation works (t, greater)", {
   n <- 40
-  rr <- critical_cortest(n, test = "t", hypothesis = ">")
+  rr <- critical_cor(n = n, test = "t", hypothesis = "greater")
   t <- rr$rc / rr$se_rc
   # should be 0.05
-  pval <- p_value(t, rr$df, test = "t", hypothesis = ">")
+  pval <- .p_value(t, rr$df, test = "t", hypothesis = "1t")
   expect_equal(pval, 0.05)
 })
 
 test_that("critical value for correlation works (t, less)", {
   n <- 40
-  rr <- critical_cortest(n, test = "t", hypothesis = "<")
+  rr <- critical_cor(n = n, test = "t", hypothesis = "less")
   t <- rr$rc / rr$se_rc
   # should be 0.05
-  pval <- p_value(t, rr$df, test = "t", hypothesis = "<")
+  pval <- .p_value(t, rr$df, test = "t", hypothesis = "1t")
   expect_equal(pval, 0.05)
 })
 
@@ -32,9 +32,9 @@ test_that("critical value for correlation works (t, less)", {
 test_that("critical value for correlation works (t, two sided)", {
   n <- 40
   conf.level <- 0.8
-  rr <- critical_cortest(n, test = "t", conf.level = conf.level)
+  rr <- critical_cor(n = n, test = "t", conf.level = conf.level)
   t <- rr$rc / rr$se_rc
-  pval <- p_value(t, rr$df, test = "t", hypothesis = "><", conf.level = conf.level)
+  pval <- .p_value(t, rr$df, test = "t", hypothesis = "2t", conf.level = conf.level)
   pval_expected <- 1 - conf.level
   expect_equal(pval, pval_expected)
 })
[…]

### 2024-08-21 README.Rmd
updating

@@
 
-# criticalvalue
+# criticalESvalue
 
@@
 
-You can install the development version of criticalvalue like so:
+You can install the development version of criticalESvalue like so:
 
@@
 # require(devtools)
-devtools::install_github("filippogambarota/criticalvalue")
+devtools::install_github("psicostat/criticalESvalue")
 ```

### 2024-08-21 R/criticalvalue-funs.R

@@ -22,7 +22,7 @@ crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
   }
   alpha <- .get_alpha(conf.level, hypothesis)
   df <- n1 + n2 - 2
-  tc <- abs(qt(alpha, df))
+  tc <- abs(stats::qt(alpha, df))
   if(is.null(t)){
     warning("When t is NULL, d cannot be computed, returning NA")
     d <- NA
@@ -83,7 +83,7 @@ crit_from_data_t2s <- function(m1, m2,
     if(is.null(se)) se <- s * sqrt(1/n1 + 1/n2)
     if(is.null(df)) df <- n1 + n2 - 2
   }
-  tc <- abs(qt(alpha, df))
+  tc <- abs(stats::qt(alpha, df))
   d <- b / s
   bc <- tc * se
   dc <- tc * sqrt(1/n1 + 1/n2)
@@ -111,7 +111,7 @@ crit_from_t_t1s <- function(t = NULL, n, se = NULL,
                             hypothesis){
   alpha <- .get_alpha(conf.level, hypothesis)
   df <- n - 1
-  tc <- abs(qt(alpha, df))
+  tc <- abs(stats::qt(alpha, df))
   dc <- tc * sqrt(1/n)
   if(is.null(t)){
     warning("When t is NULL, d cannot be computed, returning NA")
@@ -149,7 +149,7 @@ crit_from_data_t1s <- function(m, s, n, se = NULL, df = NULL,
                                conf.level, hypothesis){
   alpha <- .get_alpha(conf.level, hypothesis)
   if(is.null(df)) df <- n - 1
-  tc <- abs(qt(alpha, df))
+  tc <- abs(stats::qt(alpha, df))
   if(is.null(se)) se <- s / sqrt(n)
   d <- m / s
   dc <- tc * sqrt(1/n)
@@ -176,7 +176,7 @@ crit_from_data_t1s <- function(m, s, n, se = NULL, df = NULL,
 crit_from_t_t2sp <- function(t = NULL, n, se = NULL, r12 = NULL, hypothesis, conf.level){
   alpha <- .get_alpha(conf.level, hypothesis)
   df <- n - 1
-  tc <- abs(qt(alpha, df))
+  tc <- abs(stats::qt(alpha, df))
   
   # d on differences
   dzc <- tc * sqrt(1/n)
@@ -213,12 +213,13 @@ crit_from_t_t2sp <- function(t = NULL, n, se = NULL, r12 = NULL, hypothesis, con
[…]

### 2024-08-21 R/criticalvalue-methods.R

@@ -13,10 +13,9 @@ critical <- function(x, ...){
 #' Compute critical effect size values for t-test and correlation tests
 #' @description Compute critical effect size values for objects of classes htest computed with t.test or cor.test
 #' @param x an object of class htest (available for t.test, cor.test)
-#' @param type a value among "Two Sample", "One Sample", "correlation", "Paired"
 #' @return an object of class critvalue
 #' @export
-critical.htest <- function(x, type = NULL){
+critical.htest <- function(x){
   # all implemented subtype of htest objects
   mtds <- c("Two Sample", "One Sample", "correlation", "Paired")
   
@@ -39,7 +38,7 @@ critical.htest <- function(x, type = NULL){
     
     alpha <- .get_alpha(conf.level, hypothesis)
     df <- x$parameter
-    tc <- qt(alpha, df)
+    tc <- stats::qt(alpha, df)
     t <- x$statistic
     se <- x$stderr
     b <- x$estimate
@@ -48,7 +47,8 @@ critical.htest <- function(x, type = NULL){
   
   if(grepl("correlation", method)){
     n <- df + 2
-    cc <- critical_cor(n = n, conf.level = conf.level, hypothesis = hypothesis)
+    r <- x$estimate
+    cc <- critical_cor(r = r, n = n, conf.level = conf.level, hypothesis = hypothesis)
     dc <- bc <- cc$rc
     d <- b
     class(x) <- c("critvalue", "ctest", "htest")
@@ -56,7 +56,7 @@ critical.htest <- function(x, type = NULL){
     if(grepl("Two Sample", method)){
       n <- tapply(D$x, D$y, length)
       if(grepl("Welch", method)){
-        ss <- tapply(D$x, D$y, sd)
+        ss <- tapply(D$x, D$y, stats::sd)
         mm <- tapply(D$x, D$y, mean)
         tt <- critical_t2s(m1 = mm[1], m2 = mm[2],
                            sd1 = ss[1], sd2 = ss[2], 
@@ -69,11 +69,11 @@ critical.htest <- function(x, type = NULL){
                            conf.level = conf.level, hypothesis = hypothesis)
       }
     } else if(grepl("One Sample", method)){
-      n <- length(D$x)
+      n <- nrow(D)
[…]

### 2024-08-21 R/utils.R

@@ -7,9 +7,9 @@
   test <- match.arg(test)
   # two tails
   if(test == "t"){
-    p <- 2 * pt(-abs(q), df)
+    p <- 2 * stats::pt(-abs(q), df)
   }else{
-    p <- 2 * pnorm(-abs(q))
+    p <- 2 * stats::pnorm(-abs(q))
   }
   if(hypothesis == "1t"){
     p <- p / 2
@@ -49,7 +49,7 @@
   ifelse(hypothesis %in% c("two.sided", "greater"), x, -x) 
 }
 
-sim_t_data <- function(m1, m2 = NULL, 
+.sim_t_data <- function(m1, m2 = NULL, 
                        sd1, sd2 = NULL, 
                        n1, n2 = NULL,
                        type = c("1s", "2s", "2sp"),
@@ -73,4 +73,16 @@ sim_t_data <- function(m1, m2 = NULL,
   
   return(data)
   
+}
+
+.get_s <- function(sd1, sd2, n1 = NULL, n2 = NULL, var.equal = TRUE){
+  if(var.equal){
+    if(is.null(n1) | is.null(n2)){
+      stop("When pooled = TRUE, sample size need to be provided!")
+    }
+    s <- sqrt((sd1^2 * (n1 - 1) + sd2^2 * (n2 - 1)) / (n1 + n2 - 2))
+  } else{
+    s <- sqrt(0.5*(sd1^2 + sd2^2))
+  }
+  return(s)
 }

### 2024-08-21 deprecated/critical-value-class.R

@@ -1,132 +1,0 @@
-set.seed(2021)
-x <- rnorm(100)
-y <- rnorm(100)
-ht <- t.test(x, y)
-attributes(ht$conf.int)
-attr(ht, "class") <- c("feonyte", "htest")
-class(ht)
-
-print.feonyte <- function(x, ...){
-    NextMethod(x, ...)
-    alpha <- attr(x$conf.int, "conf.level")
-    cat("confidence level:", alpha, "\n\n")
-    invisible(NULL)
-}
-
-stats:::print.htest(ht)
-
-fit <- t.test(x, y, var.equal = TRUE)
-
-
-t.test()
-
-.all_null <- function(...){
-    all(sapply(list(...), is.null))
-}
-
-critical_t_test <- function(sd1 = NULL, 
-                            sd2 = NULL, 
-                            n1 = NULL,
-                            n2 = NULL,
-                            se = NULL, 
-                            df = NULL,
-                            alpha = 0.05,
-                            var.equal = FALSE){
-    if(is.null(se) & is.null(df)){
-        if(is.null(sd1) | is.null(sd2) | is.null(n1) | is.null(n2)){
-            stop("when se/df are NULL, sd1, sd2, n1 and n2 must be specified!")
-        }
-        stats <- t_test_se_df(n1, n2, sd1, sd2, var.equal)
-        se <- stats$se
-        df <- stats$df
-    }
-    tc <- critical_t(df, alpha/2)
-    critical_b <- se * tc
-    dc <- abs(critical_b)
-    list(critical = dc, se = se, df = df)
-}
-
-critical_cor_test <- function(n = NULL, alpha = 0.05){
-    df <- n - 2
-    tc <- critical_t(df, alpha/2)
-    abs(tc / sqrt(df + tc))
-}
-
-critical_t <- function(df, alpha = 0.05){
-    qt(alpha, df = df)
-}
-
-
-x <- rnorm(100)
-y <- rnorm(100)
-
-fit <- t.test(x, y)
-d <- critical_t_test(fit$stderr, fit$parameter)
-pt(d/fit$stderr, fit$parameter, lower.tail = FALSE)
-
-
-se <- sqrt(var(x)/100 + var(y)/100)
-fit$stderr
-
[…]

### 2024-08-21 deprecated/test.R

@@ -1,20 +1,0 @@
-t.test <- function(x, den = "pooled", ...){
-    res <- stats::t.test(x, ...)
-    alpha <- 1 - attributes(res$conf.int)$conf.level
-    nu <- res$parameter
-    res$tc <- qt(p = alpha/2, df = nu)
-    res$bc <- res$tc * res$stderr
-    class(res) <- c("critval", class(res))
-    return(res)
-}
-
-print.critval <- function(x, digits = getOption("digits"), prefix = "\t", ...){
-    NextMethod(x, ...)
-    critical_b <- sprintf("critical |b| = %.3f", abs(x$bc))
-    cat(critical_b, "\n")
-    invisible(NULL)
-}
-
-x <- t.test(rnorm(100), rnorm(100), paired = TRUE)
-
-x

### 2024-08-21 deprecated/test2.R

@@ -1,30 +1,0 @@
-get_htest <- function(x){
-    patterns <- c("ttest" = "t-test", 
-                  "cortest" = "correlation")
-    if(!"htest" %in% class(x)){
-        stop("x need to be of class htest!")
-    }else{
-        p <- sapply(patterns, function(p) grepl(p, x$method))
-        names(patterns)[p]
-    }
-}
-
-add_critical <- function(x){
-    htest <- get_htest(x)
-    class(x) <- c("critval", class(x))
-    return(x)
-}
-
-print.critval <- function(x, ...){
-    NextMethod(x, ...)
-    critical_b <- sprintf("critical b = %.3f", 0.45)
-    critical_d <- sprintf("critical d = %.3f", 0.1)
-    cat(critical_b)
-    invisible(NULL)
-}
-
-x <- t.test(rnorm(100))
-y <- cor.test(rnorm(100), rnorm(100))
-
-x <- add_critical(x)
-x

### 2024-08-21 example.R

@@ -1,38 +1,0 @@
-
-devtools::load_all()
-library(criticalvalue)
-
-# two sample t.test (var.equal = TRUE)
-
-x <- rnorm(500, 0.5, 1)
-y <- rnorm(500, 0, 1)
-
-tt <- t.test(x, y)
-tt
-cc = cor.test(x,y)
-fit <- lm(y ~ x)
-
-critical(fit)
-critical(fit, standardize = TRUE)
-critical(tt)
-critical(cc)
-
-n = 20
-yi = rnorm(n, .5, .2)
-vi = runif(n,.05,.10)
-xi = rnorm(n, 2, 1)
-library(metafor)
-fitrma = rma(yi=yi,vi=vi,mods=~xi)
-summary(fitrma)
-critical(fitrma)
-
-
-
-niter = 1e4
-n = 500
-ps = rep(NA,niter)
-
-pwr::pwr.t.test(n=500,power=0.8)
-metafor:::.cmicalc()
-
-

### 2024-08-21 tests/testthat.R

@@ -7,6 +7,6 @@
 # * https://testthat.r-lib.org/articles/special-files.html
 
 library(testthat)
-library(criticalvalue)
+library(criticalESvalue)
 
-test_check("criticalvalue")
+test_check("criticalESvalue")

### 2024-08-21 tests/testthat/test-utils2.R

@@ -1,302 +1,320 @@
-# cor test, t, conf.level = 0.95 -------------------------------
+# COR TEST ----------------------------------------------------------------
 
-test_that("critical value for correlation works (t, two sided)", {
-  n <- 40
-  rr <- critical_cor(n = n, test = "t")
-  t <- rr$rc / rr$se_rc
-  # should be 0.05
-  pval <- .p_value(q = t, df = rr$df, test = "t", hypothesis = "2t")
-  expect_equal(pval, 0.05)
-})
+## is the critical value correct? ---------------------------------------
 
-test_that("critical value for correlation works (t, greater)", {
-  n <- 40
-  rr <- critical_cor(n = n, test = "t", hypothesis = "greater")
-  t <- rr$rc / rr$se_rc
-  # should be 0.05
-  pval <- .p_value(t, rr$df, test = "t", hypothesis = "1t")
-  expect_equal(pval, 0.05)
-})
+# the p value calculated with the critical value should be equal to alpha
 
-test_that("critical value for correlation works (t, less)", {
-  n <- 40
-  rr <- critical_cor(n = n, test = "t", hypothesis = "less")
-  t <- rr$rc / rr$se_rc
-  # should be 0.05
-  pval <- .p_value(t, rr$df, test = "t", hypothesis = "1t")
-  expect_equal(pval, 0.05)
-})
-
-# cor test, t, conf.level = 0.80 -------------------------------
-
-test_that("critical value for correlation works (t, two sided)", {
-  n <- 40
-  conf.level <- 0.8
-  rr <- critical_cor(n = n, test = "t", conf.level = conf.level)
-  t <- rr$rc / rr$se_rc
-  pval <- .p_value(t, rr$df, test = "t", hypothesis = "2t", conf.level = conf.level)
-  pval_expected <- 1 - conf.level
-  expect_equal(pval, pval_expected)
-})
-
-test_that("critical value for correlation works (t, greater)", {
-  n <- 40
-  conf.level <- 0.8
-  rr <- critical_cor(n = n, test = "t", hypothesis = "greater", conf.level = conf.level)
-  t <- rr$rc / rr$se_rc
-  pval_expected <- 1 - conf.level
[…]

### 2024-08-22 README.Rmd
updating

@@
 The goal of criticalvalue is to ...
+<!-- TODO aggiungi qualche riga qui -->
 
@@
 ``` r
-# require(devtools)
-devtools::install_github("psicostat/criticalESvalue")
+# require(remotes)
+remotes::install_github("psicostat/criticalESvalue")
 ```
 
-## Installation (temporary)
+## Examples
 
-The package is not production-ready thus to test the feature we suggest to clone or download the repository, open the R Project and load the functions with `devtools::load_all()`.
-
-## Example from fitted objects
-
-This is a basic example which shows you how to solve a common problem:
+```{r}
+# loading the package
+library(criticalESvalue)
+```
 
@@
 ```{r example}
-# library(criticalvalue) # when the package will be ready
-devtools::load_all() # loading all functions for testing
-
 # t-test (welch)
@@
 summary(fit)
-
-# linear model (standardized)
-fit <- lm(y ~ x + q + z, data = dat)
-fit <- critical(fit, standardize = TRUE)
-fit
-summary(fit)
 ```
@@
 ```{r}
-require(metafor)
+library(metafor)
 dat <- escalc(measure="RR", ai=tpos, bi=tneg, ci=cpos, di=cneg, data=dat.bcg)

### 2024-08-22 paper/paper.Rmd
updating

@@
       - "Writing - Review & Editing"
+      - "Methodology"
+      - "Software"
   - name          : "Filippo Gambarota"
@@
       - "Methodology"
+      - "Software"
   - name          : "Enrico Toffalini"
@@
       - "Writing - Review & Editing"
+      - "Software"
       - "Supervision"
@@
 These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest.
-Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty on the expected effect size.
-To assist researchers in calculating critical values, we have developed an R package that allows to report the critical effect size values of model parameters.
+Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty about the expected effect size.
+To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
 The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
@@
 keywords          : critical effect size values, statistical significance, fucntions implementation 
-wordcount         : "X"
+wordcount         : "3912"
 
@@
 
-In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance.
-The "critical effect size values" refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample [@lakens2022sample].
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's $r = -0.44$ and $r = 0.44$.
-As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
+The "critical effect size value(s)" refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample [@lakens2022sample].
+To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's $r = -0.44$ and $r = 0.44$.
+As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
+This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
 
-Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null].
+Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null].
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size ($n$) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
+These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
-For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, that should subsequently be validated by "replication science".
+For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, which should subsequently be validated by "replication science".
 
-In psychology, power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings.
+In psychology, statistical power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings.
 In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology.
@@
 On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts [@wilson2020science].
+However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts [@wilson2020science].
 Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.
+In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
 
-There are two clear benefits of reporting the critical effect size for a corresponding test.
-First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
-If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant.
+There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
+First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
+Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
 This will focus the attention of researchers on the difference between statistical significance and practical significance.
@@
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
-Subsequently, they compute the critical values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
+Subsequently, they compute the critical effect size values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
 This indicates that any estimated |Cohen's $d| < 0.75$ will certainly fail to reach statistical significance.
-By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
+By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
 In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained on a small sample.
-Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained with a small sample.
+Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
 
 In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes.
-However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
+With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
+However, they legitimately  opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
 In this context, the critical effect size values for significance are determined to be correlations of $r = \pm 0.03$.
@@
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
+While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
 
-To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research.
-The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
+To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
+The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
 The first case pertains to "Study 5" conducted by @mccrea2008self.
@@
 With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, $t(26) = 1.87, p (one-tailed) < .05$.
+The resulting effect was statistically significant, $t(26) = 1.87$, $p$ (one-tailed) $< .05$.
 Based on the reported data, Cohen's $d = 0.736$, indicating a considerable effect size.
@@
 
-In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package "criticalESvalues" accessible at: https://github.com/filippogambarota/criticalvalue, and are elaborated upon in the subsequent section.
+In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+These formulas have been incorporated into R functions of the package "criticalESvalue" accessible at: https://github.com/psicostat/criticalESvalue, and are elaborated upon in the subsequent section.
 
@@
 Where $t_c$ is the critical $t$ value calculated using a certain $\alpha$ value (e.g., $0.05$) and $n_1 + n_2 - 2$ degrees of freedom.
-Then the $b_c$ is simply divided by $s$ obtaining the standardized critical value.
+Then the $b_c$ is simply divided by $s$ obtaining the standardized critical effect size value.
 For the two-sample case ($s = s_p$ assuming equal variances) Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.
@@
 Where $\rho$ is the correlation between the two paired samples.
-Even if the hypothesis testing is computed using $s_D$, we reported the critical value both using the pooled standard deviation and the standard deviation of the differences.
+Even if the hypothesis testing is computed using $s_D$, we reported the critical effect size value both using the pooled standard deviation and the standard deviation of the differences.
 Beyond this difference, the equations are still the same as the previous example.
@@
 Then a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept.
-The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical value as reported in Equation \@ref(eq:eq-critical-lm).
+The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical effect size value as reported in Equation \@ref(eq:eq-critical-lm).
 
@@
 Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests.
-The calculation of critical values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm).
+The calculation of critical effect size values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm).
 The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.
@@
 Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models.
-In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.
+In fact, we simply need to multiply the critical effect size value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.
 
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalues" in R.
-The complete package can be accessed at: https://github.com/filippogambarota/criticalvalue.
-Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size.
-In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalue" in R.
+The complete package can be accessed at: https://github.com/psicostat/criticalESvalue.
+Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
+It is important to note that, in general, depending on the researcher's hypotheses, the package allows for the calculation of either two critical effect size values for two-tailed tests or a single critical effect size value for one-tailed tests, for which the direction must be specified.
+In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients.
 
@@
 ```{r, download-package}
-# devtools::install_github("filippogambarota/criticalvalue")
-library(criticalvalue)
+# require(remotes)
+# remotes::install_github("psicostat/criticalESvalue")
+library(criticalESvalue)
 ```
@@
 
-We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception.
-In this case, it can be easily done with the same procedure using the 't.test' function:
+We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the 't.test' function:
 
@@
 tt <- t.test(Holz$t01_visperc[Holz$female == 1], 
-             Holz$t01_visperc[Holz$female == 2])
+             Holz$t01_visperc[Holz$female == 2], var.equal = T)
 critical(tt)
@@
 The output gives a wide range of values: the Cohen's $d$ calculated on the data ($d$), the critical Cohen's $d$ ($dc$), the numerator of the formula for the critical Cohen's $d$ ($bc$), the Cohen's $d$ adjusted for small samples ($g$) and the critical Cohen's $d$ adjusted for small samples ($gc$).
+The variance is set to be equal, but if that is not the case of your data, you can also run Welch two sample $t$-test and obtain the critical effect size values for that.
 
-In the next example we will show the use of the package's function `critical_cor` to calculate the critical value for a correlation in a prospective framework.
+In the next example we will show the use of the package's function `critical_cor` to calculate the critical effect size values for a correlation in a prospective framework.
 
@@
 The direction of the hypothesis and the test to apply, either $t$-test or $z$-test, should be specified.
-The output will return the critical correlation value, the degrees of freedom and the type of test used.
+The output will return the critical correlation value(s), the degrees of freedom and the type of test used.
 
@@
 This is not intended to replace other strategies aimed at enhancing the NHST approach to inference.
-Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the a priori planning for statistical power are valuable in their own right.
+Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the \textit{a priori} planning for statistical power are valuable in their own right.
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).
-Interestingly, the critical value can be retrospectively applied even to already published studies.
+Interestingly, critical effect size values can be retrospectively applied even to already published studies.
 This possible application facilitates potential reframing of the original interpretations.
-Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.
+Serving as a tool for retrospective analysis, critical effect size values may enable a reconsideration of the relevance of previously reported findings.
 
-An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
-The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
+An advantage of reporting critical effect size values is that they can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
+The critical effect size value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
 For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects.
 To what extent can the reported results be interpreted, precisely?
-Knowing the critical value provides us with a clear benchmark.
+Knowing the critical effect size value provides us with a clear benchmark.
 Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this.
@@
 If in comparing two groups, such as controls versus treatment, any Cohen's $d > 0.07$ would reach significance, is statistical significance enough to signal a "successful" treatment?
-Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical value certainly prompts some appropriate interpretive caution.
+Maybe yes, even if effects are tiny [e.g., @funder2019evaluating], but knowing the critical effect size value certainly prompts some appropriate interpretive caution.
 
-Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
+Reporting the critical effect size value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
@@
 
-Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical value(s) can also serve an educational purpose.
+Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical effect size values can also serve an educational purpose.
 It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size.
-By highlighting a critical value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant.
-Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical value(s) may serves as a reminder that any observed effect larger than a trivially small value will likely achieve significance.
+By highlighting a critical effect size value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant.
+Conversely, in studies with exceedingly large samples and in many meta-analyses, the critical effect size value(s) may serve as a reminder that any observed effect larger than a trivially small value will likely achieve significance.
 This emphasizes that the mere attainment of statistical significance in a test is not particularly surprising, especially in non-experimental studies.
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical significance values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients.
-This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
+Hence, we chose to expand the application of computing critical effect size values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients.
+This serves as a first step in computing critical effect size values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
 A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance.
@@
 In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward.
-However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical values for the obtained sample can help provide context for interpretation.
-Critical values can be computed and interpreted even retrospectively or for studies that have already been published.
+However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical effect size values for the obtained sample can help provide context for interpretation.
+Critical effect size values can be computed and interpreted even retrospectively or for studies that have already been published.
 
-In conclusion, the reporting of the critical value in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings.
+In conclusion, reporting critical effect size values in empirical articles serves as a valuable addition to researchers' toolkit, aimed at augmenting transparency and facilitating the interpretability of their findings.
 While not designed to supplant existing practices, it provides a useful aid in interpreting newly presented and previously published results, thus advancing the understanding of research outcomes.

### 2024-08-22 paper/paper.tex
updating

@@
 \fi
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: X}
+\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: 3912}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 The authors made the following contributions.
-Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Conceptualization, Writing - Review \& Editing, Supervision; Livio Finos: Conceptualization, Writing - Review \& Editing, Supervision; Psicostat: Conceptualization, Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
+Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Software, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Conceptualization, Writing - Review \& Editing, Supervision; Livio Finos: Conceptualization, Writing - Review \& Editing, Supervision; Psicostat: Conceptualization, Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
 
@@
 
-In the present paper, we suggest that researchers systematically report the critical effect size values when they make inference using statistical significance.
-The ``critical effect size values'' refers to the smallest statistically significant effect sizes, which depends on the test performed and the characteristics of the sample (Lakens, 2022).
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
-As we will argue, the critical values can serve as a useful complement to aid in the interpretation of study results, communicate their relevance, help assessing their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size or when statistical power may be difficult to determine or deviates from an optimal level.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
+The ``critical effect size value(s)'' refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample (Lakens, 2022).
+To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
+As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
+This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
 
-Despite longstanding criticisms, the Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
+Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
 Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size (\(n\)) a priori based on theory-informed target effects, and reporting of effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
+These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
 Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
 For instance, Wilson et al.
-(2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, that should subsequently be validated by ``replication science''.
+(2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, which should subsequently be validated by ``replication science''.
 
-In psychology, power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings.
+In psychology, statistical power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings.
 In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size (Lakens, 2022), and this scenario is arguably widespread in psychology.
@@
 On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they just reflect minor procedural artifacts (Wilson et al., 2020).
+However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts (Wilson et al., 2020).
 Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical values can help put into perspective and convey the relevance of the results.
+In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
 
-There are two clear benefits of reporting the critical effect size for a corresponding test.
-First, when sample sizes are smaller, the critical values inform readers about whether the effect sizes that could lead to a rejection of the null hypothesis are in line with realistic expectations.
-If the sample size is small, and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realized they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting the critical effect sizes focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect sizes will make clear that trivially small effect sizes will be statistically significant.
+There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
+First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
+Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
 This will focus the attention of researchers on the difference between statistical significance and practical significance.
@@
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
-Subsequently, they compute the critical values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
+Subsequently, they compute the critical effect size values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
 This indicates that any estimated \textbar Cohen's \(d| < 0.75\) will certainly fail to reach statistical significance.
-By reporting these critical values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
+By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
 In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will probably be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained on a small sample.
-Knowing from the critical values that a specific result must necessarily be associated with an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained with a small sample.
+Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
 
 In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead just focus on effect sizes.
-However, they opt to legitimately use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
+With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
+However, they legitimately opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
 In this context, the critical effect size values for significance are determined to be correlations of \(r = \pm 0.03\).
@@
 It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, preliminarily signaling the critical value level can serve as a clear warning that the significance filter might be somehow permissive, thus urging interpretive caution.
+While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis as well, reporting the critical values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
 
-To provide more tangible illustrations of the practical application of critical values, let us examine two real-world instances drawn from published research.
-The first one features a small sample, and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
+To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
+The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
 The first case pertains to ``Study 5'' conducted by McCrea (2008).
@@
 With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, \(t(26) = 1.87, p (one-tailed) < .05\).
+The resulting effect was statistically significant, \(t(26) = 1.87\), \(p\) (one-tailed) \(< .05\).
 Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size.
@@
 
-\hypertarget{how-to-compute-critical-effect-size-values}{%
-\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}}
+\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}
 
-In this section, we provide guidance and formulas for computing critical values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package ``criticalESvalues'' accessible at: \url{https://github.com/filippogambarota/criticalvalue}, and are elaborated upon in the subsequent section.
+In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+These formulas have been incorporated into R functions of the package ``criticalESvalue'' accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 
-\hypertarget{t-test}{%
-\subsection{t-test}\label{t-test}}
+\subsection{t-test}\label{t-test}
 
@@
 
-\hypertarget{two-sample-t-test}{%
-\subsubsection{two-sample t-test}\label{two-sample-t-test}}
+\subsubsection{two-sample t-test}\label{two-sample-t-test}
 
@@
 Where \(t_c\) is the critical \(t\) value calculated using a certain \(\alpha\) value (e.g., \(0.05\)) and \(n_1 + n_2 - 2\) degrees of freedom.
-Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical value.
+Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical effect size value.
 For the two-sample case (\(s = s_p\) assuming equal variances) Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.
@@
 
-\hypertarget{one-sample-t-test}{%
-\subsubsection{one sample t-test}\label{one-sample-t-test}}
+\subsubsection{one sample t-test}\label{one-sample-t-test}
 
@@
 
-\hypertarget{paired-sample-t-test}{%
-\subsubsection{paired sample t-test}\label{paired-sample-t-test}}
+\subsubsection{paired sample t-test}\label{paired-sample-t-test}
 
@@
 Where \(\rho\) is the correlation between the two paired samples.
-Even if the hypothesis testing is computed using \(s_D\), we reported the critical value both using the pooled standard deviation and the standard deviation of the differences.
+Even if the hypothesis testing is computed using \(s_D\), we reported the critical effect size value both using the pooled standard deviation and the standard deviation of the differences.
 Beyond this difference, the equations are still the same as the previous example.
 
-\hypertarget{hedgess-correction}{%
-\subsubsection{Hedges's correction}\label{hedgess-correction}}
+\subsubsection{Hedges's correction}\label{hedgess-correction}
 
@@
 
-\hypertarget{correlation-test}{%
-\subsection{Correlation Test}\label{correlation-test}}
+\subsection{Correlation Test}\label{correlation-test}
 
@@
 
-\hypertarget{linear-regression}{%
-\subsection{Linear Regression}\label{linear-regression}}
+\subsection{Linear Regression}\label{linear-regression}
 
@@
 Then a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept.
-The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical value as reported in Equation \eqref{eq:eq-critical-lm}.
+The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical effect size value as reported in Equation \eqref{eq:eq-critical-lm}.
 
@@
 
-\hypertarget{meta-analysis}{%
-\subsubsection{Meta-analysis}\label{meta-analysis}}
+\subsubsection{Meta-analysis}\label{meta-analysis}
 
@@
 Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests.
-The calculation of critical values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}.
+The calculation of critical effect size values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}.
 The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.
 
-\hypertarget{other-models}{%
-\subsubsection{Other models}\label{other-models}}
+\subsubsection{Other models}\label{other-models}
 
 Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models.
-In fact, we simply need to multiply the critical value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.
+In fact, we simply need to multiply the critical effect size value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.
 
-\hypertarget{examples-in-r}{%
-\section{Examples in R}\label{examples-in-r}}
+\section{Examples in R}\label{examples-in-r}
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalues'' in R.
-The complete package can be accessed at: \url{https://github.com/filippogambarota/criticalvalue}.
-Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical value for a correlation from sample size.
-In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients, both from data and from sample size.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalue'' in R.
+The complete package can be accessed at: \url{https://github.com/psicostat/criticalESvalue}.
+Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
+It is important to note that, in general, depending on the researcher's hypotheses, the package allows for the calculation of either two critical effect size values for two-tailed tests or a single critical effect size value for one-tailed tests, for which the direction must be specified.
+In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# devtools::install\_github("filippogambarota/criticalvalue")}
-\FunctionTok{library}\NormalTok{(criticalvalue)}
+\CommentTok{\# require(remotes)}
+\CommentTok{\# remotes::install\_github("psicostat/criticalESvalue")}
+\FunctionTok{library}\NormalTok{(criticalESvalue)}
 \end{Highlighting}
@@
 
-We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception.
-In this case, it can be easily done with the same procedure using the `t.test' function:
+We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the `t.test' function:
 
@@
 \NormalTok{tt }\OtherTok{\textless{}{-}} \FunctionTok{t.test}\NormalTok{(Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{1}\NormalTok{], }
-\NormalTok{             Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{2}\NormalTok{])}
+\NormalTok{             Holz}\SpecialCharTok{$}\NormalTok{t01\_visperc[Holz}\SpecialCharTok{$}\NormalTok{female }\SpecialCharTok{==} \DecValTok{2}\NormalTok{], }\AttributeTok{var.equal =}\NormalTok{ T)}
 \FunctionTok{critical}\NormalTok{(tt)}
@@
 #> 
-#>  Welch Two Sample t-test
+#>  Two Sample t-test
 #> 
 #> data:  Holz$t01_visperc[Holz$female == 1] and Holz$t01_visperc[Holz$female == 2]
-#> t = 1.4095, df = 298.9, p-value = 0.1597
+#> t = 1.4062, df = 299, p-value = 0.1607
 #> alternative hypothesis: true difference in means is not equal to 0
 #> 95 percent confidence interval:
-#>  -0.06419229  0.38823118
+#>  -0.06472157  0.38876046
 #> sample estimates:
@@
 #> |== Effect Size and Critical Value ==| 
-#> d = 0.1623665 dc = ± 0.2269609 bc = ± 0.2262117 
-#> g = 0.1619587 gc = ± 0.2263909
+#> d = 0.1621764 dc = ± 0.2269606 bc = ± 0.226741 
+#> g = 0.1617692 gc = ± 0.2263908
 \end{verbatim}
@@
 The output gives a wide range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).
+The variance is set to be equal, but if that is not the case of your data, you can also run Welch two sample \(t\)-test and obtain the critical effect size values for that.
 
-In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical value for a correlation in a prospective framework.
+In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical effect size values for a correlation in a prospective framework.
 
@@
 #> 
+#> $rzc
+#> [1] 0.2596036
+#> 
 #> $df
 #> [1] 58
+#> 
+#> $se_r
+#> numeric(0)
+#> 
+#> $se_rc
+#> [1] 0.1270027
+#> 
+#> $se_rzc
+#> [1] 0.1324532
 #> 
@@
 The direction of the hypothesis and the test to apply, either \(t\)-test or \(z\)-test, should be specified.
-The output will return the critical correlation value, the degrees of freedom and the type of test used.
+The output will return the critical correlation value(s), the degrees of freedom and the type of test used.
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 This is not intended to replace other strategies aimed at enhancing the NHST approach to inference.
-Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the a priori planning for statistical power are valuable in their own right.
+Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the \textit{a priori} planning for statistical power are valuable in their own right.
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).
-Interestingly, the critical value can be retrospectively applied even to already published studies.
+Interestingly, critical effect size values can be retrospectively applied even to already published studies.
 This possible application facilitates potential reframing of the original interpretations.
-Serving as a tool for retrospective analysis, the critical value may enable a reconsideration of the relevance of previously reported findings.
+Serving as a tool for retrospective analysis, critical effect size values may enable a reconsideration of the relevance of previously reported findings.
 
-An advantage of reporting the critical value is that it can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
-The critical value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
+An advantage of reporting critical effect size values is that they can be precisely computed in any scenario, without requiring assumptions about the expected effect size, as is the case with power calculations.
+The critical effect size value represents a directly interpretable benchmark that is especially useful in situations where statistical power is below the desired level and researchers are left otherwise uncertain about how to proceed with the interpretation of a study findings.
 For example, let us say that we read a published article reporting some effects as statistically significant, while others as not: we suspect that the study may be underpowered, but we are widely uncertain about the magnitude of possible true effects.
 To what extent can the reported results be interpreted, precisely?
-Knowing the critical value provides us with a clear benchmark.
+Knowing the critical effect size value provides us with a clear benchmark.
 Conversely, let us say that an effect achieves significance in a very large sample: researchers tend to draw substantive conclusions based on this.
@@
 If in comparing two groups, such as controls versus treatment, any Cohen's \(d > 0.07\) would reach significance, is statistical significance enough to signal a ``successful'' treatment?
-Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical value certainly prompts some appropriate interpretive caution.
+Maybe yes, even if effects are tiny (e.g., Funder \& Ozer, 2019), but knowing the critical effect size value certainly prompts some appropriate interpretive caution.
 
-Reporting the critical value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
+Reporting the critical effect size value(s) can also be an efficient way to allow researchers to evaluate which findings are statistically significant.
 For example, in a correlation table researchers customarily add an asterisk to all statistically significant correlations.
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical significance values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients.
-This serves as a first step in computing critical values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
+Hence, we chose to expand the application of computing critical effect size values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients.
+This serves as a first step in computing critical effect size values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
 A prerequisite is that researchers must be able to identify what parameters in their statistical models reflect the effect sizes of interest, and that they are able to assess their relevance.
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-altoe2020enhancing}{}}%
+\bibitem[\citeproctext]{ref-altoe2020enhancing}
 Altoè, G., Bertoldo, G., Zandonella Callegher, C., Toffalini, E., Calcagnı̀, A., Finos, L., \& Pastore, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
+\bibitem[\citeproctext]{ref-cohen1994}
 Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.
 
-\leavevmode\vadjust pre{\hypertarget{ref-open2015estimating}{}}%
+\bibitem[\citeproctext]{ref-open2015estimating}
 Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
 
-\leavevmode\vadjust pre{\hypertarget{ref-funder2019evaluating}{}}%
+\bibitem[\citeproctext]{ref-funder2019evaluating}
 Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: Sense and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008-zp}{}}%
+\bibitem[\citeproctext]{ref-Gelman2008-zp}
 Gelman, A. (2008). Scaling regression inputs by dividing by two standard deviations. \emph{Statistics in Medicine}, \emph{27}, 2865--2873. \url{https://doi.org/10.1002/sim.3107}
 
-\leavevmode\vadjust pre{\hypertarget{ref-gigerenzer2004null}{}}%
+\bibitem[\citeproctext]{ref-gigerenzer2004null}
 Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
+\bibitem[\citeproctext]{ref-Hedges1981-za}
 Hedges, L.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hedges1984estimation}{}}%
+\bibitem[\citeproctext]{ref-hedges1984estimation}
 Hedges, L.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kramer2014experimental}{}}%
+\bibitem[\citeproctext]{ref-kramer2014experimental}
 Kramer, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
+\bibitem[\citeproctext]{ref-krueger2001null}
 Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
+\bibitem[\citeproctext]{ref-lakens2013calculating}
 Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
+\bibitem[\citeproctext]{ref-lakens2022sample}
 Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.
 
-\leavevmode\vadjust pre{\hypertarget{ref-mccrea2008self}{}}%
+\bibitem[\citeproctext]{ref-mccrea2008self}
 McCrea, S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
+\bibitem[\citeproctext]{ref-morris2002combining}
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
 
-\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
+\bibitem[\citeproctext]{ref-orben_crud_2020}
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
 
-\leavevmode\vadjust pre{\hypertarget{ref-szucs2017empirical}{}}%
+\bibitem[\citeproctext]{ref-szucs2017empirical}
 Szucs, D., \& Ioannidis, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-transue2019apa}{}}%
+\bibitem[\citeproctext]{ref-transue2019apa}
 Transue, B. (2019). \emph{APA style 7th edition}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-wilson2020science}{}}%
+\bibitem[\citeproctext]{ref-wilson2020science}
 Wilson, B. M., Harris, C. R., \& Wixted, J. T. (2020). Science is not a signal detection problem. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(11), 5559--5567.

### 2024-08-22 R/criticalvalue-funs.R

@@ -10,9 +10,8 @@
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #' @param var.equal a logical variable indicating whether to treat the two variances as being equal.
 #'
-#' @export
-#' @return the output returns a "d" which is the cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the "bc" is the numerator of the formula from which the d is calculated and "df" are the degrees of freedom.
-#'
+#' @return the output returns a `d` which is the Cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the `bc` is the numerator of the formula from which the d is calculated, `df` are the degrees of freedom and `se` is the standard error.
+
 crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
                             conf.level, 
                             hypothesis,
@@ -56,10 +55,8 @@ crit_from_t_t2s <- function(t = NULL, n1, n2, se = NULL,
 #' @param df degrees of freedom.
 #' @param var.equal a logical variable indicating whether to treat the two variances as being equal.
 #'
-#' @return the output returns a "d" which is the cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the "bc" is the numerator of the formula from which the d is calculated and "df" are the degrees of freedom.
-#' @export
+#' @return the output returns a `d` which is the Cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample size, the `bc` is the numerator of the formula from which the d is calculated, `df` are the degrees of freedom, and `se` is the standard error.
 #'
-#' @examples
 crit_from_data_t2s <- function(m1, m2, 
                                sd1, sd2, 
                                n1, n2, 
@@ -102,10 +99,8 @@ crit_from_data_t2s <- function(m1, m2,
 #' @param conf.level confidence level of the interval.
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #'
-#' @return the output returns a "d" which is the cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the "bc" is the numerator of the formula from which the d is calculated and "df" are the degrees of freedom.
-#' @export
[…]

### 2024-08-22 R/criticalvalue-methods.R

@@ -1,4 +1,3 @@
-
 #' Compute critical effect size values for a range of different objects
 #' @description Compute critical effect size values for objects of classes htest (Student's t-test, correlation test), lm, or rma
 #' @param x an object of class htest (available for t.test, cor.test), lm (a linear regression model), or rma (a meta-analytic linear model fitted with the "metafor" package)
@@ -13,9 +12,10 @@ critical <- function(x, ...){
 #' Compute critical effect size values for t-test and correlation tests
 #' @description Compute critical effect size values for objects of classes htest computed with t.test or cor.test
 #' @param x an object of class htest (available for t.test, cor.test)
+#' @param ... Additional arguments (currently unused).
 #' @return an object of class critvalue
 #' @export
-critical.htest <- function(x){
+critical.htest <- function(x, ...){
   # all implemented subtype of htest objects
   mtds <- c("Two Sample", "One Sample", "correlation", "Paired")
   
@@ -108,9 +108,10 @@ critical.htest <- function(x){
 #' @description Compute critical effect size values for linear model coefficients of objects of class lm
 #' @param x an object of class lm
 #' @param conf.level the confidence interval level, needed to compute the smallest significant coefficient (default is 0.95, equaling a critical alpha = 0.05)
+#' @param ... Additional arguments (currently unused).
 #' @return an object of class critvalue
 #' @export
-critical.lm <- function(x, conf.level = 0.95){
+critical.lm <- function(x, conf.level = 0.95, ...){
   
   # always two.sided
   hypothesis <- "two.sided"
@@ -133,9 +134,10 @@ critical.lm <- function(x, conf.level = 0.95){
 #' @description Compute critical effect size values for coefficients of objects of class rma (fitted with the "metafor" package)
 #' @param x an object of class rma
 #' @param conf.level the confidence interval level, needed to compute the smallest significant coefficient (default is 0.95, equaling a critical alpha = 0.05)
+#' @param ... Additional arguments (currently unused).
 #' @return an object of class critvalue
 #' @export
-critical.rma <- function(x, conf.level = 0.95){
+critical.rma <- function(x, conf.level = 0.95, ...){
   if(inherits(x, "rma.uni")){
     hypothesis <- "two.sided"
     se <- x$se
[…]

### 2024-08-22 paper/paper.Rmd
updating

@@
 In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package "criticalESvalue" accessible at: https://github.com/psicostat/criticalESvalue, and are elaborated upon in the subsequent section.
+These formulas have been incorporated into R functions of the package `criticalESvalue` accessible at: [https://github.com/psicostat/criticalESvalue](https://github.com/psicostat/criticalESvalue), and are elaborated upon in the subsequent section.
 

### 2024-08-22 paper/paper.tex
updating

@@
 In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package ``criticalESvalue'' accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
+These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 

### 2024-08-22 R/criticalvalue-funs.R
adding examples

@@ -271,7 +271,7 @@ crit_from_data_t2sp <- function(m1, m2 = NULL, # if m2 = NULL, m1 is the mean of
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #' @param conf.level the confidence level to set the confidence interval, default is set to 0.95.
 #'
-#' @return the output returns, based on the inserted values, mean, variance, confidence interval and hypothesis or t-value, numerosity, standard error, confidence interval and hypothesis. 
+#' @return the output returns a `d` which is the Cohen's d, the critical d which is the minimum value for which to get a significant result with a given sample, the `bc` is the numerator of the formula from which the d is calculated, `df` are the degrees of freedom and `se` is the standard error, then it also gives the `g` and `gc` which are respectively `d` and `dc` with Hedfer's Correction for small samples.
 #' @export
 #'
 #' @examples
@@ -285,8 +285,6 @@ crit_from_data_t2sp <- function(m1, m2 = NULL, # if m2 = NULL, m1 is the mean of
 #' t <- m / se
 #' critical_t1s(t = t, n = n, se = se) # se only required for calculating bc
 
-# TODO manca documentare g e gc che sono la versione con small sample correction (Hedges'g) del valore critico e dell'effect size calcolato
-# TODO manca esempio (qui faccio io un esempio, prova a fare gli altri)
 critical_t1s <- function(m = NULL, s = NULL, t = NULL,
                          n, se = NULL,
                          hypothesis = c("two.sided", "greater", "less"),
@@ -329,12 +327,24 @@ critical_t1s <- function(m = NULL, s = NULL, t = NULL,
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #' @param conf.level the confidence level to set the confidence interval, default is set to 0.95.
 #'
-#' @return the output returns, based on the inserted values, means of the two groups, standard deviations, numerosity of the two groups, confidence interval, hypothesis, standard error, degrees of freedom and the set parameter for var.equal or the t-value, numerosity of the two groups, standard error, confidence interval, hypothesis and the set parameter for var.equal.
[…]

### 2024-08-23 R/criticalvalue-funs.R
m

@@ -404,14 +404,14 @@ critical_t2s <- function(m1 = NULL, m2 = NULL, t = NULL,
 #'
 #' @examples
 #' # critical value from summary statistics
-#' m1 <- 0.5
-#' m2 <- 1.0
-#' sd1 <- 1
-#' sd2 <- 1.5
+#' m1 <- 10
+#' m2 <- 15
+#' sd1 <- 5
+#' sd2 <- 4.25
 #' n <- 30
 #' critical_t2sp(m1 = m1, m2 = m2, sd1 = sd1, sd2 = sd2, n = n)
 #' # critical value from the t statistic
-#' se <- sqrt(sd1^2 / n1 + sd2^2 / n2)
+#' se <- sqrt((sd1^2 + sd2^2) / n)
 #' t <- (m1 - m2) / se
 #' critical_t2sp(t = t, n = n, se = se) # se only required for calculating bc
 # TODO controllare formula dell'esemptio from t test

### 2024-09-16 paper/paper.Rmd
updating how to compute es section

@@
     \makeatother
+    \usepackage{xcolor}
 ---
@@
 knitr::opts_chunk$set(echo = TRUE,
-                      comment = "#>")
+                      comment = "#>",
+                      size = "footnotesize")
+
+def.chunk.hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def.chunk.hook(x, options)
+  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
+})
 ```
@@
 For example, in the two sample t-test or the single mean for a one sample t-test.
-The denominator is the standard error of the numerator.
+The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
 
@@
 
-### two-sample t-test
-
-For the two sample t-test the critical value ($b_c$) is calculated using Equation \@ref(eq:eq-tw-sp).
-Where $t_c$ is the critical $t$ value calculated using a certain $\alpha$ value (e.g., $0.05$) and $n_1 + n_2 - 2$ degrees of freedom.
-Then the $b_c$ is simply divided by $s$ obtaining the standardized critical effect size value.
-For the two-sample case ($s = s_p$ assuming equal variances) Equation \@ref(eq:eq-s-p-twsmp) reported the pooled standard deviation.
+The critical $t_c$ is the test statistic associated with a p-value equal or lower to $\alpha$.
+Substituting $t_c$ into Equation \@ref(eq:eq-gen-t) we realize that there is a $b_c$ that with a certain $SE_b$ produce the critical test statistics.
+The general form of the critical effect size $d_c$ is defined in Equation \@ref(eq:eq-gen-crit).
+The $b_c$ is derived solving Equation \@ref(eq:eq-gen-t) for $b$.
 
 \begin{equation}
-    (\#eq:eq-tw-sp)
-    b_c = t_c \times \text{SE}_{b}
+\begin{gathered}
+    (\#eq:eq-gen-crit)
+    d_c = \frac{b_c}{s} \\
+    b_c = t_c \times SE_b
+\end{gathered}
 \end{equation}
+
+NOT SURE TO KEEP THIS!
+
+Another way of conceptualizing the critical effect size is by removing the sample size from the $t$ statistics (see Equation \@ref(eq:eq-gen-t)).
+In fact, $SE_b$ can be generally defined as $\frac{s}{\sqrt{N}}$ where $s$ is the denominator in Equation \@ref(eq:eq-gen-d) and $n$ is the sample size.
+Thus, $d_c = t_c \times \frac{1}{\sqrt{n}}$ ($\frac{1}{\sqrt{n}}$ because the standard deviation of a standardized effect size is one by definition) is equivalent to \@ref(eq:eq-gen-crit).
+The only caveat is that calculating $d$ or $d_c$ in this way assume that the standard deviation used to calculate the test statistics and the effect size is the same.
+While this is true most of the time, as explained in the paired t-test section, there are more than one methods to calculate the (critical) effect size using different standard deviations.
+Furthemore, this method assume that the two groups (in case of a two-sample t-test) have the same size.
+
+### One sample t-test
+
+The easiest case is the one-sample t-test.
+$b$ is the sample mean and $s$ is the sample standard deviation and $n$ is the sample size.
+$SE_b = \frac{s}{\sqrt{n}}$ and the $t_c$ is based on $n - 1$ degrees of freedom.
+Thus $b_c = t_c \times SE_b$ and $d_c = \frac{b_c}{s}$.
+Equivalently, $d_c = t_c \times \frac{1}{\sqrt{n}}$
+
+### Two-sample t-test
+
+We can apply the one-sample approach to the two-sample case.
+When assuming homogeneity of the variances between the two groups we have $n_1 + n_2 - 2$ degrees of freedom to calculate $t_c$ and $s$ is the pooled standard deviation ($s_p$) calculated using Equation \@ref(eq:eq-s-p-twsmp).
+The $s_p$ can be intepreted as the square root of the weighted average of the samples standard deviations.
+The effect size $d$ is calculated as $\frac{b}{s_p}$ and the critical effect size can be calculated as $d_c = \frac{b_c}{s_p}$ with $b_c = t_c \times SE_b$.
+$SE_b$ is calculated as $s_p \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}$.
+As for the one-sample case, we can calculate the critical effect size from the t statistics as $d_c = t_c \times \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}$.
 
@@
 
-When relaxing the assumption of equal variances (i.e., Welch's test) $s$ is simply the square root of the average between the two variances.
+When relaxing the assumption of equal variances (i.e., using the Welch's t-test) $s$ is calculated as the square root of the average between the two samples variances, $SE_b$ is calculated as $\sqrt{\frac{s^2_{x_1}}{n_1} + \frac{s^2_{x_2}}{n_2}}$.
+The degress of freedom for $t_c$ are calculated in a more complex way using the Welch-Satterthwaite equation [@Welch1947-wh; @Satterthwaite1946-aq].
+For the Welch's t-test we can calculate the critical effect size removing the sample size from $t_c$ (as done above) only when $n_1 = n_2$.
 
-### one sample t-test
-
-For the one sample case, the equations are the same. The only difference is that $b$ is the sample mean and $s$ is the sample standard deviation.
-
-### paired sample t-test
+### Paired sample t-test
 
 For the paired sample case the situation is less straightforward.
-The reason is that the $t$ statistic is computed using $b$ being the average of the paired differences between the two (paired) samples and $s$ being the standard deviation/error of the differences.
-While for hypothesis testing and power analysis it is common to calculate the $d$ dividing by the standard deviation of the differences [@lakens2013calculating], this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation [@morris2002combining].
-Practically, Equation \@ref(eq:eq-sp-sd) depicts the relationship between the pooled standard deviation and the standard deviation of the differences.
+The reason is that the $t$ statistic is computed using $b$ being the average of the paired differences between the two (paired) samples and $s$ being the standard deviation of the differences.
+Using these values, we can just compute the critical effects size as for the one-sample case, in fact a paired t-test is a one-sample test on the vector of paired differences.
+Test statistics and effect size based on the standard deviation of differences (the so called $d_z$) are used for hypothesis testing and power analysis [@lakens2013calculating].
+The problem is that the $d_z$ cannot be directly compared to the effect size calculated using the pooled standard deviation [@morris2002combining].
+If the correlation between the two paired samples is known there is a direct relationship between the pooled standard deviation $s_p$ and the standard deviation of differences $s_D$ as reported in Equation \@ref(eq:eq-sp-sd)
 
@@
 
-Where $\rho$ is the correlation between the two paired samples.
 Even if the hypothesis testing is computed using $s_D$, we reported the critical effect size value both using the pooled standard deviation and the standard deviation of the differences.
-Beyond this difference, the equations are still the same as the previous example.
 
@@
 
-The effect size calculated as in the previous step is known to be inflated thus we calculated also the corrected version usually known as Hedges's $g$ [@Hedges1981-za].
+The effect size calculated as in the previous step is known to be inflated especially for small samples.
+For this reason, there is a corrected version of the effect size called Hedges's $g$ [@Hedges1981-za, @Viechtbauer2007-tq].
+The Hedges's $g$ can be calculated for all the t-test scenarios.
+The Hegdes's correction is implemented in Equation \@ref(eq:eq-g-correction) where $\Gamma$ is the gamma function and $m$ the degrees of freedom.
+
+\begin{equation}
+(\#eq:eq-g-correction)
+c(m) = \frac{\Gamma\left(\frac{m}{2}\right)}{\sqrt{\frac{m}{2}} \Gamma\left(\frac{m-1}{2}\right)} \approx 1 - \frac{3}{4m - 1}
+\end{equation}
+
+The correction is applied as $g = c(m) \times d$. The hypothesis testing is not taking into account the Hedges's correction thus the critical $g$ will be different from the critical $d$.
 
@@
 
-For the Pearson's correlation test, the critical value is computed using Equation \@ref(eq:eq-cor-c) where $t_c$ is still the critical value for the $t$ distribution and $n$ is the sample size.
+Hypothesis testing for the Pearson's correlation coefficient is usually done using the t statistics or using a z statistics.
+The general approach presented for the t-test is still valid.
+The only difference is that the correlation is already a standardized effect size thus there is no need of the standardization term $s$.
+Equation \@ref(eq:eq-gen-t) can be used substituing $b$ with $r$ (the sample correlation coefficient) as shown in Equation \@ref(eq:eq-rc).
+The standard error is calculated as $SE_r = \sqrt{\frac{1 - r^2}{n - 2}}$ [@Bowley1928-yt].
 
 \begin{equation}
-    (\#eq:eq-cor-c)
+(\#eq:eq-rc)
+    \begin{gathered}
+        t = \frac{r}{SE_r} = r \sqrt{\frac{n - 2}{1 - r^2}} \\
         r_c = \frac{t_c}{\sqrt{n - 2 + t_c^2}}
+    \end{gathered}
 \end{equation}
 
-Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's $z$ transformation.
-Equation \@ref(eq:eq-f-rc) shows how to calculate the critical $F(r_c)$ value where $z_c$ is the critical value of the standard normal distribution with a chosen $\alpha$ level.
-Then the $F(r_c)$ can be transformed back into a correlation coefficient using Equation \@ref(eq:eq-bk-r).
+Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's $z$ transformation [@Fisher1921-qz; @Fisher1915-ks] reported in Equation \@ref(eq:eq-z-transf) and a $z$ test statistics.
+
+\begin{equation}
+    (\#eq:eq-z-transf)
+    \begin{gathered}
+    r_z = \frac{1}{2} \ln \left(\frac{1 + r}{1 - r} \right) =                   \text{artanh}(r) \\
+    r = \frac{\exp(2r_z - 1)}{\exp(2r_z + 1)} = \text{tanh}(r_z)
+    \end{gathered}
+\end{equation}
+
+We can still use Equation \@ref(eq:eq-gen-t) substituting $b$ with $r_z$.
+The standard error of the $r_z$ correlation is calculated as $SE_{r_z} = \frac{1}{\sqrt{n - 3}}$.
+Equation \@ref(eq:eq-f-rc) shows how to calculate the critical $r_{z_c}$ value where $z_c$ is the critical value of the standard normal distribution with a certain $\alpha$ level.
+Finally, $r_{z_c}$ can be transformed back from the Fisher's $z$ transformation using Equation \@ref(eq:eq-z-transf)
 
@@
     \begin{gathered}
-        F(r_c) = \frac{z_c}{\sqrt{n - 3}}
+        z = \frac{r_z}{SE_z} = r_z \sqrt{n - 3} \\
+        r_{z_c} = \frac{z_c}{\sqrt{n - 3}}
     \end{gathered}
@@
 
-\begin{equation}
-    (\#eq:eq-bk-r)
-    r_c = \frac{\exp(2z_c)-1}{\exp(2z_c)+1} = \operatorname{tanh}(z_c)
-\end{equation}
-
-<!-- %% see better this part -->
-
 ## Linear Regression
 
-Hypothesis testing on single coefficients in linear regression (using the \texttt{lm} function in R) is performed using the Wald test.
-This test is basically a t-test where the $t$ value is calculated dividing the regression parameter $\beta_0, \dots, \beta_j$ by the standard error.
-Then a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept.
-The $t$ value is calculated as reported in Equation \@ref(eq:eq-wald-lm) and the critical effect size value as reported in Equation \@ref(eq:eq-critical-lm).
+Hypothesis testing on regression coefficients (e.g., using the \texttt{lm} function in R) is performed using the so-called Wald test.
+The test can be considered a one-sample t-test where the $t$ value is calculated dividing the regression parameter $\beta_0, \dots, \beta_j$ by its standard error $SE_{\beta_j}$.
+The critical $t$ value is calculated from a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept.
+We can substitute $b$ with $\beta_j$ and $SE_{b}$ with $SE_{\beta_j}$ in Equation \@ref(eq:eq-gen-crit) to find the critical unstandardized regression coefficient ($\beta_{c_j} = t_c \times SE_{\beta_j}$).
+There is no an unique way to calculate the standardized version of the critical regression coefficient.
+The main reason is that in regression modelling we can include categorical and numerical independent variables with main effects and interactions.
+For example, @Gross2023-fl proposed a way to calculate a generalized Cohen's $d$ from linear regression with or without the presence of other variables dividing the estimated parameter with the residual standard deviation.
+@Gelman2008-zp proposed to standardize using two standard deviations because when both numeric and categorical variables are included the coefficients are not on the same metric.
+Another complication is that in some cases, researchers standardize the dependent variable, the independent variables or both.
+In general, also the unstandardized critical regression coefficient can be usually easily intepreted.
+For example, assuming we are regressing some scores on a cognitive tests using the age of participant.
+The critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
 
-\begin{equation}
-(\#eq:eq-wald-lm)
-    t = \frac{\beta_j}{SE_{\beta_j}}
-\end{equation}
-
-\begin{equation}
-(\#eq:eq-critical-lm)
-    \beta_{jc} = t_c SE_{\beta_j}
-\end{equation}
-
-<!-- maybe worth adding more about standardization -->
-
-In this case $\beta_{jc}$ is the unstandardized critical regression coefficient (as $b$ in the $t$-test section).
-We implemented also the possibility to use standardized regression coefficients.
-We can identify two types of standardization, namely full and partial.
-The full standardization involves both the response variable and the predictors.
-Regression coefficients are interpreted as the increase in standard deviations of the response variable for an increase of one standard deviation in the predictor, keeping fixed all other predictors.
-The partial standardization involves only the predictors and regression coefficients are interpreted as the increase in the (raw) response variable for a increase in one standard deviation of the predictor.
-By default, standardizing means dividing a quantity by the standard deviation and eventually centering (i.e., subtracting the mean).
-Gelman [@Gelman2008-zp] suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric.
-Equation \@ref(eq:eq-critical-lm) can be used but to calculate the critical value for the standardized regression coefficients the result need to be interpreted according to the chosen approach.
-
-<!-- %% -->
-
-### Meta-analysis
-
-<!-- %% da rivedere -->
+## Meta-analysis
 
@@
 From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances.
-Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests.
-The calculation of critical effect size values is the same as reported in Equations \@ref(eq:eq-wald-lm) and \@ref(eq:eq-critical-lm).
-The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.
+Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests [@Borenstein2009-mo].
+For this reason, we can apply the same equations used for the linear regression.
+The main difference and advantage is that the meta-analysis model usually works directly with standardized effect size measures and regression coefficients are already standardized.
 
-### Other models
+## Other models
 
-Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models.
-In fact, we simply need to multiply the critical effect size value of the chosen distribution (e.g., Student $t$ or Standard Normal) by the standard error of the regression coefficient.
+Despite we discussed only linear models, the same approach could be applied to other types of models such as generalized linear models.
+In fact, we simply need to multiply the critical value of the chosen distribution (e.g., $t$ or $z$) by the standard error of the regression coefficient.
+For example, in a logistic regression with a binary predictor the estimated regression coefficient is the (log) odds ratio.
+We can obtain the critical odds ratio by multiplying the critical $z$ statistics (the default in R with the `glm` function) with the standard error of the regression coefficient.
 

### 2024-09-16 paper/paper.tex
updating how to compute es section

@@
 \makeatother
+\usepackage{xcolor}
+\newcommand{\brev}{\color{red}}
+\newcommand{\erev}{\color{black}}
 \ifLuaTeX
@@
 
-\subsubsection{two-sample t-test}\label{two-sample-t-test}
+\color{red}
 
-For the two sample t-test the critical value (\(b_c\)) is calculated using Equation \eqref{eq:eq-tw-sp}.
-Where \(t_c\) is the critical \(t\) value calculated using a certain \(\alpha\) value (e.g., \(0.05\)) and \(n_1 + n_2 - 2\) degrees of freedom.
-Then the \(b_c\) is simply divided by \(s\) obtaining the standardized critical effect size value.
-For the two-sample case (\(s = s_p\) assuming equal variances) Equation \eqref{eq:eq-s-p-twsmp} reported the pooled standard deviation.
+The critical \(t_c\) is the test statistic associated with a p-value equal or lower to \(\alpha\).
+Substituting \(t_c\) into Equation \eqref{eq:eq-gen-t} we realize that there is a \(b_c\) that with a certain \(SE_b\) produce the critical test statistics.
+The general form of the critical effect size \(d_c\) is defined in Equation \eqref{eq:eq-gen-crit}.
+The \(b_c\) is derived solving Equation \eqref{eq:eq-gen-t} for \(b\).
 
 \begin{equation}
-    \label{eq:eq-tw-sp}
-    b_c = t_c \times \text{SE}_{b}
+\begin{gathered}
+    \label{eq:eq-gen-crit}
+    d_c = \frac{b_c}{s} \\
+    b_c = t_c \times SE_b
+\end{gathered}
 \end{equation}
+
+NOT SURE TO KEEP THIS!
+
+Another way of conceptualizing the critical effect size is by removing the sample size from the \(t\) statistics (see Equation \eqref{eq:eq-gen-t}).
+In fact, \(SE_b\) can be generally defined as \(\frac{s}{\sqrt{N}}\) where \(s\) is the denominator in Equation \eqref{eq:eq-gen-d} and \(n\) is the sample size.
+Thus, \(d_c = t_c \times \frac{1}{\sqrt{n}}\) (\(\frac{1}{\sqrt{n}}\) because the standard deviation of a standardized effect size is one by definition) is equivalent to \eqref{eq:eq-gen-crit}.
+The only caveat is that calculating \(d\) or \(d_c\) in this way assume that the standard deviation used to calculate the test statistics and the effect size is the same.
+While this is true most of the time, as explained in the paired t-test section, there are more than one methods to calculate the (critical) effect size using different standard deviations.
+Furthemore, this method assume that the two groups (in case of a two-sample t-test) have the same size.
+
+\color{black}
+
+\color{red}
+
+\subsubsection{One sample t-test}\label{one-sample-t-test}
+
+The easiest case is the one-sample t-test.
+\(b\) is the sample mean and \(s\) is the sample standard deviation and \(n\) is the sample size.
+\(SE_b = \frac{s}{\sqrt{n}}\) and the \(t_c\) is based on \(n - 1\) degrees of freedom.
+Thus \(b_c = t_c \times SE_b\) and \(d_c = \frac{b_c}{s}\).
+Equivalently, \(d_c = t_c \times \frac{1}{\sqrt{n}}\)
+
+\color{black}
+
+\subsubsection{Two-sample t-test}\label{two-sample-t-test}
+
+We can apply the one-sample approach to the two-sample case.
+When assuming homogeneity of the variances between the two groups we have \(n_1 + n_2 - 2\) degrees of freedom to calculate \(t_c\) and \(s\) is the pooled standard deviation (\(s_p\)) calculated using Equation \eqref{eq:eq-s-p-twsmp}.
+The \(s_p\) can be intepreted as the square root of the weighted average of the samples standard deviations.
+The effect size \(d\) is calculated as \(\frac{b}{s_p}\) and the critical effect size can be calculated as \(d_c = \frac{b_c}{s_p}\) with \(b_c = t_c \times SE_b\).
+\(SE_b\) is calculated as \(s_p \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}\).
+As for the one-sample case, we can calculate the critical effect size from the t statistics as \(d_c = t_c \times \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}\).
 
@@
 
-When relaxing the assumption of equal variances (i.e., Welch's test) \(s\) is simply the square root of the average between the two variances.
+When relaxing the assumption of equal variances (i.e., using the Welch's t-test) \(s\) is calculated as the square root of the average between the two samples variances, \(SE_b\) is calculated as \(\sqrt{\frac{s^2_{x_1}}{n_1} + \frac{s^2_{x_2}}{n_2}}\).
+The degress of freedom for \(t_c\) are calculated in a more complex way using the Welch-Satterthwaite equation (Satterthwaite, 1946; Welch, 1947).
+For the Welch's t-test we can calculate the critical effect size removing the sample size from \(t_c\) (as done above) only when \(n_1 = n_2\).
 
-\subsubsection{one sample t-test}\label{one-sample-t-test}
-
-For the one sample case, the equations are the same. The only difference is that \(b\) is the sample mean and \(s\) is the sample standard deviation.
-
-\subsubsection{paired sample t-test}\label{paired-sample-t-test}
+\subsubsection{Paired sample t-test}\label{paired-sample-t-test}
 
 For the paired sample case the situation is less straightforward.
-The reason is that the \(t\) statistic is computed using \(b\) being the average of the paired differences between the two (paired) samples and \(s\) being the standard deviation/error of the differences.
-While for hypothesis testing and power analysis it is common to calculate the \(d\) dividing by the standard deviation of the differences (Lakens, 2013), this effect size cannot be directly compared to the effect size used in the independent samples t-test, which divides by the pooled standard deviation (Morris \& DeShon, 2002).
-Practically, Equation \eqref{eq:eq-sp-sd} depicts the relationship between the pooled standard deviation and the standard deviation of the differences.
+The reason is that the \(t\) statistic is computed using \(b\) being the average of the paired differences between the two (paired) samples and \(s\) being the standard deviation of the differences.
+Using these values, we can just compute the critical effects size as for the one-sample case, in fact a paired t-test is a one-sample test on the vector of paired differences.
+Test statistics and effect size based on the standard deviation of differences (the so called \(d_z\)) are used for hypothesis testing and power analysis (Lakens, 2013).
+The problem is that the \(d_z\) cannot be directly compared to the effect size calculated using the pooled standard deviation (Morris \& DeShon, 2002).
+If the correlation between the two paired samples is known there is a direct relationship between the pooled standard deviation \(s_p\) and the standard deviation of differences \(s_D\) as reported in Equation \eqref{eq:eq-sp-sd}
 
@@
 
-Where \(\rho\) is the correlation between the two paired samples.
 Even if the hypothesis testing is computed using \(s_D\), we reported the critical effect size value both using the pooled standard deviation and the standard deviation of the differences.
-Beyond this difference, the equations are still the same as the previous example.
 
@@
 
-The effect size calculated as in the previous step is known to be inflated thus we calculated also the corrected version usually known as Hedges's \(g\) (Hedges, 1981).
+The effect size calculated as in the previous step is known to be inflated especially for small samples.
+For this reason, there is a corrected version of the effect size called Hedges's \(g\) Viechtbauer (2007).
+The Hedges's \(g\) can be calculated for all the t-test scenarios.
+The Hegdes's correction is implemented in Equation \eqref{eq:eq-g-correction} where \(\Gamma\) is the gamma function and \(m\) the degrees of freedom.
+
+\begin{equation}
+\label{eq:eq-g-correction}
+c(m) = \frac{\Gamma\left(\frac{m}{2}\right)}{\sqrt{\frac{m}{2}} \Gamma\left(\frac{m-1}{2}\right)} \approx 1 - \frac{3}{4m - 1}
+\end{equation}
+
+The correction is applied as \(g = c(m) \times d\). The hypothesis testing is not taking into account the Hedges's correction thus the critical \(g\) will be different from the critical \(d\).
 
@@
 
-For the Pearson's correlation test, the critical value is computed using Equation \eqref{eq:eq-cor-c} where \(t_c\) is still the critical value for the \(t\) distribution and \(n\) is the sample size.
+Hypothesis testing for the Pearson's correlation coefficient is usually done using the t statistics or using a z statistics.
+The general approach presented for the t-test is still valid.
+The only difference is that the correlation is already a standardized effect size thus there is no need of the standardization term \(s\).
+Equation \eqref{eq:eq-gen-t} can be used substituing \(b\) with \(r\) (the sample correlation coefficient) as shown in Equation \eqref{eq:eq-rc}.
+The standard error is calculated as \(SE_r = \sqrt{\frac{1 - r^2}{n - 2}}\) (Bowley, 1928).
 
 \begin{equation}
-    \label{eq:eq-cor-c}
+\label{eq:eq-rc}
+    \begin{gathered}
+        t = \frac{r}{SE_r} = r \sqrt{\frac{n - 2}{1 - r^2}} \\
         r_c = \frac{t_c}{\sqrt{n - 2 + t_c^2}}
+    \end{gathered}
 \end{equation}
 
-Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's \(z\) transformation.
-Equation \eqref{eq:eq-f-rc} shows how to calculate the critical \(F(r_c)\) value where \(z_c\) is the critical value of the standard normal distribution with a chosen \(\alpha\) level.
-Then the \(F(r_c)\) can be transformed back into a correlation coefficient using Equation \eqref{eq:eq-bk-r}.
+Another approach for hypothesis testing of the Pearson's correlation coefficient is using the Fisher's \(z\) transformation (Fisher, 1915, 1921) reported in Equation \eqref{eq:eq-z-transf} and a \(z\) test statistics.
+
+\begin{equation}
+    \label{eq:eq-z-transf}
+    \begin{gathered}
+    r_z = \frac{1}{2} \ln \left(\frac{1 + r}{1 - r} \right) =                   \text{artanh}(r) \\
+    r = \frac{\exp(2r_z - 1)}{\exp(2r_z + 1)} = \text{tanh}(r_z)
+    \end{gathered}
+\end{equation}
+
+We can still use Equation \eqref{eq:eq-gen-t} substituting \(b\) with \(r_z\).
+The standard error of the \(r_z\) correlation is calculated as \(SE_{r_z} = \frac{1}{\sqrt{n - 3}}\).
+Equation \eqref{eq:eq-f-rc} shows how to calculate the critical \(r_{z_c}\) value where \(z_c\) is the critical value of the standard normal distribution with a certain \(\alpha\) level.
+Finally, \(r_{z_c}\) can be transformed back from the Fisher's \(z\) transformation using Equation \eqref{eq:eq-z-transf}
 
@@
     \begin{gathered}
-        F(r_c) = \frac{z_c}{\sqrt{n - 3}}
+        z = \frac{r_z}{SE_z} = r_z \sqrt{n - 3} \\
+        r_{z_c} = \frac{z_c}{\sqrt{n - 3}}
     \end{gathered}
-\end{equation}
-
-\begin{equation}
-    \label{eq:eq-bk-r}
-    r_c = \frac{\exp(2z_c)-1}{\exp(2z_c)+1} = \operatorname{tanh}(z_c)
 \end{equation}
@@
 
-Hypothesis testing on single coefficients in linear regression (using the \texttt{lm} function in R) is performed using the Wald test.
-This test is basically a t-test where the \(t\) value is calculated dividing the regression parameter \(\beta_0, \dots, \beta_j\) by the standard error.
-Then a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept.
-The \(t\) value is calculated as reported in Equation \eqref{eq:eq-wald-lm} and the critical effect size value as reported in Equation \eqref{eq:eq-critical-lm}.
+Hypothesis testing on regression coefficients (e.g., using the \texttt{lm} function in R) is performed using the so-called Wald test.
+The test can be considered a one-sample t-test where the \(t\) value is calculated dividing the regression parameter \(\beta_0, \dots, \beta_j\) by its standard error \(SE_{\beta_j}\).
+The critical \(t\) value is calculated from a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept.
+We can substitute \(b\) with \(\beta_j\) and \(SE_{b}\) with \(SE_{\beta_j}\) in Equation \eqref{eq:eq-gen-crit} to find the critical unstandardized regression coefficient (\(\beta_{c_j} = t_c \times SE_{\beta_j}\)).
+There is no an unique way to calculate the standardized version of the critical regression coefficient.
+The main reason is that in regression modelling we can include categorical and numerical independent variables with main effects and interactions.
+For example, Groß and Möller (2023) proposed a way to calculate a generalized Cohen's \(d\) from linear regression with or without the presence of other variables dividing the estimated parameter with the residual standard deviation.
+Gelman (2008) proposed to standardize using two standard deviations because when both numeric and categorical variables are included the coefficients are not on the same metric.
+Another complication is that in some cases, researchers standardize the dependent variable, the independent variables or both.
+In general, also the unstandardized critical regression coefficient can be usually easily intepreted.
+For example, assuming we are regressing some scores on a cognitive tests using the age of participant.
+The critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
 
-\begin{equation}
-\label{eq:eq-wald-lm}
-    t = \frac{\beta_j}{SE_{\beta_j}}
-\end{equation}
-
-\begin{equation}
-\label{eq:eq-critical-lm}
-    \beta_{jc} = t_c SE_{\beta_j}
-\end{equation}
-
-In this case \(\beta_{jc}\) is the unstandardized critical regression coefficient (as \(b\) in the \(t\)-test section).
-We implemented also the possibility to use standardized regression coefficients.
-We can identify two types of standardization, namely full and partial.
-The full standardization involves both the response variable and the predictors.
-Regression coefficients are interpreted as the increase in standard deviations of the response variable for an increase of one standard deviation in the predictor, keeping fixed all other predictors.
-The partial standardization involves only the predictors and regression coefficients are interpreted as the increase in the (raw) response variable for a increase in one standard deviation of the predictor.
-By default, standardizing means dividing a quantity by the standard deviation and eventually centering (i.e., subtracting the mean).
-Gelman (Gelman, 2008) suggested to standardize using two standard deviations because when numeric and binary variables are included the coefficients are not on the same metric.
-Equation \eqref{eq:eq-critical-lm} can be used but to calculate the critical value for the standardized regression coefficients the result need to be interpreted according to the chosen approach.
-
-\subsubsection{Meta-analysis}\label{meta-analysis}
+\subsection{Meta-analysis}\label{meta-analysis}
 
@@
 From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances.
-Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests.
-The calculation of critical effect size values is the same as reported in Equations \eqref{eq:eq-wald-lm} and \eqref{eq:eq-critical-lm}.
-The only difference is that the standard error of the meta-analysis parameters included both the within-studies variances and eventually the between-studies heterogeneity.
+Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests (Borenstein et al., 2009).
+For this reason, we can apply the same equations used for the linear regression.
+The main difference and advantage is that the meta-analysis model usually works directly with standardized effect size measures and regression coefficients are already standardized.
 
-\subsubsection{Other models}\label{other-models}
+\subsection{Other models}\label{other-models}
 
-Despite we discussed only linear models, the same approach could be theoretically applied to other types of models such as generalized linear models.
-In fact, we simply need to multiply the critical effect size value of the chosen distribution (e.g., Student \(t\) or Standard Normal) by the standard error of the regression coefficient.
+Despite we discussed only linear models, the same approach could be applied to other types of models such as generalized linear models.
+In fact, we simply need to multiply the critical value of the chosen distribution (e.g., \(t\) or \(z\)) by the standard error of the regression coefficient.
+For example, in a logistic regression with a binary predictor the estimated regression coefficient is the (log) odds ratio.
+We can obtain the critical odds ratio by multiplying the critical \(z\) statistics (the default in R with the \texttt{glm} function) with the standard error of the regression coefficient.
 
@@
 
+\footnotesize
+
 \begin{Shaded}
@@
 
+\normalsize
+
 For our examples on real data we used from the package `psych' the dataset ``holzinger.swineford'' which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects.
@@
 For simplicity we decided to rename it with a shorter name.
+
+\footnotesize
 
@@
 
+\normalsize
+
 We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the `t.test' function:
+
+\footnotesize
 
@@
 
+\normalsize
+
 The output gives a wide range of values: the Cohen's \(d\) calculated on the data (\(d\)), the critical Cohen's \(d\) (\(dc\)), the numerator of the formula for the critical Cohen's \(d\) (\(bc\)), the Cohen's \(d\) adjusted for small samples (\(g\)) and the critical Cohen's \(d\) adjusted for small samples (\(gc\)).
@@
 In the next example we will show the use of the package's function \texttt{critical\_cor} to calculate the critical effect size values for a correlation in a prospective framework.
+
+\footnotesize
 
@@
 \end{verbatim}
+
+\normalsize
 
@@
 
+\bibitem[\citeproctext]{ref-Borenstein2009-mo}
+Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
+
+\bibitem[\citeproctext]{ref-Bowley1928-yt}
+Bowley, A. L. (1928). The standard deviation of the correlation coefficient. \emph{Journal of the American Statistical Association}, \emph{23}, 31. \url{https://doi.org/10.2307/2277400}
+
 \bibitem[\citeproctext]{ref-cohen1994}
@@
 Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
+
+\bibitem[\citeproctext]{ref-Fisher1915-ks}
+Fisher, R.
+A.
+(1915).
+Frequency distribution of the values of the correlation coefficient in samples from an indefinitely large population.
+\emph{Biometrika}, \emph{10}, 507.
+\url{https://doi.org/10.2307/2331838}
+
+\bibitem[\citeproctext]{ref-Fisher1921-qz}
+Fisher, R. A. (1921). On the "probable error" of a coefficient of correlation deduced from a small sample. \emph{Metron}.
 
@@
 Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.
+
+\bibitem[\citeproctext]{ref-Gross2023-fl}
+Groß, J., \& Möller, A.
+(2023).
+A note on cohen's d from a partitioned linear regression model.
+\emph{Journal of Statistical Theory and Practice}, \emph{17}.
+\url{https://doi.org/10.1007/s42519-023-00323-w}
 
@@
 
+\bibitem[\citeproctext]{ref-Satterthwaite1946-aq}
+Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
+
 \bibitem[\citeproctext]{ref-szucs2017empirical}
@@
 
+\bibitem[\citeproctext]{ref-Viechtbauer2007-tq}
+Viechtbauer, W.
+(2007).
+Approximate confidence intervals for standardized effect sizes in the two-independent and two-dependent samples design.
+\emph{Journal of Educational and Behavioral Statistics: A Quarterly Publication Sponsored by the American Educational Research Association and the American Statistical Association}, \emph{32}, 39--60.
+\url{https://doi.org/10.3102/1076998606298034}
+
+\bibitem[\citeproctext]{ref-Welch1947-wh}
+Welch, B. L. (1947). The generalization of {``student's''} problem when several different population variances are involved. \emph{Biometrika}, \emph{34}, 28. \url{https://doi.org/10.2307/2332510}
+
 \bibitem[\citeproctext]{ref-wilson2020science}

### 2024-10-01 paper/paper.Rmd
^^

@@
 abstract: |
-  Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size.
-These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest.
-Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty about the expected effect size.
+  Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size and alpha level.
+It can be useful to calculate the smallest effect that can be statistically significant and evaluate whether such effects are plausible, and if desireable that smaller effects can be detected.
+Reporting critical effect size values may be especially useful while adopting NHST (Null Hypothesis Significance Testing) when the sample size has not been planned a priori, or when there is uncertainty about the expected effect size that can be collected.
 To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
 The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
-What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations.
+What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations, especially non significant results.
 
@@
 
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
 The "critical effect size value(s)" refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample [@lakens2022sample].

### 2024-10-01 paper/paper.Rmd
^^

@@
 In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
-The "critical effect size value(s)" refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample [@lakens2022sample].
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's $r = -0.44$ and $r = 0.44$.
+The "critical effect size" refers to the smallest effect size that can be statistically significant, given the test performed, the sample size, and the alpha level [@lakens2022sample].
+Let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance.
+In this case, the critical effect sizes are Pearson's $r = -0.44$ and $r = 0.44$.
 As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
@@
 
-In psychology, statistical power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true replicable findings.
+In psychology, statistical power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true findings.
 In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology.
@@
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
-If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
@@
 Due to severe resource constraints, they are only able to collect a limited sample size of $n = 30$ ($n = 15$ per group).
+The authors did not perform a power, analysis, but their sample size justification is purely based on how many participants they can collect.
 No statistically significant effect is detected with $p < 0.05$.

### 2024-11-26 paper/paperDL.Rmd
^^

--- 2024-10-01 paper/paper.Rmd
@@
 abstract: |
-  Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size and alpha level.
-It can be useful to calculate the smallest effect that can be statistically significant and evaluate whether such effects are plausible, and if desireable that smaller effects can be detected.
-Reporting critical effect size values may be especially useful while adopting NHST (Null Hypothesis Significance Testing) when the sample size has not been planned a priori, or when there is uncertainty about the expected effect size that can be collected.
-To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
-The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
-What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations, especially non significant results.
+  Critical effect size values represent the smallest detectable effect that can reach statistical significance given a specific sample size and alpha level.
+It can be useful to calculate the critical effect size when designing a study, and evaluate whether such effects are plausible.
+Reporting critical effect size values may be useful when the sample size has not been planned a priori, or when there is uncertainty about the expected sample size that can be collected, when researchers plan to analyze the data with a statistical hypothesis test.
+To assist researchers in calculating critical effect size values we have developed an R package that allows researchers to report critical effect size values for group comparisons, correlations, linear regressions, and meta-analyses.
+Reflecting on critical effect size values could benefit researchers during the planning phase of the study by helping them to understand the limitations of their research design.
+Critical effect size values are also useful when evaluating studies performed by other researchers when a-priori power analyses were not performed, especially when non-significant results were observed.
 
-keywords          : critical effect size values, statistical significance, fucntions implementation 
+keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
 wordcount         : "3912"
@@
 
+The "critical effect size" refers to the smallest effect size that can be statistically significant, given the test performed, the sample size, and the alpha level [@lakens2022sample].
+Consider a study that tests a bivariate correlation with $n = 20$ participants and uses $\alpha < 0.05$ as the threshold for statistical significance in a two-sided test.
+The critical effect size values for this test are Pearson's $r = -0.44$ and $r = 0.44$, which means that the statistical test will yield a significant result for effects smaller than $r = -0.44$ or larger than $r = 0.44$, while observed effect sizes between these two values will lead to a statistically non-significant result.
 In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
-The "critical effect size" refers to the smallest effect size that can be statistically significant, given the test performed, the sample size, and the alpha level [@lakens2022sample].
-Let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance.
-In this case, the critical effect sizes are Pearson's $r = -0.44$ and $r = 0.44$.
-As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
+Critical effect size values can serve as a useful complement interpret study results, communicate the informational value of a study design, and contextualizes the interporetation of non-signiicant results, particularly when statistical significance is relied upon for inference.
+Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
 
-Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null].
-Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
-Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
-For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, which should subsequently be validated by "replication science".
+Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests [@cohen1994; @gigerenzer2004null; @krueger2001null].
+Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest.
+Researchers often publish studies with low power for medium to small effect sizes [@szucs2017empirical], with the main justification for the sample size being resource limitations [@lakens2022sample].
+However, low power makes it challenging to distinguish signal from noise, and combined with the selective reporting of statistically significant results in the published literature leads to an overestimation of effect sizes [@altoe2020enhancing].
+Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant.
 
-In psychology, statistical power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true findings.
-In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology.
-However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference [@altoe2020enhancing].
-On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts [@wilson2020science].
-Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
-
-There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
+We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
-If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
-This will focus the attention of researchers on the difference between statistical significance and practical significance.
-This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the 'crud factor' [@orben_crud_2020].
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform an informative hypothesis test.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
+Second, when sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
+This will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
 
 Let us consider the following two scenarios as examples.
-First, imagine researchers conducting a study involving a between-group comparison.
+First, imagine researchers who conduct a study involving a between-group comparison.
 Due to severe resource constraints, they are only able to collect a limited sample size of $n = 30$ ($n = 15$ per group).
-The authors did not perform a power, analysis, but their sample size justification is purely based on how many participants they can collect.
-No statistically significant effect is detected with $p < 0.05$.
-Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
+The authors did not perform an a-priori power analysis, and their sample size justification is purely based on how many participants they can collect.
+The statistical tests yields a non-significant result, $p > 0.05$.
+Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though without a theoretically expected effect size, it remains uncertain how low the statistical power is.
 Subsequently, they compute the critical effect size values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
-This indicates that any estimated |Cohen's $d| < 0.75$ will certainly fail to reach statistical significance.
-By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
-In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained with a small sample.
-Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+This indicates that any observed Cohen's $d$ between -0.75 and 0.75 will certainly fail to reach statistical significance.
+By reporting these critical effect size values the researchers transparently convey that estimated effects much larger than most effects in the psychological literature will always fall short of significance.
+This means that the hypothesis test will not be informative.
+Non-significant effects are likely, even if there is a true effect of interest.
+If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated  [@hedges1984estimation], and this should be taken into account when interpreting the effect size.
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
-However, they legitimately  opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
-In this context, the critical effect size values for significance are determined to be correlations of $r = \pm 0.03$.
-This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance.
+In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore bivariate correlations bwetween variables.
+With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are $r = \pm 0.03$.
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
-
 A similar scenario may arise in meta-analysis.
-Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance.
+Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, and as a consequence, even very small average effect sizes can reach significance.
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results.
 
 To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
+The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second study had a very large sample, and has attracted attention of the scientific community.
 The first case pertains to "Study 5" conducted by @mccrea2008self.
@@
 The resulting effect was statistically significant, $t(26) = 1.87$, $p$ (one-tailed) $< .05$.
-Based on the reported data, Cohen's $d = 0.736$, indicating a considerable effect size.
+Based on the reported data, Cohen's $d = 0.736$, the findings indicated a considerable effect size.
 However, the critical value for Cohen's $d$ in this context was $0.645$, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
+As the study was not a Registered Report, selective reporting of statistically significant findings was common at the time of publication, and most effects in psychology are substantially smaller, the effect size estimate is likely upwardly biased.
 Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample ($n = 61$), $t(59)=2.325$.
 The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428).
+THIS IS NOT A GOOD EXAMPLE, BECAUSE THESE TWO EFFECT SIZE ESTIMATES DO NOT DIFFER STATISTICALLY FROM EACH OTHER, SO WE CAN NOT SAY THE TWO EFFECT SIZES DIFFER!
+THERE SHOULD BE OTHER EXAMPLES IN THE LITERATURE, MAYBE FROM REGISTERED REPLICATION REPORTS.
+
 The second scenario involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of $n = 689'003$, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
-Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's $d$ in this case are $-0.006$ and $+0.006$, emphasizing the importance of considering effect size beyond mere significance.
-This prompts reflection on whether an effect of such magnitude on such a large sample (actual $d = 0.02$) is truly meaningful in individuals' lives.
+With a notably large sample size of $n = 689.003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
+Based on this and other results from the study, the authors conclude that emotions expressed by other users on Facebook influence our own emotions [CHANGE A BIT, STAY CLOSER TO WHAT THE RESEARCHERS SAY, MAYBE QUOTE THEM DIRECTLY].
+The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significancee.
+The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size ($d = 0.02$) is theoretically or practically interesting.
 
@@
 
-In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing critical effect size values in general [BUT WE FOCUS ON STANDARDIZED, NOT RAW, EFFECT SIZES - SO WE EITHER GIVE A REASON FOR THIS, OR WE ALSO ADD RAW EFFECTS], with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's $r$), and raw and standardized coefficients in linear models.
 These formulas have been incorporated into R functions of the package `criticalESvalue` accessible at: [https://github.com/psicostat/criticalESvalue](https://github.com/psicostat/criticalESvalue), and are elaborated upon in the subsequent section.
@@
 
-For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the $t$ statistics is computed as reported in Equation \@ref(eq:eq-gen-t).
+For the t-test we considered the two-sample, paired, and one-sample tests. As a general approach the $t$ statistic is computed as reported in Equation \@ref(eq:eq-gen-t).
 
@@
 Where $b$ is the unstandardized effect size that depends on the type of test.
-For example, in the two sample t-test or the single mean for a one sample t-test.
+[THE NEXT SENTENCE NEEDS TO BE REWRITTEN] For example, in the two sample $t$-test or the single mean for a one sample $t$-test.
 The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
 
-Similarly, Equation \@ref(eq:eq-gen-d) formalize a general form of the effect size.
+Similarly, Equation \@ref(eq:eq-gen-d) [REWRITE THE REST OF THIS SENTENCE] formalize a general form of the effect size.
 
@@
 
-Where $b$ is still the unstandardized effect size and $s$ is the standardization term.
-For example, in the two-sample case, $s$ is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.
+Where $b$ is the unstandardized effect size and $s$ is the standardization term.
+For example, in the two-sample case, $s$ is the pooled standard deviation between the two samples, while in the paired samples case, $s$ is the standard deviation of the differences for paired samples.
 
@@
 
-NOT SURE TO KEEP THIS!
+NOT SURE TO KEEP THIS! [DL: I THINK IT IS MAYBE FOR AN APPENDIX]
 
@@
 
-The easiest case is the one-sample t-test.
-$b$ is the sample mean and $s$ is the sample standard deviation and $n$ is the sample size.
+For the one-sample $t$-test $b$ is the sample mean and $s$ is the sample standard deviation and $n$ is the sample size.
 $SE_b = \frac{s}{\sqrt{n}}$ and the $t_c$ is based on $n - 1$ degrees of freedom.
 Thus $b_c = t_c \times SE_b$ and $d_c = \frac{b_c}{s}$.
-Equivalently, $d_c = t_c \times \frac{1}{\sqrt{n}}$
+Equivalently, $d_c = t_c \times \frac{1}{\sqrt{n}}$.
 
@@
 Another complication is that in some cases, researchers standardize the dependent variable, the independent variables or both.
-In general, also the unstandardized critical regression coefficient can be usually easily intepreted.
-For example, assuming we are regressing some scores on a cognitive tests using the age of participant.
-The critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
+In general, the unstandardized critical regression coefficient can be usually also be easily interpreted.
+For example, assuming we are regressing some scores on a cognitive tests using the age of participant,the critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
 
@@
 
-Despite we discussed only linear models, the same approach could be applied to other types of models such as generalized linear models.
+Despite the fact that we discussed only linear models, the same approach could be applied to other types of models such as generalized linear models.
 In fact, we simply need to multiply the critical value of the chosen distribution (e.g., $t$ or $z$) by the standard error of the regression coefficient.

### 2024-11-26 paper/paperDL.tex
^^

--- 2024-09-16 paper/paper.tex
@@
 \fi
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
   % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, fucntions implementation\newline\indent Word count: 3912}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 3912}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \usepackage{xcolor}
-\newcommand{\brev}{\color{red}}
-\newcommand{\erev}{\color{black}}
 \ifLuaTeX
@@
 \fi
-\usepackage{bookmark}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
   pdflang={en-EN},
-  pdfkeywords={critical effect size values, statistical significance, fucntions implementation},
+  pdfkeywords={critical effect size values, statistical significance, hypothesis testing, effect sizes},
   hidelinks,
@@
 \abstract{%
-Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size.
-These values are highly applicable in various scenarios, as they do not require specifying a plausible effect size of interest.
-Reporting critical effect size values may be especially useful when adopting NHST (Null Hypothesis Significance Testing) and the sample size has not been planned a priori, or when there is uncertainty about the expected effect size.
-To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
-The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
-What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations.
+Critical effect size values represent the smallest detectable effect that can reach statistical significance given a specific sample size and alpha level.
+It can be useful to calculate the critical effect size when designing a study, and evaluate whether such effects are plausible.
+Reporting critical effect size values may be useful when the sample size has not been planned a priori, or when there is uncertainty about the expected sample size that can be collected, when researchers plan to analyze the data with a statistical hypothesis test.
+To assist researchers in calculating critical effect size values we have developed an R package that allows researchers to report critical effect size values for group comparisons, correlations, linear regressions, and meta-analyses.
+Reflecting on critical effect size values could benefit researchers during the planning phase of the study by helping them to understand the limitations of their research design.
+Critical effect size values are also useful when evaluating studies performed by other researchers when a-priori power analyses were not performed, especially when non-significant results were observed.
 }
@@
 
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
-The ``critical effect size value(s)'' refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample (Lakens, 2022).
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
-As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
+The ``critical effect size'' refers to the smallest effect size that can be statistically significant, given the test performed, the sample size, and the alpha level (Lakens, 2022).
+Consider a study that tests a bivariate correlation with \(n = 20\) participants and uses \(\alpha < 0.05\) as the threshold for statistical significance in a two-sided test.
+The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects smaller than \(r = -0.44\) or larger than \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
+Critical effect size values can serve as a useful complement interpret study results, communicate the informational value of a study design, and contextualizes the interporetation of non-signiicant results, particularly when statistical significance is relied upon for inference.
+Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
 
-Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
-Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
-Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
-For instance, Wilson et al.
-(2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, which should subsequently be validated by ``replication science''.
+Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
+Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest.
+Researchers often publish studies with low power for medium to small effect sizes (Szucs \& Ioannidis, 2017), with the main justification for the sample size being resource limitations (Lakens, 2022).
+However, low power makes it challenging to distinguish signal from noise, and combined with the selective reporting of statistically significant results in the published literature leads to an overestimation of effect sizes (Altoè et al., 2020).
+Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant.
 
-In psychology, statistical power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings.
-In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size (Lakens, 2022), and this scenario is arguably widespread in psychology.
-However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference (Altoè et al., 2020).
-On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts (Wilson et al., 2020).
-Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
-
-There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
+We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
-If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
-This will focus the attention of researchers on the difference between statistical significance and practical significance.
-This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the `crud factor' (Orben \& Lakens, 2020).
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform an informative hypothesis test.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
+Second, when sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
+This will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
 
 Let us consider the following two scenarios as examples.
-First, imagine researchers conducting a study involving a between-group comparison.
+First, imagine researchers who conduct a study involving a between-group comparison.
 Due to severe resource constraints, they are only able to collect a limited sample size of \(n = 30\) (\(n = 15\) per group).
-No statistically significant effect is detected with \(p < 0.05\).
-Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
+The authors did not perform an a-priori power analysis, and their sample size justification is purely based on how many participants they can collect.
+The statistical tests yields a non-significant result, \(p > 0.05\).
+Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though without a theoretically expected effect size, it remains uncertain how low the statistical power is.
 Subsequently, they compute the critical effect size values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
-This indicates that any estimated \textbar Cohen's \(d| < 0.75\) will certainly fail to reach statistical significance.
-By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
-In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained with a small sample.
-Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+This indicates that any observed Cohen's \(d\) between -0.75 and 0.75 will certainly fail to reach statistical significance.
+By reporting these critical effect size values the researchers transparently convey that estimated effects much larger than most effects in the psychological literature will always fall short of significance.
+This means that the hypothesis test will not be informative.
+Non-significant effects are likely, even if there is a true effect of interest.
+If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated (Hedges, 1984), and this should be taken into account when interpreting the effect size.
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
-However, they legitimately opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
-In this context, the critical effect size values for significance are determined to be correlations of \(r = \pm 0.03\).
-This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance.
+In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore bivariate correlations bwetween variables.
+With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are \(r = \pm 0.03\).
 It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
-
 A similar scenario may arise in meta-analysis.
-Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance.
+Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, and as a consequence, even very small average effect sizes can reach significance.
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results.
 
 To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
+The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second study had a very large sample, and has attracted attention of the scientific community.
 The first case pertains to ``Study 5'' conducted by McCrea (2008).
@@
 The resulting effect was statistically significant, \(t(26) = 1.87\), \(p\) (one-tailed) \(< .05\).
-Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size.
+Based on the reported data, Cohen's \(d = 0.736\), the findings indicated a considerable effect size.
 However, the critical value for Cohen's \(d\) in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
+As the study was not a Registered Report, selective reporting of statistically significant findings was common at the time of publication, and most effects in psychology are substantially smaller, the effect size estimate is likely upwardly biased.
 Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\).
 The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428).
+THIS IS NOT A GOOD EXAMPLE, BECAUSE THESE TWO EFFECT SIZE ESTIMATES DO NOT DIFFER STATISTICALLY FROM EACH OTHER, SO WE CAN NOT SAY THE TWO EFFECT SIZES DIFFER!
+THERE SHOULD BE OTHER EXAMPLES IN THE LITERATURE, MAYBE FROM REGISTERED REPLICATION REPORTS.
+
 The second scenario involves the study by Kramer et al.
 (2014), which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of \(n = 689'003\), the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
-Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's \(d\) in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance.
-This prompts reflection on whether an effect of such magnitude on such a large sample (actual \(d = 0.02\)) is truly meaningful in individuals' lives.
+With a notably large sample size of \(n = 689.003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
+Based on this and other results from the study, the authors conclude that emotions expressed by other users on Facebook influence our own emotions {[}CHANGE A BIT, STAY CLOSER TO WHAT THE RESEARCHERS SAY, MAYBE QUOTE THEM DIRECTLY{]}.
+The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significancee.
+The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size (\(d = 0.02\)) is theoretically or practically interesting.
 
-\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}
+\hypertarget{how-to-compute-critical-effect-size-values}{%
+\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}}
 
-In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing critical effect size values in general {[}BUT WE FOCUS ON STANDARDIZED, NOT RAW, EFFECT SIZES - SO WE EITHER GIVE A REASON FOR THIS, OR WE ALSO ADD RAW EFFECTS{]}, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
 These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 
-\subsection{t-test}\label{t-test}
+\hypertarget{t-test}{%
+\subsection{t-test}\label{t-test}}
 
-For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the \(t\) statistics is computed as reported in Equation \eqref{eq:eq-gen-t}.
+For the t-test we considered the two-sample, paired, and one-sample tests. As a general approach the \(t\) statistic is computed as reported in Equation \eqref{eq:eq-gen-t}.
 
@@
 Where \(b\) is the unstandardized effect size that depends on the type of test.
-For example, in the two sample t-test or the single mean for a one sample t-test.
+{[}THE NEXT SENTENCE NEEDS TO BE REWRITTEN{]} For example, in the two sample \(t\)-test or the single mean for a one sample \(t\)-test.
 The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
 
-Similarly, Equation \eqref{eq:eq-gen-d} formalize a general form of the effect size.
+Similarly, Equation \eqref{eq:eq-gen-d} {[}REWRITE THE REST OF THIS SENTENCE{]} formalize a general form of the effect size.
 
@@
 
-Where \(b\) is still the unstandardized effect size and \(s\) is the standardization term.
-For example, in the two-sample case, \(s\) is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.
-
-\color{red}
+Where \(b\) is the unstandardized effect size and \(s\) is the standardization term.
+For example, in the two-sample case, \(s\) is the pooled standard deviation between the two samples, while in the paired samples case, \(s\) is the standard deviation of the differences for paired samples.
 
@@
 
-NOT SURE TO KEEP THIS!
+NOT SURE TO KEEP THIS! {[}DL: I THINK IT IS MAYBE FOR AN APPENDIX{]}
 
@@
 
-\color{black}
+\hypertarget{one-sample-t-test}{%
+\subsubsection{One sample t-test}\label{one-sample-t-test}}
 
-\color{red}
-
-\subsubsection{One sample t-test}\label{one-sample-t-test}
-
-The easiest case is the one-sample t-test.
-\(b\) is the sample mean and \(s\) is the sample standard deviation and \(n\) is the sample size.
+For the one-sample \(t\)-test \(b\) is the sample mean and \(s\) is the sample standard deviation and \(n\) is the sample size.
 \(SE_b = \frac{s}{\sqrt{n}}\) and the \(t_c\) is based on \(n - 1\) degrees of freedom.
 Thus \(b_c = t_c \times SE_b\) and \(d_c = \frac{b_c}{s}\).
-Equivalently, \(d_c = t_c \times \frac{1}{\sqrt{n}}\)
+Equivalently, \(d_c = t_c \times \frac{1}{\sqrt{n}}\).
 
-\color{black}
-
-\subsubsection{Two-sample t-test}\label{two-sample-t-test}
+\hypertarget{two-sample-t-test}{%
+\subsubsection{Two-sample t-test}\label{two-sample-t-test}}
 
@@
 
-\subsubsection{Paired sample t-test}\label{paired-sample-t-test}
+\hypertarget{paired-sample-t-test}{%
+\subsubsection{Paired sample t-test}\label{paired-sample-t-test}}
 
@@
 
-\subsubsection{Hedges's correction}\label{hedgess-correction}
+\hypertarget{hedgess-correction}{%
+\subsubsection{Hedges's correction}\label{hedgess-correction}}
 
@@
 
-\subsection{Correlation Test}\label{correlation-test}
+\hypertarget{correlation-test}{%
+\subsection{Correlation Test}\label{correlation-test}}
 
@@
 
-\subsection{Linear Regression}\label{linear-regression}
+\hypertarget{linear-regression}{%
+\subsection{Linear Regression}\label{linear-regression}}
 
@@
 Another complication is that in some cases, researchers standardize the dependent variable, the independent variables or both.
-In general, also the unstandardized critical regression coefficient can be usually easily intepreted.
-For example, assuming we are regressing some scores on a cognitive tests using the age of participant.
-The critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
+In general, the unstandardized critical regression coefficient can be usually also be easily interpreted.
+For example, assuming we are regressing some scores on a cognitive tests using the age of participant,the critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
 
-\subsection{Meta-analysis}\label{meta-analysis}
+\hypertarget{meta-analysis}{%
+\subsection{Meta-analysis}\label{meta-analysis}}
 
@@
 
-\subsection{Other models}\label{other-models}
+\hypertarget{other-models}{%
+\subsection{Other models}\label{other-models}}
 
-Despite we discussed only linear models, the same approach could be applied to other types of models such as generalized linear models.
+Despite the fact that we discussed only linear models, the same approach could be applied to other types of models such as generalized linear models.
 In fact, we simply need to multiply the critical value of the chosen distribution (e.g., \(t\) or \(z\)) by the standard error of the regression coefficient.
@@
 
-\section{Examples in R}\label{examples-in-r}
+\hypertarget{examples-in-r}{%
+\section{Examples in R}\label{examples-in-r}}
 
@@
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-\phantomsection\label{refs}
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-altoe2020enhancing}
+\leavevmode\vadjust pre{\hypertarget{ref-altoe2020enhancing}{}}%
 Altoè, G., Bertoldo, G., Zandonella Callegher, C., Toffalini, E., Calcagnı̀, A., Finos, L., \& Pastore, M.
@@
 
-\bibitem[\citeproctext]{ref-Borenstein2009-mo}
+\leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
 Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
 
-\bibitem[\citeproctext]{ref-Bowley1928-yt}
+\leavevmode\vadjust pre{\hypertarget{ref-Bowley1928-yt}{}}%
 Bowley, A. L. (1928). The standard deviation of the correlation coefficient. \emph{Journal of the American Statistical Association}, \emph{23}, 31. \url{https://doi.org/10.2307/2277400}
 
-\bibitem[\citeproctext]{ref-cohen1994}
+\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
 Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.
 
-\bibitem[\citeproctext]{ref-open2015estimating}
+\leavevmode\vadjust pre{\hypertarget{ref-open2015estimating}{}}%
 Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
 
-\bibitem[\citeproctext]{ref-Fisher1915-ks}
+\leavevmode\vadjust pre{\hypertarget{ref-Fisher1915-ks}{}}%
 Fisher, R.
@@
 
-\bibitem[\citeproctext]{ref-Fisher1921-qz}
+\leavevmode\vadjust pre{\hypertarget{ref-Fisher1921-qz}{}}%
 Fisher, R. A. (1921). On the "probable error" of a coefficient of correlation deduced from a small sample. \emph{Metron}.
 
-\bibitem[\citeproctext]{ref-funder2019evaluating}
+\leavevmode\vadjust pre{\hypertarget{ref-funder2019evaluating}{}}%
 Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: Sense and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
 
-\bibitem[\citeproctext]{ref-Gelman2008-zp}
+\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008-zp}{}}%
 Gelman, A. (2008). Scaling regression inputs by dividing by two standard deviations. \emph{Statistics in Medicine}, \emph{27}, 2865--2873. \url{https://doi.org/10.1002/sim.3107}
 
-\bibitem[\citeproctext]{ref-gigerenzer2004null}
+\leavevmode\vadjust pre{\hypertarget{ref-gigerenzer2004null}{}}%
 Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.
 
-\bibitem[\citeproctext]{ref-Gross2023-fl}
+\leavevmode\vadjust pre{\hypertarget{ref-Gross2023-fl}{}}%
 Groß, J., \& Möller, A.
@@
 
-\bibitem[\citeproctext]{ref-Hedges1981-za}
+\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
 Hedges, L.
@@
 
-\bibitem[\citeproctext]{ref-hedges1984estimation}
+\leavevmode\vadjust pre{\hypertarget{ref-hedges1984estimation}{}}%
 Hedges, L.
@@
 
-\bibitem[\citeproctext]{ref-kramer2014experimental}
+\leavevmode\vadjust pre{\hypertarget{ref-kramer2014experimental}{}}%
 Kramer, A.
@@
 
-\bibitem[\citeproctext]{ref-krueger2001null}
+\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
 Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.
 
-\bibitem[\citeproctext]{ref-lakens2013calculating}
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
 Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.
 
-\bibitem[\citeproctext]{ref-lakens2022sample}
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
 Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.
 
-\bibitem[\citeproctext]{ref-mccrea2008self}
+\leavevmode\vadjust pre{\hypertarget{ref-mccrea2008self}{}}%
 McCrea, S.
@@
 
-\bibitem[\citeproctext]{ref-morris2002combining}
+\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
 
-\bibitem[\citeproctext]{ref-orben_crud_2020}
+\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
 
-\bibitem[\citeproctext]{ref-Satterthwaite1946-aq}
+\leavevmode\vadjust pre{\hypertarget{ref-Satterthwaite1946-aq}{}}%
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
 
-\bibitem[\citeproctext]{ref-szucs2017empirical}
+\leavevmode\vadjust pre{\hypertarget{ref-szucs2017empirical}{}}%
 Szucs, D., \& Ioannidis, J.
@@
 
-\bibitem[\citeproctext]{ref-transue2019apa}
+\leavevmode\vadjust pre{\hypertarget{ref-transue2019apa}{}}%
 Transue, B. (2019). \emph{APA style 7th edition}.
 
-\bibitem[\citeproctext]{ref-Viechtbauer2007-tq}
+\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2007-tq}{}}%
 Viechtbauer, W.
@@
 
-\bibitem[\citeproctext]{ref-Welch1947-wh}
+\leavevmode\vadjust pre{\hypertarget{ref-Welch1947-wh}{}}%
 Welch, B. L. (1947). The generalization of {``student's''} problem when several different population variances are involved. \emph{Biometrika}, \emph{34}, 28. \url{https://doi.org/10.2307/2332510}
 
-\bibitem[\citeproctext]{ref-wilson2020science}
+\leavevmode\vadjust pre{\hypertarget{ref-wilson2020science}{}}%
 Wilson, B. M., Harris, C. R., \& Wixted, J. T. (2020). Science is not a signal detection problem. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(11), 5559--5567.

### 2024-12-11 paper/paperDL_old.Rmd
FG reply to DL and adding new files (old are _old)

--- 2024-11-26 paper/paperDL.Rmd
@@
 
-To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second study had a very large sample, and has attracted attention of the scientific community.
-The first case pertains to "Study 5" conducted by @mccrea2008self.
-This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure.
-One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward.
-With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, $t(26) = 1.87$, $p$ (one-tailed) $< .05$.
-Based on the reported data, Cohen's $d = 0.736$, the findings indicated a considerable effect size.
-However, the critical value for Cohen's $d$ in this context was $0.645$, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-As the study was not a Registered Report, selective reporting of statistically significant findings was common at the time of publication, and most effects in psychology are substantially smaller, the effect size estimate is likely upwardly biased.
-Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample ($n = 61$), $t(59)=2.325$.
-The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428).
-THIS IS NOT A GOOD EXAMPLE, BECAUSE THESE TWO EFFECT SIZE ESTIMATES DO NOT DIFFER STATISTICALLY FROM EACH OTHER, SO WE CAN NOT SAY THE TWO EFFECT SIZES DIFFER!
-THERE SHOULD BE OTHER EXAMPLES IN THE LITERATURE, MAYBE FROM REGISTERED REPLICATION REPORTS.
+\textcolor{red}{To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine two real-world instances drawn from published research.
+The first study was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
+Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+A two-tailed $t$-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+}
 
@@
 With a notably large sample size of $n = 689.003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
-Based on this and other results from the study, the authors conclude that emotions expressed by other users on Facebook influence our own emotions [CHANGE A BIT, STAY CLOSER TO WHAT THE RESEARCHERS SAY, MAYBE QUOTE THEM DIRECTLY].
-The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
-It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significancee.
+Based on this and other results from the study, the authors conclude that \textcolor{red}{"the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks".} The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
 The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size ($d = 0.02$) is theoretically or practically interesting.

### 2024-12-11 paper/paperDL_old.tex
FG reply to DL and adding new files (old are _old)

--- 2024-11-26 paper/paperDL.tex
@@
 
-To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second study had a very large sample, and has attracted attention of the scientific community.
-The first case pertains to ``Study 5'' conducted by McCrea (2008).
-This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure.
-One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward.
-With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, \(t(26) = 1.87\), \(p\) (one-tailed) \(< .05\).
-Based on the reported data, Cohen's \(d = 0.736\), the findings indicated a considerable effect size.
-However, the critical value for Cohen's \(d\) in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-As the study was not a Registered Report, selective reporting of statistically significant findings was common at the time of publication, and most effects in psychology are substantially smaller, the effect size estimate is likely upwardly biased.
-Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\).
-The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428).
-THIS IS NOT A GOOD EXAMPLE, BECAUSE THESE TWO EFFECT SIZE ESTIMATES DO NOT DIFFER STATISTICALLY FROM EACH OTHER, SO WE CAN NOT SAY THE TWO EFFECT SIZES DIFFER!
-THERE SHOULD BE OTHER EXAMPLES IN THE LITERATURE, MAYBE FROM REGISTERED REPLICATION REPORTS.
+\textcolor{red}{To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine two real-world instances drawn from published research.
+The first study was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
+Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+A two-tailed $t$-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+}
 
@@
 With a notably large sample size of \(n = 689.003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
-Based on this and other results from the study, the authors conclude that emotions expressed by other users on Facebook influence our own emotions {[}CHANGE A BIT, STAY CLOSER TO WHAT THE RESEARCHERS SAY, MAYBE QUOTE THEM DIRECTLY{]}.
-The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
-It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significancee.
+Based on this and other results from the study, the authors conclude that \textcolor{red}{"the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks".} The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
 The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size (\(d = 0.02\)) is theoretically or practically interesting.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-open2015estimating}{}}%
-Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
-
 \leavevmode\vadjust pre{\hypertarget{ref-Fisher1915-ks}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mccrea2008self}{}}%
-McCrea, S.
-M.
-(2008).
-Self-handicapping, excuse making, and counterfactual thinking: Consequences for self-esteem and future motivation.
-\emph{Journal of Personality and Social Psychology}, \emph{95}(2), 274.
-
 \leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%

### 2024-12-14 paper/paperDL.Rmd
^^

@@
     \usepackage{xcolor}
+    \newcommand{\bfilippo}{\color{blue}}
+    \newcommand{\efilippo}{\color{black}}
 ---
@@
 Consider a study that tests a bivariate correlation with $n = 20$ participants and uses $\alpha < 0.05$ as the threshold for statistical significance in a two-sided test.
-The critical effect size values for this test are Pearson's $r = -0.44$ and $r = 0.44$, which means that the statistical test will yield a significant result for effects smaller than $r = -0.44$ or larger than $r = 0.44$, while observed effect sizes between these two values will lead to a statistically non-significant result.
+The critical effect size values for this test are Pearson's $r = -0.44$ and $r = 0.44$, which means that the statistical test will yield a significant result for effects larger than $r = -0.44$ or $r = 0.44$, while observed effect sizes between these two values will lead to a statistically non-significant result.
 In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
-Critical effect size values can serve as a useful complement interpret study results, communicate the informational value of a study design, and contextualizes the interporetation of non-signiicant results, particularly when statistical significance is relied upon for inference.
+Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
 Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
 
-Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests [@cohen1994; @gigerenzer2004null; @krueger2001null].
+Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests [@lakens2021practical; @chow1988significance; @cortina1997logic; @hagen1997praise; @haig2017tests; @miller2017hypothesis; @nickerson2000null; @cohen1994; @gigerenzer2004null; @krueger2001null].
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
@@
 We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
+In the following sections we will illustrate examples of common scenarios and we will go through the main problems and possible solutions, highlighting the utility of critical effect size values in such contexts.
+
+## Small Sample Sizes and Uncertain sample size determination
+
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-Second, when sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
-This will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
-This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
+\textcolor{red}{In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
+The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.}
 
@@
 This means that the hypothesis test will not be informative.
-Non-significant effects are likely, even if there is a true effect of interest.
+
+\textcolor{red}{To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine two real-world instances drawn from published research.
+The first study was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
+Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+A two-tailed $t$-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+} Non-significant effects are likely, even if there is a true effect of interest.
 If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated  [@hedges1984estimation], and this should be taken into account when interpreting the effect size.
+\textcolor{red}{Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn, and for those cases where sample size constrains will make it impossible to find an effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multi-lab studies (\cite{AAA}), should be explored.}
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore bivariate correlations bwetween variables.
+## Large Sample Sizes and Meta-analysis
+
+When sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
+Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
+
+Imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore bivariate correlations between variables.
 With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are $r = \pm 0.03$.
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
+
+The second real-world instance involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
+With a notably large sample size of $n = 689.003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
+Based on this and other results from the study, the authors conclude that \textcolor{red}{"the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks".} The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
+The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size ($d = 0.02$) is theoretically or practically interesting.
+
 A similar scenario may arise in meta-analysis.
@@
 
-To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second study had a very large sample, and has attracted attention of the scientific community.
-The first case pertains to "Study 5" conducted by @mccrea2008self.
-This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure.
-One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward.
-With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, $t(26) = 1.87$, $p$ (one-tailed) $< .05$.
-Based on the reported data, Cohen's $d = 0.736$, the findings indicated a considerable effect size.
-However, the critical value for Cohen's $d$ in this context was $0.645$, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-As the study was not a Registered Report, selective reporting of statistically significant findings was common at the time of publication, and most effects in psychology are substantially smaller, the effect size estimate is likely upwardly biased.
-Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample ($n = 61$), $t(59)=2.325$.
-The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428).
-THIS IS NOT A GOOD EXAMPLE, BECAUSE THESE TWO EFFECT SIZE ESTIMATES DO NOT DIFFER STATISTICALLY FROM EACH OTHER, SO WE CAN NOT SAY THE TWO EFFECT SIZES DIFFER!
-THERE SHOULD BE OTHER EXAMPLES IN THE LITERATURE, MAYBE FROM REGISTERED REPLICATION REPORTS.
-
-The second scenario involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of $n = 689.003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
-Based on this and other results from the study, the authors conclude that emotions expressed by other users on Facebook influence our own emotions [CHANGE A BIT, STAY CLOSER TO WHAT THE RESEARCHERS SAY, MAYBE QUOTE THEM DIRECTLY].
-The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
-It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significancee.
-The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size ($d = 0.02$) is theoretically or practically interesting.
-
 # How to Compute Critical Effect Size Values
 
-In this section, we provide guidance and formulas for computing critical effect size values in general [BUT WE FOCUS ON STANDARDIZED, NOT RAW, EFFECT SIZES - SO WE EITHER GIVE A REASON FOR THIS, OR WE ALSO ADD RAW EFFECTS], with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's $r$), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package `criticalESvalue` accessible at: [https://github.com/psicostat/criticalESvalue](https://github.com/psicostat/criticalESvalue), and are elaborated upon in the subsequent section.
+\bfilippo In this section, we provide guidance and formulas for computing standardized critical effect size with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's $r$), and raw and standardized coefficients in linear models.
+The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
+\efilippo These formulas have been incorporated into R functions of the package `criticalESvalue` accessible at: [https://github.com/psicostat/criticalESvalue](https://github.com/psicostat/criticalESvalue), and are elaborated upon in the subsequent section.
 
@@
 Where $b$ is the unstandardized effect size that depends on the type of test.
-[THE NEXT SENTENCE NEEDS TO BE REWRITTEN] For example, in the two sample $t$-test or the single mean for a one sample $t$-test.
-The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
+\bfilippo For example, in the two sample $t$-test the numerator is the difference between the two sample means while in the one-sample case is the difference between the sample mean and the population value.
+The denominator is the standard error of the numerator that depends on the sample size and the samples variances.
+\efilippo
 
-Similarly, Equation \@ref(eq:eq-gen-d) [REWRITE THE REST OF THIS SENTENCE] formalize a general form of the effect size.
+\bfilippo
+Equation \@ref(eq:eq-gen-d) describe a general formulation of a standardized effect size measure.
+Where $b$ is the unstandardized effect size (e.g., the difference between two means) and $s$ is the standardization term \efilippo.
+For example, in the two-samples case, $s$ is the pooled standard deviation between the two samples, while in the paired samples case, $s$ is the standard deviation of the differences for paired samples.
 
@@
 \end{equation}
-
-Where $b$ is the unstandardized effect size and $s$ is the standardization term.
-For example, in the two-sample case, $s$ is the pooled standard deviation between the two samples, while in the paired samples case, $s$ is the standard deviation of the differences for paired samples.
 
@@
 \end{equation}
-
-NOT SURE TO KEEP THIS! [DL: I THINK IT IS MAYBE FOR AN APPENDIX]
 

### 2024-12-14 paper/paperDL.tex
^^

@@
 \usepackage{xcolor}
+\newcommand{\bfilippo}{\color{blue}}
+\newcommand{\efilippo}{\color{black}}
 \ifLuaTeX
@@
 Consider a study that tests a bivariate correlation with \(n = 20\) participants and uses \(\alpha < 0.05\) as the threshold for statistical significance in a two-sided test.
-The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects smaller than \(r = -0.44\) or larger than \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
+The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects larger than \(r = -0.44\) or \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
 In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
-Critical effect size values can serve as a useful complement interpret study results, communicate the informational value of a study design, and contextualizes the interporetation of non-signiicant results, particularly when statistical significance is relied upon for inference.
+Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
 Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
 
-Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
+Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests (Chow, 1988; Cohen, 1994; Cortina \& Dunlap, 1997; Gigerenzer et al., 2004; Hagen, 1997; Haig, 2017; Krueger, 2001; Lakens, 2021; Miller, 2017; Nickerson, 2000).
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
@@
 We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
+In the following sections we will illustrate examples of common scenarios and we will go through the main problems and possible solutions, highlighting the utility of critical effect size values in such contexts.
+
+\hypertarget{small-sample-sizes-and-uncertain-sample-size-determination}{%
+\subsection{Small Sample Sizes and Uncertain sample size determination}\label{small-sample-sizes-and-uncertain-sample-size-determination}}
+
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-Second, when sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
-This will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
-This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
+\textcolor{red}{In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
+The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to lack of power in a very intuitive way.}
 
@@
 This means that the hypothesis test will not be informative.
-Non-significant effects are likely, even if there is a true effect of interest.
+
+\textcolor{red}{To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine two real-world instances drawn from published research.
+The first study was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
+Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+A two-tailed $t$-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+} Non-significant effects are likely, even if there is a true effect of interest.
 If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated (Hedges, 1984), and this should be taken into account when interpreting the effect size.
+\textcolor{red}{Beyhond critical effect size values, researchers should reason before conducting a study on what conclusions can be deawn, and for those cases where sample size constrains will make it impossible to find an effect bacuse it is likely smaller than the critical effect size value, alternative options for data collection, such as multi-lab studies (\cite{AAA}), should be explored.}
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore bivariate correlations bwetween variables.
+\hypertarget{large-sample-sizes-and-meta-analysis}{%
+\subsection{Large Sample Sizes and Meta-analysis}\label{large-sample-sizes-and-meta-analysis}}
+
+When sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
+This will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
+
+Imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore bivariate correlations bwetween variables.
 With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are \(r = \pm 0.03\).
@@
 
-To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second study had a very large sample, and has attracted attention of the scientific community.
-The first case pertains to ``Study 5'' conducted by McCrea (2008).
-This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure.
-One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward.
-With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, \(t(26) = 1.87\), \(p\) (one-tailed) \(< .05\).
-Based on the reported data, Cohen's \(d = 0.736\), the findings indicated a considerable effect size.
-However, the critical value for Cohen's \(d\) in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-As the study was not a Registered Report, selective reporting of statistically significant findings was common at the time of publication, and most effects in psychology are substantially smaller, the effect size estimate is likely upwardly biased.
-Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\).
-The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428).
-THIS IS NOT A GOOD EXAMPLE, BECAUSE THESE TWO EFFECT SIZE ESTIMATES DO NOT DIFFER STATISTICALLY FROM EACH OTHER, SO WE CAN NOT SAY THE TWO EFFECT SIZES DIFFER!
-THERE SHOULD BE OTHER EXAMPLES IN THE LITERATURE, MAYBE FROM REGISTERED REPLICATION REPORTS.
-
-The second scenario involves the study by Kramer et al.
+The second real-world instance involves the study by Kramer et al.
 (2014), which explored the impact of emotional content on Facebook users' experiences.
 With a notably large sample size of \(n = 689.003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
-Based on this and other results from the study, the authors conclude that emotions expressed by other users on Facebook influence our own emotions {[}CHANGE A BIT, STAY CLOSER TO WHAT THE RESEARCHERS SAY, MAYBE QUOTE THEM DIRECTLY{]}.
-The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
-It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significancee.
+Based on this and other results from the study, the authors conclude that \textcolor{red}{"the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks".} The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
 The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size (\(d = 0.02\)) is theoretically or practically interesting.
@@
 
-In this section, we provide guidance and formulas for computing critical effect size values in general {[}BUT WE FOCUS ON STANDARDIZED, NOT RAW, EFFECT SIZES - SO WE EITHER GIVE A REASON FOR THIS, OR WE ALSO ADD RAW EFFECTS{]}, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
+\color{blue} In this section, we provide guidance and formulas for computing standardized critical effect size with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
+The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
+\color{black} These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 
@@
 Where \(b\) is the unstandardized effect size that depends on the type of test.
-{[}THE NEXT SENTENCE NEEDS TO BE REWRITTEN{]} For example, in the two sample \(t\)-test or the single mean for a one sample \(t\)-test.
-The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
+\color{blue} For example, in the two sample \(t\)-test the numerator is the difference between the two sample means while in the one-sample case is the difference between the sample mean and the population value.
+The denominator is the standard error of the numerator that depends on the sample size and the samples variances.
+\color{black}
 
-Similarly, Equation \eqref{eq:eq-gen-d} {[}REWRITE THE REST OF THIS SENTENCE{]} formalize a general form of the effect size.
+\color{blue}
+
+Equation \eqref{eq:eq-gen-d} describe a general formulation of a standardized effect size measure.
+Where \(b\) is the unstandardized effect size (e.g., the difference between two means) and \(s\) is the standardization term \color{black}.
+For example, in the two-samples case, \(s\) is the pooled standard deviation between the two samples, while in the paired samples case, \(s\) is the standard deviation of the differences for paired samples.
 
@@
 \end{equation}
-
-Where \(b\) is the unstandardized effect size and \(s\) is the standardization term.
-For example, in the two-sample case, \(s\) is the pooled standard deviation between the two samples, while in the paired samples case, \(s\) is the standard deviation of the differences for paired samples.
 
@@
 \end{equation}
-
-NOT SURE TO KEEP THIS! {[}DL: I THINK IT IS MAYBE FOR AN APPENDIX{]}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-chow1988significance}{}}%
+Chow, S. L. (1988). Significance test or effect size? \emph{Psychological Bulletin}, \emph{103}(1), 105.
+
 \leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-open2015estimating}{}}%
-Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
+\leavevmode\vadjust pre{\hypertarget{ref-cortina1997logic}{}}%
+Cortina, J. M., \& Dunlap, W. P. (1997). On the logic and purpose of significance testing. \emph{Psychological Methods}, \emph{2}(2), 161.
 
@@
 \url{https://doi.org/10.1007/s42519-023-00323-w}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hagen1997praise}{}}%
+Hagen, R. L. (1997). \emph{In praise of the null hypothesis statistical test.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-haig2017tests}{}}%
+Haig, B. D. (2017). Tests of statistical significance made sound. \emph{Educational and Psychological Measurement}, \emph{77}(3), 489--506.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2021practical}{}}%
+Lakens, D. (2021). The practical alternative to the p value is the correctly used p value. \emph{Perspectives on Psychological Science}, \emph{16}(3), 639--648.
+
 \leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mccrea2008self}{}}%
-McCrea, S.
-M.
-(2008).
-Self-handicapping, excuse making, and counterfactual thinking: Consequences for self-esteem and future motivation.
-\emph{Journal of Personality and Social Psychology}, \emph{95}(2), 274.
+\leavevmode\vadjust pre{\hypertarget{ref-miller2017hypothesis}{}}%
+Miller, J. (2017). Hypothesis testing in the real world. \emph{Educational and Psychological Measurement}, \emph{77}(4), 663--672.
 
@@
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
+
+\leavevmode\vadjust pre{\hypertarget{ref-nickerson2000null}{}}%
+Nickerson, R. S. (2000). Null hypothesis significance testing: A review of an old and continuing controversy. \emph{Psychological Methods}, \emph{5}(2), 241.
 

### 2024-12-14 paper/paper.tex
pulitp un po' la repo

@@
 \usepackage{xcolor}
-\newcommand{\brev}{\color{red}}
-\newcommand{\erev}{\color{black}}
+\newcommand{\bfilippo}{\color{blue}}
+\newcommand{\efilippo}{\color{black}}
 \ifLuaTeX
@@
 
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using statistical significance.
-The ``critical effect size value(s)'' refers to the smallest statistically significant effect sizes, which depend on the test performed and the characteristics of the sample (Lakens, 2022).
-To cut to the chase, let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with \(n = 20\) participants and uses \(p < 0.05\) as the threshold for statistical significance: in this case, the smallest significant effects (critical effect size values) are Pearson's \(r = -0.44\) and \(r = 0.44\).
-As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
+The ``critical effect size'' refers to the smallest effect size that can be statistically significant, given the test performed, the sample size, and the alpha level (Lakens, 2022).
+Consider a study that tests a bivariate correlation with \(n = 20\) participants and uses \(\alpha < 0.05\) as the threshold for statistical significance in a two-sided test.
+The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects larger than \(r = -0.44\) or \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
+Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
+Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
 
-Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science (Cohen, 1994; Gigerenzer et al., 2004; Krueger, 2001).
-Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance (Transue, 2019).
-Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
-For instance, Wilson et al.
-(2020) propose that it could serve as a useful preliminary filter utilized by ``original science'' for screening potentially interesting effects, which should subsequently be validated by ``replication science''.
+Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests (Chow, 1988; Cohen, 1994; Cortina \& Dunlap, 1997; Gigerenzer et al., 2004; Hagen, 1997; Haig, 2017; Krueger, 2001; Lakens, 2021; Miller, 2017; Nickerson, 2000).
+Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest.
+Researchers often publish studies with low power for medium to small effect sizes (Szucs \& Ioannidis, 2017), with the main justification for the sample size being resource limitations (Lakens, 2022).
+However, low power makes it challenging to distinguish signal from noise, and combined with the selective reporting of statistically significant results in the published literature leads to an overestimation of effect sizes (Altoè et al., 2020).
+Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant.
 
-In psychology, statistical power is often severely limited for medium to small effect sizes (Szucs \& Ioannidis, 2017), which are generally expected to represent true replicable findings.
-In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size (Lakens, 2022), and this scenario is arguably widespread in psychology.
-However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference (Altoè et al., 2020).
-On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an ``everything-is-significant'' scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts (Wilson et al., 2020).
-Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
+We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
+In the following sections we will illustrate examples of common scenarios and we will go through the main problems and possible solutions, highlighting the utility of critical effect size values in such contexts.
 
-There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
+\hypertarget{small-sample-sizes-and-uncertain-sample-size-determination}{%
+\subsection{Small Sample Sizes and Uncertain sample size determination}\label{small-sample-sizes-and-uncertain-sample-size-determination}}
+
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
-If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, in these cases, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
-This will focus the attention of researchers on the difference between statistical significance and practical significance.
-This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the `crud factor' (Orben \& Lakens, 2020).
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform an informative hypothesis test.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
+\textcolor{red}{In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
+The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.}
 
 Let us consider the following two scenarios as examples.
-First, imagine researchers conducting a study involving a between-group comparison.
+First, imagine researchers who conduct a study involving a between-group comparison.
 Due to severe resource constraints, they are only able to collect a limited sample size of \(n = 30\) (\(n = 15\) per group).
-No statistically significant effect is detected with \(p < 0.05\).
-Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
+The authors did not perform an a-priori power analysis, and their sample size justification is purely based on how many participants they can collect.
+The statistical tests yields a non-significant result, \(p > 0.05\).
+Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though without a theoretically expected effect size, it remains uncertain how low the statistical power is.
 Subsequently, they compute the critical effect size values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
-This indicates that any estimated \textbar Cohen's \(d| < 0.75\) will certainly fail to reach statistical significance.
-By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
-In essence, this scenario exemplifies a version of the ``winner's curse'' (Hedges, 1984), made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a ``winner'') to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the ``curse''), especially if such initial finding was obtained with a small sample.
-Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+This indicates that any observed Cohen's \(d\) between -0.75 and 0.75 will certainly fail to reach statistical significance.
+By reporting these critical effect size values the researchers transparently convey that estimated effects much larger than most effects in the psychological literature will always fall short of significance.
+This means that the hypothesis test will not be informative.
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
-However, they legitimately opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
-In this context, the critical effect size values for significance are determined to be correlations of \(r = \pm 0.03\).
-This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance.
+\textcolor{red}{To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine two real-world instances drawn from published research.
+The first study was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
+Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+A two-tailed $t$-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+} Non-significant effects are likely, even if there is a true effect of interest.
+If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated (Hedges, 1984), and this should be taken into account when interpreting the effect size.
+\textcolor{red}{Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn, and for those cases where sample size constrains will make it impossible to find an effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multi-lab studies (\cite{AAA}), should be explored.}
+
+\hypertarget{large-sample-sizes-and-meta-analysis}{%
+\subsection{Large Sample Sizes and Meta-analysis}\label{large-sample-sizes-and-meta-analysis}}
+
+When sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
+Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
+
+Imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore bivariate correlations between variables.
+With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are \(r = \pm 0.03\).
 It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
+
+The second real-world instance involves the study by Kramer et al.
+(2014), which explored the impact of emotional content on Facebook users' experiences.
+With a notably large sample size of \(n = 689.003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
+Based on this and other results from the study, the authors conclude that \textcolor{red}{"the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks".} The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
+The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size (\(d = 0.02\)) is theoretically or practically interesting.
 
 A similar scenario may arise in meta-analysis.
-Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance.
+Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, and as a consequence, even very small average effect sizes can reach significance.
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results.
 
-To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration (Collaboration, 2015), while the second one presents with a very large sample and has already attracted attention in the scientific community.
-The first case pertains to ``Study 5'' conducted by McCrea (2008).
-This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure.
-One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward.
-With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, \(t(26) = 1.87\), \(p\) (one-tailed) \(< .05\).
-Based on the reported data, Cohen's \(d = 0.736\), indicating a considerable effect size.
-However, the critical value for Cohen's \(d\) in this context was \(0.645\), suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-Subsequent replication efforts (Collaboration, 2015) reaffirmed the significance of the effect, and used a somewhat larger sample (\(n = 61\)), \(t(59)=2.325\).
-The estimated effect size was somehow lower than that originally reported, however, \(d = 0.605\) (here, critical d = 0.428).
-The second scenario involves the study by Kramer et al.
-(2014), which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of \(n = 689'003\), the researchers observed a significant effect worsened emotional states when positive posts were reduced, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
-Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's \(d\) in this case are \(-0.006\) and \(+0.006\), emphasizing the importance of considering effect size beyond mere significance.
-This prompts reflection on whether an effect of such magnitude on such a large sample (actual \(d = 0.02\)) is truly meaningful in individuals' lives.
+\hypertarget{how-to-compute-critical-effect-size-values}{%
+\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}}
 
-\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}
+\color{blue} In this section, we provide guidance and formulas for computing standardized critical effect size with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
+The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
+\color{black} These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 
-In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's r), and raw and standardized coefficients in linear models.
-These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
+\hypertarget{t-test}{%
+\subsection{t-test}\label{t-test}}
 
-\subsection{t-test}\label{t-test}
-
-For the t-test we considered the two-sample, paired and one-sample tests. As a general approach the \(t\) statistics is computed as reported in Equation \eqref{eq:eq-gen-t}.
+For the t-test we considered the two-sample, paired, and one-sample tests. As a general approach the \(t\) statistic is computed as reported in Equation \eqref{eq:eq-gen-t}.
 
@@
 Where \(b\) is the unstandardized effect size that depends on the type of test.
-For example, in the two sample t-test or the single mean for a one sample t-test.
-The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
+\color{blue} For example, in the two sample \(t\)-test the numerator is the difference between the two sample means while in the one-sample case is the difference between the sample mean and the population value.
+The denominator is the standard error of the numerator that depends on the sample size and the samples variances.
+\color{black}
 
-Similarly, Equation \eqref{eq:eq-gen-d} formalize a general form of the effect size.
+\color{blue}
+
+Equation \eqref{eq:eq-gen-d} describe a general formulation of a standardized effect size measure.
+Where \(b\) is the unstandardized effect size (e.g., the difference between two means) and \(s\) is the standardization term \color{black}.
+For example, in the two-samples case, \(s\) is the pooled standard deviation between the two samples, while in the paired samples case, \(s\) is the standard deviation of the differences for paired samples.
 
@@
 \end{equation}
-
-Where \(b\) is still the unstandardized effect size and \(s\) is the standardization term.
-For example, in the two-sample case, \(s\) is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.
-
-\color{red}
 
@@
 
-NOT SURE TO KEEP THIS!
-
 Another way of conceptualizing the critical effect size is by removing the sample size from the \(t\) statistics (see Equation \eqref{eq:eq-gen-t}).
@@
 
-\bibitem[\citeproctext]{ref-Borenstein2009-mo}
+\leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
 Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
 
-\bibitem[\citeproctext]{ref-Bowley1928-yt}
+\leavevmode\vadjust pre{\hypertarget{ref-Bowley1928-yt}{}}%
 Bowley, A. L. (1928). The standard deviation of the correlation coefficient. \emph{Journal of the American Statistical Association}, \emph{23}, 31. \url{https://doi.org/10.2307/2277400}
 
-\bibitem[\citeproctext]{ref-cohen1994}
+\leavevmode\vadjust pre{\hypertarget{ref-chow1988significance}{}}%
+Chow, S. L. (1988). Significance test or effect size? \emph{Psychological Bulletin}, \emph{103}(1), 105.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
 Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.
 
-\bibitem[\citeproctext]{ref-open2015estimating}
-Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
+\leavevmode\vadjust pre{\hypertarget{ref-cortina1997logic}{}}%
+Cortina, J. M., \& Dunlap, W. P. (1997). On the logic and purpose of significance testing. \emph{Psychological Methods}, \emph{2}(2), 161.
 
-\bibitem[\citeproctext]{ref-Fisher1915-ks}
+\leavevmode\vadjust pre{\hypertarget{ref-Fisher1915-ks}{}}%
 Fisher, R.
@@
 
-\bibitem[\citeproctext]{ref-Hedges1981-za}
+\leavevmode\vadjust pre{\hypertarget{ref-hagen1997praise}{}}%
+Hagen, R. L. (1997). \emph{In praise of the null hypothesis statistical test.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-haig2017tests}{}}%
+Haig, B. D. (2017). Tests of statistical significance made sound. \emph{Educational and Psychological Measurement}, \emph{77}(3), 489--506.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
 Hedges, L.
@@
 
-\bibitem[\citeproctext]{ref-krueger2001null}
+\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
 Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.
 
-\bibitem[\citeproctext]{ref-lakens2013calculating}
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
 Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.
 
-\bibitem[\citeproctext]{ref-lakens2022sample}
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2021practical}{}}%
+Lakens, D. (2021). The practical alternative to the p value is the correctly used p value. \emph{Perspectives on Psychological Science}, \emph{16}(3), 639--648.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
 Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.
 
-\bibitem[\citeproctext]{ref-mccrea2008self}
-McCrea, S.
-M.
-(2008).
-Self-handicapping, excuse making, and counterfactual thinking: Consequences for self-esteem and future motivation.
-\emph{Journal of Personality and Social Psychology}, \emph{95}(2), 274.
+\leavevmode\vadjust pre{\hypertarget{ref-miller2017hypothesis}{}}%
+Miller, J. (2017). Hypothesis testing in the real world. \emph{Educational and Psychological Measurement}, \emph{77}(4), 663--672.
 
-\bibitem[\citeproctext]{ref-morris2002combining}
+\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
 
-\bibitem[\citeproctext]{ref-orben_crud_2020}
+\leavevmode\vadjust pre{\hypertarget{ref-nickerson2000null}{}}%
+Nickerson, R. S. (2000). Null hypothesis significance testing: A review of an old and continuing controversy. \emph{Psychological Methods}, \emph{5}(2), 241.
+
+\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
 
-\bibitem[\citeproctext]{ref-Satterthwaite1946-aq}
+\leavevmode\vadjust pre{\hypertarget{ref-Satterthwaite1946-aq}{}}%
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
 
-\bibitem[\citeproctext]{ref-szucs2017empirical}
+\leavevmode\vadjust pre{\hypertarget{ref-szucs2017empirical}{}}%
 Szucs, D., \& Ioannidis, J.

### 2024-12-16 paper/paper.Rmd
final

@@
 abstract: |
-  Critical effect size values represent the smallest detectable effect that would reach statistical significance given a certain sample size and alpha level.
-It can be useful to calculate the smallest effect that can be statistically significant and evaluate whether such effects are plausible, and if desireable that smaller effects can be detected.
-Reporting critical effect size values may be especially useful while adopting NHST (Null Hypothesis Significance Testing) when the sample size has not been planned a priori, or when there is uncertainty about the expected effect size that can be collected.
-To assist researchers in calculating critical effect size values, we have developed an R package that allows to report the critical effect size values of model parameters.
-The package's functions are implemented for group comparisons, correlations, linear regression, and meta-analysis.
-What we propose could benefit researchers during the planning phase of the study by helping them to understand the limitations and strengths of their research design, and also as a retrospective tool to potentially reframe original interpretations, especially non significant results.
+  Critical effect size values represent the smallest detectable effect that can reach statistical significance given a specific sample size, alpha level and test statistic.
+It can be useful to calculate the critical effect size when designing a study, and evaluate whether such effects are plausible.
+Reporting critical effect size values may be useful when the sample size has not been planned a priori, or when there is uncertainty about the expected sample size that can be collected, when researchers plan to analyze the data with a statistical hypothesis test.
+To assist researchers in calculating critical effect size values we have developed an R package that allows researchers to report critical effect size values for group comparisons, correlations, linear regressions, and meta-analyses.
+Reflecting on critical effect size values could benefit researchers during the planning phase of the study by helping them to understand the limitations of their research design.
+Critical effect size values are also useful when evaluating studies performed by other researchers when a-priori power analyses were not performed, especially when non-significant results are observed.
 
-keywords          : critical effect size values, statistical significance, fucntions implementation 
-wordcount         : "3912"
+keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
+wordcount         : "4676"
 
@@
 
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
 The "critical effect size" refers to the smallest effect size that can be statistically significant, given the test performed, the sample size, and the alpha level [@lakens2022sample].
-Let us consider a study that tests (using for example a two-tailed test) a bivariate correlation with $n = 20$ participants and uses $p < 0.05$ as the threshold for statistical significance.
-In this case, the critical effect sizes are Pearson's $r = -0.44$ and $r = 0.44$.
-As we will argue, critical effect size values can serve as a useful complement to aid in the interpretation of study results, to communicate their relevance, and to help assess their replicability, particularly when statistical significance is relied upon for inference.
-This is especially important in cases where the sample size was not (or could not) be predetermined based on a theory-driven target effect size, or when statistical power may be difficult to determine or deviates from an optimal level.
+Consider a study that tests a bivariate correlation with $n = 20$ participants and uses $\alpha < 0.05$ as the threshold for statistical significance in a two-sided test.
+The critical effect size values for this test are Pearson's $r = -0.44$ and $r = 0.44$, which means that the statistical test will yield a significant result for effects larger than $r = -0.44$ or $r = 0.44$, while observed effect sizes between these two values will lead to a statistically non-significant result.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using significance levels in hypothesis tests to make claims about the presence or absence of effects.
+Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
+Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
 
-Despite longstanding criticisms, Null Hypothesis Significance Testing (NHST) remains a prominent approach for statistical inference in science [@cohen1994; @gigerenzer2004null; @krueger2001null].
-Over the past decades, numerous proposals have emerged to enhance inference based on this approach.
-These include conducting power analyses to determine sample size \textit{a priori} based on theory-informed target effects, and reporting effect sizes and their confidence intervals, accompanied by pertinent related comments, rather than drawing dichotomous conclusions solely based on statistical significance [@transue2019apa].
-Even for the most skeptical, there are suggestions that NHST may possess merits of its own.
-For instance, @wilson2020science propose that it could serve as a useful preliminary filter utilized by "original science" for screening potentially interesting effects, which should subsequently be validated by "replication science".
+Null hypothesis significance testing (NHST) remains the most prominent approach for statistical inference in science, even though there are widespread concerns about the misuse of hypothesis tests [@lakens2021practical; @chow1988significance; @cortina1997logic; @hagen1997praise; @haig2017tests; @miller2017hypothesis; @nickerson2000null; @cohen1994; @gigerenzer2004null; @krueger2001null].
+Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest.
+Researchers often publish studies with low power for medium to small effect sizes [@szucs2017empirical], with the main justification for the sample size being resource limitations [@lakens2022sample].
+However, low power makes it challenging to distinguish signal from noise, and combined with the selective reporting of statistically significant results in the published literature leads to an overestimation of effect sizes [@altoe2020enhancing].
+Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant.
 
-In psychology, statistical power is often severely limited for medium to small effect sizes [@szucs2017empirical], which are generally expected to represent true findings.
-In fact, resource constraints have been authoritatively listed among acceptable justifications for sample size  [@lakens2022sample], and this scenario is arguably widespread in psychology.
-However, limited power makes it challenging to distinguish signal from noise, and on average it results in an overestimation of effect sizes when relying on statistical significance for inference [@altoe2020enhancing].
-On the other hand, very large samples are occasionally available in psychological science, albeit less frequently.
-However, as some have argued, excessively large sample sizes risk creating an "everything-is-significant" scenario, where researchers report and discuss practically negligible effects solely because they are statistically significant, regardless of theoretical relevance or whether they merely reflect minor procedural artifacts [@wilson2020science].
-Thus, while statistical significance can serve as a potentially useful filter, relying solely on it for inference may introduce risks in interpretation.
-In such cases, we contend that reporting critical effect size values can help put into perspective and convey the relevance of the results.
+We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
+In the following sections we will illustrate examples of common scenarios and we will go through the main problems and possible solutions, highlighting the utility of critical effect size values in such contexts.
 
-There are two clear benefits to reporting the critical effect size value(s) for a corresponding test.
+## Small Sample Sizes and Uncertain sample size determination
+
 First, when sample sizes are small, the critical effect size values inform readers about whether the effect sizes that could lead to rejecting the null hypothesis are in line with realistic expectations.
-If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform a meaningful test.
-An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values focuses the attention more strongly on which effect sizes are reasonable to expect.
-Second, in large studies the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
-This will focus the attention of researchers on the difference between statistical significance and practical significance.
-This is especially important in correlational studies on large samples in psychology, where small uncontrolled sources of statistical noise may simultaneously affect and create some shared variance across otherwise unrelated variables, a phenomenon controversially labelled as the 'crud factor' [@orben_crud_2020].
+If the sample size is small and only very large effects would yield a statistically significant result, and the underlying mechanism that is examined is unlikely to lead to such large effect sizes, researchers will realize they are not able to collect sufficient data to perform an informative hypothesis test.
+An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
+In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
+The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
+Power analysis provides information about how likely it is to detect a specific effect size if it truly exists in the population.
+However, it does not indicate the minimum effect size required to reach statistical significance, which could provide additional insights into the strengths and limitations of a study design.
 
 Let us consider the following two scenarios as examples.
-First, imagine researchers conducting a study involving a between-group comparison.
+First, imagine researchers who conduct a study involving a between-group comparison.
 Due to severe resource constraints, they are only able to collect a limited sample size of $n = 30$ ($n = 15$ per group).
-The authors did not perform a power, analysis, but their sample size justification is purely based on how many participants they can collect.
-No statistically significant effect is detected with $p < 0.05$.
-Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though uncertain to some extent.
+The authors did not perform an a-priori power analysis, and their sample size justification is purely based on how many participants they can collect.
+The statistical tests yields a non-significant result, $p > 0.05$.
+Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though without a theoretically expected effect size, it remains uncertain how low the statistical power is.
 Subsequently, they compute the critical effect size values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
-This indicates that any estimated |Cohen's $d| < 0.75$ will certainly fail to reach statistical significance.
-By reporting these critical effect size values, the researchers transparently convey that estimated effects up to a medium-to-large magnitude will consistently fall short of significance.
-In essence, this scenario exemplifies a version of the "winner's curse" [@hedges1984estimation], made explicit.
-In this context, the winner's curse indicates the tendency for a statistically significant initial finding (i.e., a "winner") to be associated with an overestimated effect size which will likely be deflated in subsequent replications (i.e., the "curse"), especially if such initial finding was obtained with a small sample.
-Knowing from the critical effect size values that a specific result must necessarily be associated with either an estimated large effect size or be non-significant, readers are, in principle, prevented from both over-interpreting the magnitude of an effect size and from equating lack of statistical significance with lack of an effect of interest.
+This indicates that any observed Cohen's $d$ between -0.75 and 0.75 will certainly fail to reach statistical significance.
+By reporting these critical effect size values the researchers transparently convey that estimated effects much larger than most effects in the psychological literature will always fall short of significance.
+This means that the hypothesis test will not be informative.
 
-In a second scenario, imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore some bivariate correlations.
-With such a large sample size, researchers are not constrained by sample size limitations and could instead focus on effect sizes.
-However, they legitimately  opt to use statistical significance of a two-tailed test as their primary criterion for whether to comment on effects in their further discussions.
-In this context, the critical effect size values for significance are determined to be correlations of $r = \pm 0.03$.
-This implies that practically any effect differing from zero, regardless of its small magnitude, may easily attain statistical significance.
+To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine a real-world instance drawn from published research.
+The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
+Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+A two-tailed $t$-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+Non-significant effects in such scenarios are likely, even if there is a true effect of interest.
+If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated  [@hedges1984estimation], and this should be taken into account when interpreting the effect size.
+
+Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn with their experimental design, and for those cases where sample size constrains will make it impossible to find significance for a plausible effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multisite studies [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building], should be explored.
+
+## Large Sample Sizes and Meta-analysis
+
+When sample sizes are very large, the critical effect size values will make it clear that trivially small effect sizes will be statistically significant.
+Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
+
+Imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore bivariate correlations between variables.
+With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are $r = \pm 0.03$.
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
-While readers may be tempted to interpret any reported effects simply because they attain statistical significance, signaling the critical effect size value level beforehand can serve as a clear warning that the significance filter might be overly permissive, thus urging interpretive caution.
+
+Consider now another real-world instance involving the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
+With a notably large sample size of $n = 689.003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
+Based on this and other results from the study, the authors conclude that "the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks" [@kramer2014experimental].
+The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
+It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
+The critical effect size should prompt reflection on which effect sizes are truly meaningful in this context, and whether the observed effect size ($d = 0.02$) is theoretically or practically interesting.
 
 A similar scenario may arise in meta-analysis.
-Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, leading even very small average effect sizes to reach significance.
+Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, and as a consequence, even very small average effect sizes can reach significance.
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Thus, in meta-analysis, reporting the critical effect size values can caution readers that the mere attainment of statistical significance by the average effect does not automatically ensure scientific or real-life relevance.
+Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results.
 
-To provide more tangible illustrations of the practical applications of critical effect size values, let us examine two real-world instances drawn from published research.
-The first one features a small sample and has been targeted in the series of replications by the Open Science Collaboration [@open2015estimating], while the second one presents with a very large sample and has already attracted attention in the scientific community.
-The first case pertains to "Study 5" conducted by @mccrea2008self.
-This experiment compared two groups based on their performance percentage in a math test following a preliminary practice session with feedback of failure.
-One group was exposed to failure-excusing, self-handicapping thoughts, while both groups underwent the math test afterward.
-With a modest sample size of 28 participants split into two groups (13 and 15 participants, respectively), a directional one-tailed t-test was conducted on the percentage of correct answers to compare the groups.
-The resulting effect was statistically significant, $t(26) = 1.87$, $p$ (one-tailed) $< .05$.
-Based on the reported data, Cohen's $d = 0.736$, indicating a considerable effect size.
-However, the critical value for Cohen's $d$ in this context was $0.645$, suggesting that, given this sample size, statistical significance is achieved only for estimated effects of substantial magnitude.
-Subsequent replication efforts [@open2015estimating] reaffirmed the significance of the effect, and used a somewhat larger sample ($n = 61$), $t(59)=2.325$.
-The estimated effect size was somehow lower than that originally reported, however, $d = 0.605$ (here, critical d = 0.428).
-The second scenario involves the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of $n = 689'003$, the researchers observed a significant effect worsened emotional states when positive posts were reduced, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
-Based on this and other results, the authors conclude that emotions expressed by other users on Facebook influence our own emotions.
-Despite the statistical significance, however, the critical values for Cohen's $d$ in this case are $-0.006$ and $+0.006$, emphasizing the importance of considering effect size beyond mere significance.
-This prompts reflection on whether an effect of such magnitude on such a large sample (actual $d = 0.02$) is truly meaningful in individuals' lives.
+Lastly, routinely reasoning about critical effect sizes, alongside the use of commonly known practices to enhance the quality of research (i.e.
+power analysis, design analysis, data simulation etc.), will bring benefits to researchers by making clear that significance is strictly related to the sample size, the alpha level and the statistic test used to analyse the data.
+Nevertheless, in educational settings, this concept will help students to better grasp such concepts and therefore give a more critical approach to published research and for their future studies.
 
@@
 
-In this section, we provide guidance and formulas for computing critical effect size values in general, with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's r), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing standardized critical effect sizes with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's $r$), and raw and standardized coefficients in linear models.
+The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
 These formulas have been incorporated into R functions of the package `criticalESvalue` accessible at: [https://github.com/psicostat/criticalESvalue](https://github.com/psicostat/criticalESvalue), and are elaborated upon in the subsequent section.
@@
 Where $b$ is the unstandardized effect size that depends on the type of test.
-For example, in the two sample t-test or the single mean for a one sample t-test.
-The denominator is the standard error of the numerator that depends on the sample size and the sample(s) variance.
+For example, in the two sample $t$-test the numerator is the difference between the two sample means while in the one-sample case is the difference between the sample mean and the population value.
+The denominator is the standard error of the numerator that depends on the sample size and the samples variances.
 
-Similarly, Equation \@ref(eq:eq-gen-d) formalize a general form of the effect size.
+Equation \@ref(eq:eq-gen-d) describes a general formulation of a standardized effect size measure.
+Where $b$ is the unstandardized effect size (e.g., the difference between two means) and $s$ is the standardization term.
+For example, in the two-samples case, $s$ is the pooled standard deviation between the two samples, while in the paired samples case, $s$ is the standard deviation of the differences for paired samples.
 
@@
 \end{equation}
-
-Where $b$ is still the unstandardized effect size and $s$ is the standardization term.
-For example, in the two-sample case, $s$ is the pooled standard deviation between the two samples or the standard deviation of the differences for paired samples.
 
@@
 
-NOT SURE TO KEEP THIS!
-
 Another way of conceptualizing the critical effect size is by removing the sample size from the $t$ statistics (see Equation \@ref(eq:eq-gen-t)).
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical effect size values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients.
+Hence, we chose to expand the application of computing critical effect size values beyond Cohen's $d$ and correlation to include linear regression with both raw and standardized coefficients and meta-analysis.
 This serves as a first step in computing critical effect size values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.

### 2024-12-16 paper/paper.tex
final

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 3912}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4676}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \abstract{%
-Critical effect size values represent the smallest detectable effect that can reach statistical significance given a specific sample size and alpha level.
+Critical effect size values represent the smallest detectable effect that can reach statistical significance given a specific sample size, alpha level and test statistic.
 It can be useful to calculate the critical effect size when designing a study, and evaluate whether such effects are plausible.
@@
 Reflecting on critical effect size values could benefit researchers during the planning phase of the study by helping them to understand the limitations of their research design.
-Critical effect size values are also useful when evaluating studies performed by other researchers when a-priori power analyses were not performed, especially when non-significant results were observed.
+Critical effect size values are also useful when evaluating studies performed by other researchers when a-priori power analyses were not performed, especially when non-significant results are observed.
 }
@@
 The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects larger than \(r = -0.44\) or \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference use significance levels in hypothesis tests to make claims about the presence or absence of effects.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using significance levels in hypothesis tests to make claims about the presence or absence of effects.
 Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-\textcolor{red}{In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
-The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.}
+In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
+The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
+Power analysis provides information about how likely it is to detect a specific effect size if it truly exists in the population.
+However, it does not indicate the minimum effect size required to reach statistical significance, which could provide additional insights into the strengths and limitations of a study design.
 
@@
 
-\textcolor{red}{To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine two real-world instances drawn from published research.
-The first study was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine a real-world instance drawn from published research.
+The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (Weiss et al.
+(2014)).
 The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
 Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
-A two-tailed $t$-test was conducted to compare the groups.
-The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
-But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
-Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
+A two-tailed \(t\)-test was conducted to compare the groups.
+The resulting effect was statistically significant for the maximal span, \(t(55) = -2.5\), \(p = .017\), the Cohen's d calculated on summary statistics is of \(-.74\), critical \(d = \pm .62\).
+But for the total number of sequences the test did not reach statistical significance (\(t(55) = -1.8\), \(p = 0.076\)), actual \(d = -0.31\) and critical \(d\) again \(\pm.62\).
+Thus, given the limited sample size, a result must be of rather large magnitude to be associated with \(p < 0.05\), while medium to small effect sizes will simply fail to reach significance.
 This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
-} Non-significant effects are likely, even if there is a true effect of interest.
+Non-significant effects in such scenarios are likely, even if there is a true effect of interest.
 If a significant effect is observed, but most effects in a field are smaller than a study can detect, it is probable the effect size is overestimated (Hedges, 1984), and this should be taken into account when interpreting the effect size.
-\textcolor{red}{Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn, and for those cases where sample size constrains will make it impossible to find an effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multi-lab studies (\cite{AAA}), should be explored.}
+
+Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn with their experimental design, and for those cases where sample size constrains will make it impossible to find significance for a plausible effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multisite studies (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023), should be explored.
 
@@
 
-The second real-world instance involves the study by Kramer et al.
+Consider now another real-world instance involving the study by Kramer et al.
 (2014), which explored the impact of emotional content on Facebook users' experiences.
 With a notably large sample size of \(n = 689.003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
-Based on this and other results from the study, the authors conclude that \textcolor{red}{"the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks".} The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
+Based on this and other results from the study, the authors conclude that ``the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks'' (Kramer et al., 2014).
+The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
 It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
@@
 
+Lastly, routinely reasoning about critical effect sizes, alongside the use of commonly known practices to enhance the quality of research (i.e.~power analysis, design analysis, data simulation etc.), will bring benefits to researchers by making clear that significance is strictly related to the sample size, the alpha level and the statistic test used to analyse the data.
+Nevertheless, in educational settings, this concept will help students to better grasp such concepts and therefore give a more critical approach to published research and for their future studies.
+
 \hypertarget{how-to-compute-critical-effect-size-values}{%
@@
 
-\color{blue} In this section, we provide guidance and formulas for computing standardized critical effect size with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
+In this section, we provide guidance and formulas for computing standardized critical effect sizes with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
 The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
-\color{black} These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
+These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
 
@@
 Where \(b\) is the unstandardized effect size that depends on the type of test.
-\color{blue} For example, in the two sample \(t\)-test the numerator is the difference between the two sample means while in the one-sample case is the difference between the sample mean and the population value.
+For example, in the two sample \(t\)-test the numerator is the difference between the two sample means while in the one-sample case is the difference between the sample mean and the population value.
 The denominator is the standard error of the numerator that depends on the sample size and the samples variances.
-\color{black}
 
-\color{blue}
-
-Equation \eqref{eq:eq-gen-d} describe a general formulation of a standardized effect size measure.
-Where \(b\) is the unstandardized effect size (e.g., the difference between two means) and \(s\) is the standardization term \color{black}.
+Equation \eqref{eq:eq-gen-d} describes a general formulation of a standardized effect size measure.
+Where \(b\) is the unstandardized effect size (e.g., the difference between two means) and \(s\) is the standardization term.
 For example, in the two-samples case, \(s\) is the pooled standard deviation between the two samples, while in the paired samples case, \(s\) is the standard deviation of the differences for paired samples.
@@
 Real-case scenarios may not always be that simple.
-Hence, we chose to expand the application of computing critical effect size values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients.
+Hence, we chose to expand the application of computing critical effect size values beyond Cohen's \(d\) and correlation to include linear regression with both raw and standardized coefficients and meta-analysis.
 This serves as a first step in computing critical effect size values for a wider array of effects encountered in practical scenarios, where linear models and their extensions are commonly utilized for modeling purposes.
@@
 Bowley, A. L. (1928). The standard deviation of the correlation coefficient. \emph{Journal of the American Statistical Association}, \emph{23}, 31. \url{https://doi.org/10.2307/2277400}
+
+\leavevmode\vadjust pre{\hypertarget{ref-byers2020building}{}}%
+Byers-Heinlein, K., Bergmann, C., Davies, C., Frank, M.
+C., Hamlin, J.
+K., Kline, M., Kominsky, J.
+F., Kosie, J.
+E., Lew-Williams, C., Liu, L., et al.
+(2020).
+Building a collaborative psychological science: Lessons learned from ManyBabies 1.
+\emph{Canadian Psychology/Psychologie Canadienne}, \emph{61}(4), 349.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-jarke2022roadmap}{}}%
+Jarke, H., Anand-Vembar, S., Alzahawi, S., Andersen, T.
+L., Bojanić, L., Carstensen, A., Feldman, G., Garcia-Garzon, E., Kapoor, H., Lewis, S., et al.
+(2022).
+A roadmap to large-scale multi-country replications in psychology.
+\emph{Collabra: Psychology}, \emph{8}(1), 57538.
+
 \leavevmode\vadjust pre{\hypertarget{ref-kramer2014experimental}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-moshontz2018psychological}{}}%
+Moshontz, H., Campbell, L., Ebersole, C.
+R., IJzerman, H., Urry, H.
+L., Forscher, P.
+S., Grahe, J.
+E., McCarthy, R.
+J., Musser, E.
+D., Antfolk, J., et al.
+(2018).
+The psychological science accelerator: Advancing psychology through a distributed collaborative network.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(4), 501--515.
+
 \leavevmode\vadjust pre{\hypertarget{ref-nickerson2000null}{}}%
@@
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sirois2023pupil}{}}%
+Sirois, S., Brisson, J., Blaser, E., Calignano, G., Donenfeld, J., Hepach, R., Hochmann, J.-R., Kaldy, Z., Liszkowski, U., Mayer, M., et al.
+(2023).
+The pupil collaboration: A multi-lab, multi-method analysis of goal attribution in infants.
+\emph{Infant Behavior and Development}, \emph{73}, 101890.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-weiss2014spatial}{}}%
+Weiss, A. H., Biron, T., Lieder, I., Granot, R. Y., \& Ahissar, M. (2014). Spatial vision is superior in musicians when memory plays a role. \emph{Journal of Vision}, \emph{14}(9), 18--18.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Welch1947-wh}{}}%

### 2024-12-16 paper/paper.Rmd
final

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4676"
+wordcount         : "4745"
 
@@
 
-Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn with their experimental design, and for those cases where sample size constrains will make it impossible to find significance for a plausible effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multisite studies [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building], should be explored.
+Beyond critical effect size values, researchers should carefully consider, before conducting a study, what conclusions can realistically be drawn from their experimental design.
+The more complex the study design, the higher the critical effect size value.
+Researchers should assess whether their available sample size allows them to address a simpler yet still relevant question.
+If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building], should be considered.
+Lastly, if participating in or leading a multisite study is not an option, researchers should evaluate whether collecting the sample for exploratory analysis without making inferences might still be valuable.
+Such data could later be included in a meta-analysis.
 

### 2024-12-16 paper/paper.tex
final

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4676}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4745}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 
-Beyond critical effect size values, researchers should reason before conducting a study on what conclusions can be drawn with their experimental design, and for those cases where sample size constrains will make it impossible to find significance for a plausible effect because it is likely smaller than the critical effect size value, alternative options for data collection, such as multisite studies (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023), should be explored.
+Beyond critical effect size values, researchers should carefully consider, before conducting a study, what conclusions can realistically be drawn from their experimental design.
+The more complex the study design, the higher the critical effect size value.
+Researchers should assess whether their available sample size allows them to address a simpler yet still relevant question.
+If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023), should be considered.
+Lastly, if participating in or leading a multisite study is not an option, researchers should evaluate whether collecting the sample for exploratory analysis without making inferences might still be valuable.
+Such data could later be included in a meta-analysis.
 

### 2024-12-16 paper/paper.Rmd
final

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4745"
+wordcount         : "4751"
 
@@
 Beyond enhancing study design and statistical inferences based on hypothesis tests, reporting critical effect size values can also serve an educational purpose.
-It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size.
+It underscores how the distinction between a significant and non-significant result is not solely determined by the presence or absence of a true effect, but also by the sample size and the complexity of the analysis.
 By highlighting a critical effect size value, researchers can become more aware of the possibility of Type 2 errors when results are non-significant.

### 2024-12-16 paper/paper.tex
final

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4745}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4751}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}

### 2025-01-02 paper/paper.Rmd
latest version

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4751"
+wordcount         : "4761"
 
@@
 Consider a study that tests a bivariate correlation with $n = 20$ participants and uses $\alpha < 0.05$ as the threshold for statistical significance in a two-sided test.
-The critical effect size values for this test are Pearson's $r = -0.44$ and $r = 0.44$, which means that the statistical test will yield a significant result for effects larger than $r = -0.44$ or $r = 0.44$, while observed effect sizes between these two values will lead to a statistically non-significant result.
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using significance levels in hypothesis tests to make claims about the presence or absence of effects.
+The critical effect size values for this test are Pearson's $r = -0.44$ and $r = 0.44$, which means that the statistical test will yield a significant result for effects more extreme than $r = -0.44$ or $r = 0.44$, while observed effect sizes between these two values will lead to a statistically non-significant result.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make statistical inferences using significance levels in hypothesis tests to make claims about the presence or absence of effects.
 Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
-Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
+Reporting critical effect size values is especially important when the sample size was not (or could not) predetermined on the basis of a smallest effect size of interest.
 
@@
 We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
-In the following sections we will illustrate examples of common scenarios and we will go through the main problems and possible solutions, highlighting the utility of critical effect size values in such contexts.
+In the following sections we will illustrate examples of common scenarios and we will go through the main problems of certain research scenarios and possible solutions, highlighting the utility of critical effect size values in such contexts.
 
-## Small Sample Sizes and Uncertain sample size determination
+## Small Sample Sizes and Uncertain Sample Size Setermination
 
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
-The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
+In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis [@altoe2020enhancing], but this would require both knowledge of the plausible effect size and/or the smallest effect size of interest (SESOI), which are not always easily to determine.
+The use of critical effect size values can be used in a simple an efficient way to evaluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
 Power analysis provides information about how likely it is to detect a specific effect size if it truly exists in the population.
@@
 The authors did not perform an a-priori power analysis, and their sample size justification is purely based on how many participants they can collect.
-The statistical tests yields a non-significant result, $p > 0.05$.
+The statistical test yields a non-significant result, $p > 0.05$.
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though without a theoretically expected effect size, it remains uncertain how low the statistical power is.
@@
 To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine a real-world instance drawn from published research.
-The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (@weiss2014spatial).
+The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected [@weiss2014spatial].
 The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
@@
 A two-tailed $t$-test was conducted to compare the groups.
-The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's d calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's $d$ calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
 But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
 Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
-This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+This case makes it obvious that a failure to find a significant result does not imply the absence of an effect, while a statistically significant effect may represent an overestimation.
 Non-significant effects in such scenarios are likely, even if there is a true effect of interest.
@@
 
-Beyond critical effect size values, researchers should carefully consider, before conducting a study, what conclusions can realistically be drawn from their experimental design.
+Beyond critical effect size values, researchers should reason before conducting a study what conclusions can realistically be drawn from their experimental design.
 The more complex the study design, the higher the critical effect size value.
 Researchers should assess whether their available sample size allows them to address a simpler yet still relevant question.
-If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building], should be considered.
+If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies should be explored [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building].
 Lastly, if participating in or leading a multisite study is not an option, researchers should evaluate whether collecting the sample for exploratory analysis without making inferences might still be valuable.
@@
 Consider now another real-world instance involving the study by @kramer2014experimental, which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of $n = 689.003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310,044) = −5.63$, $p < 0.001$ (non-directional test).
+With a notably large sample size of $n = 689,003$, the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, $t(310.044) = −5.63$, $p < 0.001$ (non-directional test).
 Based on this and other results from the study, the authors conclude that "the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks" [@kramer2014experimental].
@@
 
-A similar scenario may arise in meta-analysis.
-Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, and as a consequence, even very small average effect sizes can reach significance.
+A similar scenario may arise in a meta-analysis.
+Despite potential loss of precision due to substantial heterogeneity across effect sizes in different studies, meta-analyses typically synthesize a large amount of evidence, and as a consequence, even very small average effect sizes can reach statistical significance.
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
@@
 Lastly, routinely reasoning about critical effect sizes, alongside the use of commonly known practices to enhance the quality of research (i.e.
-power analysis, design analysis, data simulation etc.), will bring benefits to researchers by making clear that significance is strictly related to the sample size, the alpha level and the statistic test used to analyse the data.
-Nevertheless, in educational settings, this concept will help students to better grasp such concepts and therefore give a more critical approach to published research and for their future studies.
+power analysis, design analysis, data simulation etc.), will bring benefits to researchers by making it clear that significance is strictly related to the sample size, the alpha level and the statistic test used to analyse the data.
+Nevertheless, in educational settings, critical effect size values will help students to better grasp such concepts and therefore give a more critical approach to published research and for their future studies.
 
@@
 In this section, we provide guidance and formulas for computing standardized critical effect sizes with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's $r$), and raw and standardized coefficients in linear models.
-The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
+The same equations can be used to calculate unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
 These formulas have been incorporated into R functions of the package `criticalESvalue` accessible at: [https://github.com/psicostat/criticalESvalue](https://github.com/psicostat/criticalESvalue), and are elaborated upon in the subsequent section.
 
-## t-test
+## $t$-test
 
-For the t-test we considered the two-sample, paired, and one-sample tests. As a general approach the $t$ statistic is computed as reported in Equation \@ref(eq:eq-gen-t).
+For the $t$-test we considered the two-sample, paired, and one-sample tests. As a general approach the $t$ statistic is computed as reported in Equation \@ref(eq:eq-gen-t).
 
@@
 
-The critical $t_c$ is the test statistic associated with a p-value equal or lower to $\alpha$.
+The critical $t_c$ is the test statistic associated with a $p$-value equal or lower to $\alpha$.
 Substituting $t_c$ into Equation \@ref(eq:eq-gen-t) we realize that there is a $b_c$ that with a certain $SE_b$ produce the critical test statistics.
@@
 The only caveat is that calculating $d$ or $d_c$ in this way assume that the standard deviation used to calculate the test statistics and the effect size is the same.
-While this is true most of the time, as explained in the paired t-test section, there are more than one methods to calculate the (critical) effect size using different standard deviations.
-Furthemore, this method assume that the two groups (in case of a two-sample t-test) have the same size.
+While this is true most of the time, as explained in the paired $t$-test section, there are more than one methods to calculate the (critical) effect size using different standard deviations.
+Furthemore, this method assumes that the two groups (in case of a two-sample $t$-test) have the same size.
 
-### One sample t-test
+### One sample $t$-test
 
@@
 
-### Two-sample t-test
+### Two-sample $t$-test
 
@@
 When assuming homogeneity of the variances between the two groups we have $n_1 + n_2 - 2$ degrees of freedom to calculate $t_c$ and $s$ is the pooled standard deviation ($s_p$) calculated using Equation \@ref(eq:eq-s-p-twsmp).
-The $s_p$ can be intepreted as the square root of the weighted average of the samples standard deviations.
+The $s_p$ can be interpreted as the square root of the weighted average of the samples standard deviations.
 The effect size $d$ is calculated as $\frac{b}{s_p}$ and the critical effect size can be calculated as $d_c = \frac{b_c}{s_p}$ with $b_c = t_c \times SE_b$.
@@
 
-When relaxing the assumption of equal variances (i.e., using the Welch's t-test) $s$ is calculated as the square root of the average between the two samples variances, $SE_b$ is calculated as $\sqrt{\frac{s^2_{x_1}}{n_1} + \frac{s^2_{x_2}}{n_2}}$.
-The degress of freedom for $t_c$ are calculated in a more complex way using the Welch-Satterthwaite equation [@Welch1947-wh; @Satterthwaite1946-aq].
-For the Welch's t-test we can calculate the critical effect size removing the sample size from $t_c$ (as done above) only when $n_1 = n_2$.
+When relaxing the assumption of equal variances (i.e., using the Welch's $t$-test) $s$ is calculated as the square root of the average between the two samples variances, $SE_b$ is calculated as $\sqrt{\frac{s^2_{x_1}}{n_1} + \frac{s^2_{x_2}}{n_2}}$.
+The degrees of freedom for $t_c$ are calculated in a more complex way using the Welch-Satterthwaite equation [@Welch1947-wh; @Satterthwaite1946-aq].
+For the Welch's $t$-test we can calculate the critical effect size removing the sample size from $t_c$ (as done above) only when $n_1 = n_2$.
 
-### Paired sample t-test
+### Paired sample $t$-test
 
 For the paired sample case the situation is less straightforward.
-The reason is that the $t$ statistic is computed using $b$ being the average of the paired differences between the two (paired) samples and $s$ being the standard deviation of the differences.
-Using these values, we can just compute the critical effects size as for the one-sample case, in fact a paired t-test is a one-sample test on the vector of paired differences.
+The reason is that the $t$ statistic is computed with $b$ being the average of the paired differences between the two (paired) samples and $s$ being the standard deviation of the differences.
+Using these values, we can just compute the critical effects size as for the one-sample case, as a paired $t$-test is a one-sample test on the vector of paired differences.
 Test statistics and effect size based on the standard deviation of differences (the so called $d_z$) are used for hypothesis testing and power analysis [@lakens2013calculating].
@@
 For this reason, there is a corrected version of the effect size called Hedges's $g$ [@Hedges1981-za, @Viechtbauer2007-tq].
-The Hedges's $g$ can be calculated for all the t-test scenarios.
+The Hedges's $g$ can be calculated for all the $t$-test scenarios.
 The Hegdes's correction is implemented in Equation \@ref(eq:eq-g-correction) where $\Gamma$ is the gamma function and $m$ the degrees of freedom.
@@
 
-The correction is applied as $g = c(m) \times d$. The hypothesis testing is not taking into account the Hedges's correction thus the critical $g$ will be different from the critical $d$.
+The correction is applied as $g = c(m) \times d$. The hypothesis testing does not take the Hedges's correction into account, thus the critical $g$ will be different from the critical $d$.
 
@@
 Hypothesis testing for the Pearson's correlation coefficient is usually done using the t statistics or using a z statistics.
-The general approach presented for the t-test is still valid.
+The general approach presented for the $t$-test is still valid.
 The only difference is that the correlation is already a standardized effect size thus there is no need of the standardization term $s$.
-Equation \@ref(eq:eq-gen-t) can be used substituing $b$ with $r$ (the sample correlation coefficient) as shown in Equation \@ref(eq:eq-rc).
+Equation \@ref(eq:eq-gen-t) can be used substituting $b$ with $r$ (the sample correlation coefficient) as shown in Equation \@ref(eq:eq-rc).
 The standard error is calculated as $SE_r = \sqrt{\frac{1 - r^2}{n - 2}}$ [@Bowley1928-yt].
@@
 Hypothesis testing on regression coefficients (e.g., using the \texttt{lm} function in R) is performed using the so-called Wald test.
-The test can be considered a one-sample t-test where the $t$ value is calculated dividing the regression parameter $\beta_0, \dots, \beta_j$ by its standard error $SE_{\beta_j}$.
-The critical $t$ value is calculated from a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept.
+The test can be considered a one-sample $t$-test where the $t$-value is calculated dividing the regression parameter $\beta_0, \dots, \beta_j$ by its standard error $SE_{\beta_j}$.
+The critical $t$-value is calculated from a $t$ distribution using $n - p - 1$ degrees of freedom where $n$ is the number of observations and $p$ is the number of coefficients beyond the intercept.
 We can substitute $b$ with $\beta_j$ and $SE_{b}$ with $SE_{\beta_j}$ in Equation \@ref(eq:eq-gen-crit) to find the critical unstandardized regression coefficient ($\beta_{c_j} = t_c \times SE_{\beta_j}$).
-There is no an unique way to calculate the standardized version of the critical regression coefficient.
+There is no unique way to calculate the standardized version of the critical regression coefficient.
 The main reason is that in regression modelling we can include categorical and numerical independent variables with main effects and interactions.
@@
 Another complication is that in some cases, researchers standardize the dependent variable, the independent variables or both.
-In general, the unstandardized critical regression coefficient can be usually also be easily interpreted.
+In general, the unstandardized critical regression coefficient can usually also be easily interpreted.
 For example, assuming we are regressing some scores on a cognitive tests using the age of participant,the critical regression coefficient is the minimum increase in the cognitive test for a unit age increase that would be significant.
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package "criticalESvalue" in R.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package `criticalESvalue` in R.
 The complete package can be accessed at: https://github.com/psicostat/criticalESvalue.
-Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
+Here, we demonstrate its application through two examples: one example of a $t$-test on real data and a computation of the critical effect size values for a correlation from sample size.
 It is important to note that, in general, depending on the researcher's hypotheses, the package allows for the calculation of either two critical effect size values for two-tailed tests or a single critical effect size value for one-tailed tests, for which the direction must be specified.
-In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients.
+In the Supplementary online materials additional examples can be found for correlation, $t$-test, paired $t$-test, linear models and regression coefficients.
 
-First the package should be downloaded and opened with the library function:
+First the package should be downloaded and opened with the `library` function:
 
@@
 
-For our examples on real data we used from the package 'psych' the dataset "holzinger.swineford" which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects.
-Once the package is retrieved with 'library', the dataset can be opened using 'data("name of the dataset")'.
+For our examples on real data we used from the package `psych` the dataset "holzinger.swineford" which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects.
+Once the package is retrieved with `library`, the dataset can be opened using `data`("name of the dataset").
 For simplicity we decided to rename it with a shorter name.
@@
 
-We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the 't.test' function:
+We want to know the critical value for a $t$-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the `t.test` function:
 
@@
 
-The direction of the hypothesis and the test to apply, either $t$-test or $z$-test, should be specified.
-The output will return the critical correlation value(s), the degrees of freedom and the type of test used.
+The direction of the hypothesis and the test to apply, either the $t$-test or $z$-test, should be specified.
+The output will return the critical correlation value(s), the degrees of freedom, and the type of test used.
 
@@
 In cases where optimal power can be attained with a sufficiently large sample size for an effect of a specific magnitude of interest, and this is truly determined \textit{a priori}, the interpretation of both significance and non-significance becomes straightforward.
-However, when power analysis did not inform the sample size or when power is likely but undeterminedly low, reporting critical effect size values for the obtained sample can help provide context for interpretation.
+However, when power analysis did not inform the sample size or when power is likely but indeterminably low, reporting critical effect size values for the obtained sample can help provide context for interpretation.
 Critical effect size values can be computed and interpreted even retrospectively or for studies that have already been published.

### 2025-01-02 paper/paper.tex
latest version

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4751}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4761}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 Consider a study that tests a bivariate correlation with \(n = 20\) participants and uses \(\alpha < 0.05\) as the threshold for statistical significance in a two-sided test.
-The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects larger than \(r = -0.44\) or \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
-In the present paper, we suggest that researchers should systematically report critical effect size values when they make inference using significance levels in hypothesis tests to make claims about the presence or absence of effects.
+The critical effect size values for this test are Pearson's \(r = -0.44\) and \(r = 0.44\), which means that the statistical test will yield a significant result for effects more extreme than \(r = -0.44\) or \(r = 0.44\), while observed effect sizes between these two values will lead to a statistically non-significant result.
+In the present paper, we suggest that researchers should systematically report critical effect size values when they make statistical inferences using significance levels in hypothesis tests to make claims about the presence or absence of effects.
 Critical effect size values can serve as a useful complement for interpreting study results, communicate the informational value of a study design, and contextualize the interpretation of non-significant results, particularly when statistical significance is relied upon for inference.
-Reporting critical effect size values is especially important when the sample size was not (or could not) be predetermined on the basis of a smallest effect size of interest.
+Reporting critical effect size values is especially important when the sample size was not (or could not) predetermined on the basis of a smallest effect size of interest.
 
@@
 We believe there are two clear benefits to reporting critical effect size values for a corresponding test.
-In the following sections we will illustrate examples of common scenarios and we will go through the main problems and possible solutions, highlighting the utility of critical effect size values in such contexts.
+In the following sections we will illustrate examples of common scenarios and we will go through the main problems of certain research scenarios and possible solutions, highlighting the utility of critical effect size values in such contexts.
 
-\hypertarget{small-sample-sizes-and-uncertain-sample-size-determination}{%
-\subsection{Small Sample Sizes and Uncertain sample size determination}\label{small-sample-sizes-and-uncertain-sample-size-determination}}
+\hypertarget{small-sample-sizes-and-uncertain-sample-size-setermination}{%
+\subsection{Small Sample Sizes and Uncertain Sample Size Setermination}\label{small-sample-sizes-and-uncertain-sample-size-setermination}}
 
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis, but this would require both knowledge on the plausible effect size and/or on the smallest effect size of interest (SESOI), which are not always easly to determine.
-The use of critical effect size values can be used in a simple an efficient way to evluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
+In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis (Altoè et al., 2020), but this would require both knowledge of the plausible effect size and/or the smallest effect size of interest (SESOI), which are not always easily to determine.
+The use of critical effect size values can be used in a simple an efficient way to evaluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
 Power analysis provides information about how likely it is to detect a specific effect size if it truly exists in the population.
@@
 The authors did not perform an a-priori power analysis, and their sample size justification is purely based on how many participants they can collect.
-The statistical tests yields a non-significant result, \(p > 0.05\).
+The statistical test yields a non-significant result, \(p > 0.05\).
 Given their prior expectation that the effect of interest may not be large, they acknowledge that their study was likely underpowered, though without a theoretically expected effect size, it remains uncertain how low the statistical power is.
@@
 To provide a more tangible illustration of the practical applications of the critical effect sizes, let us examine a real-world instance drawn from published research.
-The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (Weiss et al.
-(2014)).
+The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (Weiss et al., 2014).
 The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
@@
 A two-tailed \(t\)-test was conducted to compare the groups.
-The resulting effect was statistically significant for the maximal span, \(t(55) = -2.5\), \(p = .017\), the Cohen's d calculated on summary statistics is of \(-.74\), critical \(d = \pm .62\).
+The resulting effect was statistically significant for the maximal span, \(t(55) = -2.5\), \(p = .017\), the Cohen's \(d\) calculated on summary statistics is of \(-.74\), critical \(d = \pm .62\).
 But for the total number of sequences the test did not reach statistical significance (\(t(55) = -1.8\), \(p = 0.076\)), actual \(d = -0.31\) and critical \(d\) again \(\pm.62\).
 Thus, given the limited sample size, a result must be of rather large magnitude to be associated with \(p < 0.05\), while medium to small effect sizes will simply fail to reach significance.
-This case makes it obvious that a failure to find a significant result does not imply absence of an effect, while a statistically significant effect may represent an overestimation.
+This case makes it obvious that a failure to find a significant result does not imply the absence of an effect, while a statistically significant effect may represent an overestimation.
 Non-significant effects in such scenarios are likely, even if there is a true effect of interest.
@@
 
-Beyond critical effect size values, researchers should carefully consider, before conducting a study, what conclusions can realistically be drawn from their experimental design.
+Beyond critical effect size values, researchers should reason before conducting a study what conclusions can realistically be drawn from their experimental design.
 The more complex the study design, the higher the critical effect size value.
 Researchers should assess whether their available sample size allows them to address a simpler yet still relevant question.
-If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023), should be considered.
+If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies should be explored (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023).
 Lastly, if participating in or leading a multisite study is not an option, researchers should evaluate whether collecting the sample for exploratory analysis without making inferences might still be valuable.
@@
 (2014), which explored the impact of emotional content on Facebook users' experiences.
-With a notably large sample size of \(n = 689.003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310,044) = −5.63\), \(p < 0.001\) (non-directional test).
+With a notably large sample size of \(n = 689,003\), the researchers observed a statistically significant increase in the number of negative words typed by participants when positive posts were reduced in their timeline, \(t(310.044) = −5.63\), \(p < 0.001\) (non-directional test).
 Based on this and other results from the study, the authors conclude that ``the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks'' (Kramer et al., 2014).
@@
 
-Lastly, routinely reasoning about critical effect sizes, alongside the use of commonly known practices to enhance the quality of research (i.e.~power analysis, design analysis, data simulation etc.), will bring benefits to researchers by making clear that significance is strictly related to the sample size, the alpha level and the statistic test used to analyse the data.
-Nevertheless, in educational settings, this concept will help students to better grasp such concepts and therefore give a more critical approach to published research and for their future studies.
+Lastly, routinely reasoning about critical effect sizes, alongside the use of commonly known practices to enhance the quality of research (i.e.~power analysis, design analysis, data simulation etc.), will bring benefits to researchers by making it clear that significance is strictly related to the sample size, the alpha level and the statistic test used to analyse the data.
+Nevertheless, in educational settings, critical effect size values will help students to better grasp such concepts and therefore give a more critical approach to published research and for their future studies.
 
@@
 In this section, we provide guidance and formulas for computing standardized critical effect sizes with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
-The same equations can be used to calculated unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
+The same equations can be used to calculate unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
 These formulas have been incorporated into R functions of the package \texttt{criticalESvalue} accessible at: \url{https://github.com/psicostat/criticalESvalue}, and are elaborated upon in the subsequent section.
@@
 \hypertarget{t-test}{%
-\subsection{t-test}\label{t-test}}
+\subsection{\texorpdfstring{\(t\)-test}{t-test}}\label{t-test}}
 
-For the t-test we considered the two-sample, paired, and one-sample tests. As a general approach the \(t\) statistic is computed as reported in Equation \eqref{eq:eq-gen-t}.
+For the \(t\)-test we considered the two-sample, paired, and one-sample tests. As a general approach the \(t\) statistic is computed as reported in Equation \eqref{eq:eq-gen-t}.
 
@@
 
-The critical \(t_c\) is the test statistic associated with a p-value equal or lower to \(\alpha\).
+The critical \(t_c\) is the test statistic associated with a \(p\)-value equal or lower to \(\alpha\).
 Substituting \(t_c\) into Equation \eqref{eq:eq-gen-t} we realize that there is a \(b_c\) that with a certain \(SE_b\) produce the critical test statistics.
@@
 The only caveat is that calculating \(d\) or \(d_c\) in this way assume that the standard deviation used to calculate the test statistics and the effect size is the same.
-While this is true most of the time, as explained in the paired t-test section, there are more than one methods to calculate the (critical) effect size using different standard deviations.
-Furthemore, this method assume that the two groups (in case of a two-sample t-test) have the same size.
+While this is true most of the time, as explained in the paired \(t\)-test section, there are more than one methods to calculate the (critical) effect size using different standard deviations.
+Furthemore, this method assumes that the two groups (in case of a two-sample \(t\)-test) have the same size.
 
 \hypertarget{one-sample-t-test}{%
-\subsubsection{One sample t-test}\label{one-sample-t-test}}
+\subsubsection{\texorpdfstring{One sample \(t\)-test}{One sample t-test}}\label{one-sample-t-test}}
 
@@
 \hypertarget{two-sample-t-test}{%
-\subsubsection{Two-sample t-test}\label{two-sample-t-test}}
+\subsubsection{\texorpdfstring{Two-sample \(t\)-test}{Two-sample t-test}}\label{two-sample-t-test}}
 
@@
 When assuming homogeneity of the variances between the two groups we have \(n_1 + n_2 - 2\) degrees of freedom to calculate \(t_c\) and \(s\) is the pooled standard deviation (\(s_p\)) calculated using Equation \eqref{eq:eq-s-p-twsmp}.
-The \(s_p\) can be intepreted as the square root of the weighted average of the samples standard deviations.
+The \(s_p\) can be interpreted as the square root of the weighted average of the samples standard deviations.
 The effect size \(d\) is calculated as \(\frac{b}{s_p}\) and the critical effect size can be calculated as \(d_c = \frac{b_c}{s_p}\) with \(b_c = t_c \times SE_b\).
@@
 
-When relaxing the assumption of equal variances (i.e., using the Welch's t-test) \(s\) is calculated as the square root of the average between the two samples variances, \(SE_b\) is calculated as \(\sqrt{\frac{s^2_{x_1}}{n_1} + \frac{s^2_{x_2}}{n_2}}\).
-The degress of freedom for \(t_c\) are calculated in a more complex way using the Welch-Satterthwaite equation (Satterthwaite, 1946; Welch, 1947).
-For the Welch's t-test we can calculate the critical effect size removing the sample size from \(t_c\) (as done above) only when \(n_1 = n_2\).
+When relaxing the assumption of equal variances (i.e., using the Welch's \(t\)-test) \(s\) is calculated as the square root of the average between the two samples variances, \(SE_b\) is calculated as \(\sqrt{\frac{s^2_{x_1}}{n_1} + \frac{s^2_{x_2}}{n_2}}\).
+The degrees of freedom for \(t_c\) are calculated in a more complex way using the Welch-Satterthwaite equation (Satterthwaite, 1946; Welch, 1947).
+For the Welch's \(t\)-test we can calculate the critical effect size removing the sample size from \(t_c\) (as done above) only when \(n_1 = n_2\).
 
 \hypertarget{paired-sample-t-test}{%
-\subsubsection{Paired sample t-test}\label{paired-sample-t-test}}
+\subsubsection{\texorpdfstring{Paired sample \(t\)-test}{Paired sample t-test}}\label{paired-sample-t-test}}
 
 For the paired sample case the situation is less straightforward.
-The reason is that the \(t\) statistic is computed using \(b\) being the average of the paired differences between the two (paired) samples and \(s\) being the standard deviation of the differences.
-Using these values, we can just compute the critical effects size as for the one-sample case, in fact a paired t-test is a one-sample test on the vector of paired differences.
+The reason is that the \(t\) statistic is computed with \(b\) being the average of the paired differences between the two (paired) samples and \(s\) being the standard deviation of the differences.
+Using these values, we can just compute the critical effects size as for the one-sample case, as a paired \(t\)-test is a one-sample test on the vector of paired differences.
 Test statistics and effect size based on the standard deviation of differences (the so called \(d_z\)) are used for hypothesis testing and power analysis (Lakens, 2013).
@@
 For this reason, there is a corrected version of the effect size called Hedges's \(g\) Viechtbauer (2007).
-The Hedges's \(g\) can be calculated for all the t-test scenarios.
+The Hedges's \(g\) can be calculated for all the \(t\)-test scenarios.
 The Hegdes's correction is implemented in Equation \eqref{eq:eq-g-correction} where \(\Gamma\) is the gamma function and \(m\) the degrees of freedom.
@@
 
-The correction is applied as \(g = c(m) \times d\). The hypothesis testing is not taking into account the Hedges's correction thus the critical \(g\) will be different from the critical \(d\).
+The correction is applied as \(g = c(m) \times d\). The hypothesis testing does not take the Hedges's correction into account, thus the critical \(g\) will be different from the critical \(d\).
 
@@
 Hypothesis testing for the Pearson's correlation coefficient is usually done using the t statistics or using a z statistics.
-The general approach presented for the t-test is still valid.
+The general approach presented for the \(t\)-test is still valid.
 The only difference is that the correlation is already a standardized effect size thus there is no need of the standardization term \(s\).
-Equation \eqref{eq:eq-gen-t} can be used substituing \(b\) with \(r\) (the sample correlation coefficient) as shown in Equation \eqref{eq:eq-rc}.
+Equation \eqref{eq:eq-gen-t} can be used substituting \(b\) with \(r\) (the sample correlation coefficient) as shown in Equation \eqref{eq:eq-rc}.
 The standard error is calculated as \(SE_r = \sqrt{\frac{1 - r^2}{n - 2}}\) (Bowley, 1928).
@@
 Hypothesis testing on regression coefficients (e.g., using the \texttt{lm} function in R) is performed using the so-called Wald test.
-The test can be considered a one-sample t-test where the \(t\) value is calculated dividing the regression parameter \(\beta_0, \dots, \beta_j\) by its standard error \(SE_{\beta_j}\).
-The critical \(t\) value is calculated from a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept.
+The test can be considered a one-sample \(t\)-test where the \(t\)-value is calculated dividing the regression parameter \(\beta_0, \dots, \beta_j\) by its standard error \(SE_{\beta_j}\).
+The critical \(t\)-value is calculated from a \(t\) distribution using \(n - p - 1\) degrees of freedom where \(n\) is the number of observations and \(p\) is the number of coefficients beyond the intercept.
 We can substitute \(b\) with \(\beta_j\) and \(SE_{b}\) with \(SE_{\beta_j}\) in Equation \eqref{eq:eq-gen-crit} to find the critical unstandardized regression coefficient (\(\beta_{c_j} = t_c \times SE_{\beta_j}\)).
-There is no an unique way to calculate the standardized version of the critical regression coefficient.
+There is no unique way to calculate the standardized version of the critical regression coefficient.
 The main reason is that in regression modelling we can include categorical and numerical independent variables with main effects and interactions.
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package ``criticalESvalue'' in R.
+In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package \texttt{criticalESvalue} in R.
 The complete package can be accessed at: \url{https://github.com/psicostat/criticalESvalue}.
-Here, we demonstrate its application through two examples: one example of a t-test on real data and a computation of the critical effect size values for a correlation from sample size.
+Here, we demonstrate its application through two examples: one example of a \(t\)-test on real data and a computation of the critical effect size values for a correlation from sample size.
 It is important to note that, in general, depending on the researcher's hypotheses, the package allows for the calculation of either two critical effect size values for two-tailed tests or a single critical effect size value for one-tailed tests, for which the direction must be specified.
-In the Supplementary online materials additional examples can be found for correlation, t-test, paired t-test, linear models and regression coefficients.
+In the Supplementary online materials additional examples can be found for correlation, \(t\)-test, paired \(t\)-test, linear models and regression coefficients.
 
-First the package should be downloaded and opened with the library function:
+First the package should be downloaded and opened with the \texttt{library} function:
 
@@
 
-For our examples on real data we used from the package `psych' the dataset ``holzinger.swineford'' which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects.
-Once the package is retrieved with `library', the dataset can be opened using `data(``name of the dataset'')'.
+For our examples on real data we used from the package \texttt{psych} the dataset ``holzinger.swineford'' which has a series of demographics and scores of different subtests measuring intelligence on 301 subjects.
+Once the package is retrieved with \texttt{library}, the dataset can be opened using \texttt{data}(``name of the dataset'').
 For simplicity we decided to rename it with a shorter name.
@@
 
-We want to know the critical value for a t-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the `t.test' function:
+We want to know the critical value for a \(t\)-test comparing boys and girls on a cognitive variable of visual perception. In this case, it can be easily done using the \texttt{t.test} function:
 
@@
 
-The direction of the hypothesis and the test to apply, either \(t\)-test or \(z\)-test, should be specified.
-The output will return the critical correlation value(s), the degrees of freedom and the type of test used.
+The direction of the hypothesis and the test to apply, either the \(t\)-test or \(z\)-test, should be specified.
+The output will return the critical correlation value(s), the degrees of freedom, and the type of test used.
 

### 2025-01-07 R/criticalvalue-funs.R
clean

@@ -524,6 +524,7 @@ critical_cor <- function(r = NULL, n,
 #' n <- 170
 #' p = 3
 #' seb <- c(25,.25,.10)
+#' # seb is an hypothetical vector of standard errors from summary() of an lm() model 
 #' critical_coef(seb = seb,n = n, p = p)
 #' 
 # TODO controllare esempio

### 2025-01-07 paper/paper.Rmd

@@
   - name          : "Ambra Perugini"
+    corresponding: "yes"
+    address       : "Via Venezia 8, 35131, Padova, Italy"
+    email         : "ambra.perugini@phd.unipd.it"
     affiliation   : "1"
@@
       - "Writing - Review & Editing"
+      - "Methodology"
       - "Software"
@@
     affiliation   : "1"
-    corresponding: "yes"
-    address       : "Via Venezia 8, 35131, Padova, Italy"
-    email         : "gianmarco.altoe@unipd.it"
     role:

### 2025-01-07 paper/paper.tex

@@
 The authors made the following contributions.
-Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Software, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Conceptualization, Writing - Review \& Editing, Supervision; Livio Finos: Conceptualization, Writing - Review \& Editing, Supervision; Psicostat: Conceptualization, Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
+Ambra Perugini: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Filippo Gambarota: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology, Software; Enrico Toffalini: Writing -- Original Draft Preparation, Writing - Review \& Editing, Methodology, Software, Supervision; Daniël Lakens: Conceptualization, Writing - Review \& Editing, Supervision; Massimiliano Pastore: Conceptualization, Writing - Review \& Editing, Supervision; Livio Finos: Conceptualization, Writing - Review \& Editing, Supervision; Psicostat: Conceptualization, Writing - Review \& Editing, Supervision; Gianmarco Altoè: Conceptualization, Writing - Review \& Editing, Supervision.
 
-Correspondence concerning this article should be addressed to Gianmarco Altoè, Via Venezia 8, 35131, Padova, Italy. E-mail: \href{mailto:gianmarco.altoe@unipd.it}{\nolinkurl{gianmarco.altoe@unipd.it}}
+Correspondence concerning this article should be addressed to Ambra Perugini, Via Venezia 8, 35131, Padova, Italy.
+E-mail: \href{mailto:ambra.perugini@phd.unipd.it}{\nolinkurl{ambra.perugini@phd.unipd.it}}
 

### 2025-01-13 paper/paper.Rmd
^^

--- 2025-01-02 paper/paper.Rmd
@@
 ---
-title             : "The benefits of reporting critical effect size values"
+title             : "The benefits of reporting critical effect size values^[The manuscript as been posted as a pre-print and has not yet undergone peer-review.]"
 shorttitle        : "Reporting critical effect size values"

### 2025-01-14 README.Rmd
adding badges

@@
 
+<!-- badges: start -->
+[![R-CMD-check](https://github.com/psicostat/criticalESvalue/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/psicostat/criticalESvalue/actions/workflows/R-CMD-check.yaml)
+[![test-coverage](https://github.com/psicostat/criticalESvalue/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/psicostat/criticalESvalue/actions/workflows/test-coverage.yaml)
+<!-- badges: end -->
+
 # criticalESvalue
 
-<!-- badges: start -->
-<!-- badges: end -->
-
-The goal of criticalvalue is to ...
+The goal of `criticalESvalue` is to ...
 <!-- TODO aggiungi qualche riga qui -->
@@
 
-You can install the development version of criticalESvalue like so:
+You can install the development version of `criticalESvalue` like so:
 

### 2025-01-14 R/criticalvalue-funs.R
fixing typo

@@ -327,7 +327,7 @@ critical_t1s <- function(m = NULL, s = NULL, t = NULL,
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #' @param conf.level the confidence level to set the confidence interval, default is set to 0.95.
 #'
-#' @return the output returns a `d` which is the Cohen's d, the critical d (`dc`) which is the minimum value for which to get a significant result with a given sample, the `bc` is the numerator of the formula from which the d is calculated, `se` which is the standard error, `df` are the degrees of freedom, then it also gives the `g` and `gc` which are respectively `d` and `dc` with Hedfer's Correction for small samples.
+#' @return the output returns a `d` which is the Cohen's d, the critical d (`dc`) which is the minimum value for which to get a significant result with a given sample, the `bc` is the numerator of the formula from which the d is calculated, `se` which is the standard error, `df` are the degrees of freedom, then it also gives the `g` and `gc` which are respectively `d` and `dc` with Hedges's correction for small samples.
 #' @export
 #'
 #' @examples
@@ -344,7 +344,6 @@ critical_t1s <- function(m = NULL, s = NULL, t = NULL,
 #' t <- (m1 - m2) / se
 #' critical_t2s(t = t, n1 = n1, n2 = n2, se = se) # se only required for calculating bc
 
-# TODO controllare formula dell'esemptio from t test
 critical_t2s <- function(m1 = NULL, m2 = NULL, t = NULL,
                          sd1 = NULL, sd2 = NULL,
                          n1, n2, se = NULL,
@@ -399,7 +398,7 @@ critical_t2s <- function(m1 = NULL, m2 = NULL, t = NULL,
 #' @param hypothesis a character string indicating the alternative hypothesis ("less", "greater" or "two.tailed").
 #' @param conf.level the confidence level to set the confidence interval, default is set to 0.95.
 #'
[…]

### 2025-01-14 R/criticalvalue-methods.R
better formula syntax with data error

@@ -28,7 +28,7 @@ critical.htest <- function(x, ...){
     D <- insight::get_data(x)
     
     if(is.null(D)){
-      stop("insight::get_data(x) returning NULL. Are you using the formula syntax (y ~ x)? This syntax is not supported yet")
+      stop("insight::get_data(x) returning NULL. Are you using the formula syntax (y ~ x) with the data = argument? This syntax is not supported yet. See vignette('formula-syntax', package = 'criticalESvalue'). Try to call the function without the 'data = ' argument")
     }
     
     conf.level <- attributes(x$conf.int)$conf.level

### 2025-01-14 README.Rmd
adding package summary

@@
 
-The goal of `criticalESvalue` is to ...
-<!-- TODO aggiungi qualche riga qui -->
+The `criticalESvalue` package provide a set of functions to calculate the critical effects size value for the common statistical model.
+Currently, the package support the `htest` class (i.e., `t.test` and `cor.test`), the `lm` class and the `rma` class from the `metafor` package.
+
+The package contains a set of functions to work with summary data and the `critical()` method that takes objects of class `htest`, `lm` or `rma` and provide enhanced printing and summary methods with information about critical effects size.
 

### 2025-02-24 paper/paper.Rmd
adding equations table

@@
 documentclass     : "apa7"
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : no
@@
 
+\bfilippo
+Table \@ref(tab:eqsummary) summarises the equations presented above.
+
+```{r eqsummary, echo=FALSE, tab.cap="Summary of the critical effect size equations."}
+library(kableExtra)
+
+eqs <- data.frame(
+  stringsAsFactors = FALSE,
+             Model = c("One-sample t-test",
+                       "Paired t-test","Paired t-test","Between-sample t-test",
+                       "Correlation","Correlation","Meta-analysis/Regression"),
+         Numerator = c("$b_c = t_c \\times \\mbox{SE}_b$","$b_c = t_c \\times \\mbox{SE}_b$",
+                       "$b_{z_c} = t_c \\times \\mbox{SE}_b$",
+                       "$b_c = t_c \\times \\mbox{SE}_b$",
+                       "$r_c = \\frac{t_c}{\\sqrt{n - 2 + t^2_c}}$","$r_{z_c} = \\frac{z_c}{\\sqrt{n - 3}}$",
+                       "$\\beta_c = t_c \\times \\mbox{SE}_{\\beta}$"),
+       Denominator = c("$s$", NA, "$s_D$", "$s_p$", NA, NA, NA),
+          Equation = c("$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s}$",
+                       "$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s_p}$",
+                       "$d_{z_c} = \\frac{t_c \\times \\mbox{SE}_b}{s_D}$",
+                       "$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s_p}$",
+                       "$r_c = \\frac{t_c}{\\sqrt{n - 2 + t^2_c}}$","$r_{z_c} = \\frac{z_c}{\\sqrt{n - 3}}$",
+                       "$\\beta_c = t_c \\times \\mbox{SE}_{\\beta}$")
+)
+
+eqs |> 
+  dplyr::select(Model, Equation) |> 
+  kable(escape = FALSE,
+        booktabs = TRUE) |> 
+  collapse_rows(1) |> 
+  kable_styling(latex_options = "HOLD_position") |> 
+  row_spec(0, bold = TRUE)
+```
+
+\efilippo
+
 # Examples in R

### 2025-02-24 paper/paper.tex
^^
adding equations table

@@
 \documentclass[
-  man]{apa7}
+  man,floatsintext]{apa7}
 \usepackage{amsmath,amssymb}
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4761}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{csquotes}
@@
 
-\title{The benefits of reporting critical effect size values}
+\title{The benefits of reporting critical effect size values\footnote{The manuscript as been posted as a pre-print and has not yet undergone peer-review.}}
 \author{Ambra Perugini\textsuperscript{1}, Filippo Gambarota\textsuperscript{1}, Enrico Toffalini\textsuperscript{2}, Daniël Lakens\textsuperscript{3}, Massimiliano Pastore\textsuperscript{1}, Livio Finos\textsuperscript{4}, Psicostat\textsuperscript{1}, \& Gianmarco Altoè\textsuperscript{1}}
@@
 
-\hypertarget{small-sample-sizes-and-uncertain-sample-size-setermination}{%
-\subsection{Small Sample Sizes and Uncertain Sample Size Setermination}\label{small-sample-sizes-and-uncertain-sample-size-setermination}}
+\subsection{Small Sample Sizes and Uncertain Sample Size Setermination}\label{small-sample-sizes-and-uncertain-sample-size-setermination}
 
@@
 
-\hypertarget{large-sample-sizes-and-meta-analysis}{%
-\subsection{Large Sample Sizes and Meta-analysis}\label{large-sample-sizes-and-meta-analysis}}
+\subsection{Large Sample Sizes and Meta-analysis}\label{large-sample-sizes-and-meta-analysis}
 
@@
 
-\hypertarget{how-to-compute-critical-effect-size-values}{%
-\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}}
+\section{How to Compute Critical Effect Size Values}\label{how-to-compute-critical-effect-size-values}
 
@@
 
-\hypertarget{t-test}{%
-\subsection{\texorpdfstring{\(t\)-test}{t-test}}\label{t-test}}
+\subsection{\texorpdfstring{\(t\)-test}{t-test}}\label{t-test}
 
@@
 
-\hypertarget{one-sample-t-test}{%
-\subsubsection{\texorpdfstring{One sample \(t\)-test}{One sample t-test}}\label{one-sample-t-test}}
+\subsubsection{\texorpdfstring{One sample \(t\)-test}{One sample t-test}}\label{one-sample-t-test}
 
@@
 
-\hypertarget{two-sample-t-test}{%
-\subsubsection{\texorpdfstring{Two-sample \(t\)-test}{Two-sample t-test}}\label{two-sample-t-test}}
+\subsubsection{\texorpdfstring{Two-sample \(t\)-test}{Two-sample t-test}}\label{two-sample-t-test}
 
@@
 
-\hypertarget{paired-sample-t-test}{%
-\subsubsection{\texorpdfstring{Paired sample \(t\)-test}{Paired sample t-test}}\label{paired-sample-t-test}}
+\subsubsection{\texorpdfstring{Paired sample \(t\)-test}{Paired sample t-test}}\label{paired-sample-t-test}
 
@@
 
-\hypertarget{hedgess-correction}{%
-\subsubsection{Hedges's correction}\label{hedgess-correction}}
+\subsubsection{Hedges's correction}\label{hedgess-correction}
 
@@
 
-\hypertarget{meta-analysis}{%
-\subsection{Meta-analysis}\label{meta-analysis}}
+\subsection{Meta-analysis}\label{meta-analysis}
 
@@
 
-\hypertarget{other-models}{%
-\subsection{Other models}\label{other-models}}
+\subsection{Other models}\label{other-models}
 
@@
 
-\hypertarget{examples-in-r}{%
-\section{Examples in R}\label{examples-in-r}}
+\color{blue}
+
+Table \ref{tab:eqsummary} summarises the equations presented above.
+
+\footnotesize
+
+\begin{table}[H]
+
+\caption{\label{tab:eqsummary}Summary of the critical effect size equations.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+\textbf{Model} & \textbf{Equation}\\
+\midrule
+One-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s}$\\
+\cmidrule{1-2}
+ & $d_c = \frac{t_c \times \mbox{SE}_b}{s_p}$\\
+\cmidrule{2-2}
+\multirow{-2}{*}{\raggedright\arraybackslash Paired t-test} & $d_{z_c} = \frac{t_c \times \mbox{SE}_b}{s_D}$\\
+\cmidrule{1-2}
+Between-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s_p}$\\
+\cmidrule{1-2}
+ & $r_c = \frac{t_c}{\sqrt{n - 2 + t^2_c}}$\\
+\cmidrule{2-2}
+\multirow{-2}{*}{\raggedright\arraybackslash Correlation} & $r_{z_c} = \frac{z_c}{\sqrt{n - 3}}$\\
+\cmidrule{1-2}
+Meta-analysis/Regression & $\beta_c = t_c \times \mbox{SE}_{\beta}$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\normalsize
+
+\color{black}
+
+\section{Examples in R}\label{examples-in-r}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
+\bibitem[\citeproctext]{ref-Borenstein2009-mo}
 Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bowley1928-yt}{}}%
+\bibitem[\citeproctext]{ref-Bowley1928-yt}
 Bowley, A. L. (1928). The standard deviation of the correlation coefficient. \emph{Journal of the American Statistical Association}, \emph{23}, 31. \url{https://doi.org/10.2307/2277400}
 
-\leavevmode\vadjust pre{\hypertarget{ref-byers2020building}{}}%
+\bibitem[\citeproctext]{ref-byers2020building}
 Byers-Heinlein, K., Bergmann, C., Davies, C., Frank, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chow1988significance}{}}%
+\bibitem[\citeproctext]{ref-chow1988significance}
 Chow, S. L. (1988). Significance test or effect size? \emph{Psychological Bulletin}, \emph{103}(1), 105.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
+\bibitem[\citeproctext]{ref-cohen1994}
 Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cortina1997logic}{}}%
+\bibitem[\citeproctext]{ref-cortina1997logic}
 Cortina, J. M., \& Dunlap, W. P. (1997). On the logic and purpose of significance testing. \emph{Psychological Methods}, \emph{2}(2), 161.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fisher1915-ks}{}}%
+\bibitem[\citeproctext]{ref-Fisher1915-ks}
 Fisher, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fisher1921-qz}{}}%
+\bibitem[\citeproctext]{ref-Fisher1921-qz}
 Fisher, R. A. (1921). On the "probable error" of a coefficient of correlation deduced from a small sample. \emph{Metron}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-funder2019evaluating}{}}%
+\bibitem[\citeproctext]{ref-funder2019evaluating}
 Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: Sense and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008-zp}{}}%
+\bibitem[\citeproctext]{ref-Gelman2008-zp}
 Gelman, A. (2008). Scaling regression inputs by dividing by two standard deviations. \emph{Statistics in Medicine}, \emph{27}, 2865--2873. \url{https://doi.org/10.1002/sim.3107}
 
-\leavevmode\vadjust pre{\hypertarget{ref-gigerenzer2004null}{}}%
+\bibitem[\citeproctext]{ref-gigerenzer2004null}
 Gigerenzer, G., Krauss, S., \& Vitouch, O. (2004). The null ritual. \emph{The Sage Handbook of Quantitative Methodology for the Social Sciences}, 391--408.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gross2023-fl}{}}%
+\bibitem[\citeproctext]{ref-Gross2023-fl}
 Groß, J., \& Möller, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hagen1997praise}{}}%
+\bibitem[\citeproctext]{ref-hagen1997praise}
 Hagen, R. L. (1997). \emph{In praise of the null hypothesis statistical test.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-haig2017tests}{}}%
+\bibitem[\citeproctext]{ref-haig2017tests}
 Haig, B. D. (2017). Tests of statistical significance made sound. \emph{Educational and Psychological Measurement}, \emph{77}(3), 489--506.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hedges1981-za}{}}%
+\bibitem[\citeproctext]{ref-Hedges1981-za}
 Hedges, L.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-jarke2022roadmap}{}}%
+\bibitem[\citeproctext]{ref-jarke2022roadmap}
 Jarke, H., Anand-Vembar, S., Alzahawi, S., Andersen, T.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
+\bibitem[\citeproctext]{ref-krueger2001null}
 Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
+\bibitem[\citeproctext]{ref-lakens2013calculating}
 Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2021practical}{}}%
+\bibitem[\citeproctext]{ref-lakens2021practical}
 Lakens, D. (2021). The practical alternative to the p value is the correctly used p value. \emph{Perspectives on Psychological Science}, \emph{16}(3), 639--648.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
+\bibitem[\citeproctext]{ref-lakens2022sample}
 Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.
 
-\leavevmode\vadjust pre{\hypertarget{ref-miller2017hypothesis}{}}%
+\bibitem[\citeproctext]{ref-miller2017hypothesis}
 Miller, J. (2017). Hypothesis testing in the real world. \emph{Educational and Psychological Measurement}, \emph{77}(4), 663--672.
 
-\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
+\bibitem[\citeproctext]{ref-morris2002combining}
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
 
-\leavevmode\vadjust pre{\hypertarget{ref-moshontz2018psychological}{}}%
+\bibitem[\citeproctext]{ref-moshontz2018psychological}
 Moshontz, H., Campbell, L., Ebersole, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nickerson2000null}{}}%
+\bibitem[\citeproctext]{ref-nickerson2000null}
 Nickerson, R. S. (2000). Null hypothesis significance testing: A review of an old and continuing controversy. \emph{Psychological Methods}, \emph{5}(2), 241.
 
-\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
+\bibitem[\citeproctext]{ref-orben_crud_2020}
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Satterthwaite1946-aq}{}}%
+\bibitem[\citeproctext]{ref-Satterthwaite1946-aq}
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
 
-\leavevmode\vadjust pre{\hypertarget{ref-sirois2023pupil}{}}%
+\bibitem[\citeproctext]{ref-sirois2023pupil}
 Sirois, S., Brisson, J., Blaser, E., Calignano, G., Donenfeld, J., Hepach, R., Hochmann, J.-R., Kaldy, Z., Liszkowski, U., Mayer, M., et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-szucs2017empirical}{}}%
+\bibitem[\citeproctext]{ref-szucs2017empirical}
 Szucs, D., \& Ioannidis, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-transue2019apa}{}}%
+\bibitem[\citeproctext]{ref-transue2019apa}
 Transue, B. (2019). \emph{APA style 7th edition}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2007-tq}{}}%
+\bibitem[\citeproctext]{ref-Viechtbauer2007-tq}
 Viechtbauer, W.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-weiss2014spatial}{}}%
+\bibitem[\citeproctext]{ref-weiss2014spatial}
 Weiss, A. H., Biron, T., Lieder, I., Granot, R. Y., \& Ahissar, M. (2014). Spatial vision is superior in musicians when memory plays a role. \emph{Journal of Vision}, \emph{14}(9), 18--18.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Welch1947-wh}{}}%
+\bibitem[\citeproctext]{ref-Welch1947-wh}
 Welch, B. L. (1947). The generalization of {``student's''} problem when several different population variances are involved. \emph{Biometrika}, \emph{34}, 28. \url{https://doi.org/10.2307/2332510}
 
-\leavevmode\vadjust pre{\hypertarget{ref-wilson2020science}{}}%
+\bibitem[\citeproctext]{ref-wilson2020science}
 Wilson, B. M., Harris, C. R., \& Wixted, J. T. (2020). Science is not a signal detection problem. \emph{Proceedings of the National Academy of Sciences}, \emph{117}(11), 5559--5567.

### 2025-02-26 paper/paper.Rmd
updates

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4761"
+wordcount         : "4856"
 
@@
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
-These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest.
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest [@cumming2012statistical; @transue2019apa].
 Researchers often publish studies with low power for medium to small effect sizes [@szucs2017empirical], with the main justification for the sample size being resource limitations [@lakens2022sample].
@@
 Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
-Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant.
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant [@anvari2023not].
+The risk is to heuristically accept all effects as potentially important, for this reason "researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it" [@anvari2023not].
 
@@
 
-## Small Sample Sizes and Uncertain Sample Size Setermination
+## Small Sample Sizes and Uncertain Sample Size Determination
 
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis [@altoe2020enhancing], but this would require both knowledge of the plausible effect size and/or the smallest effect size of interest (SESOI), which are not always easily to determine.
+In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis [@altoe2020enhancing], but this would require both knowledge of the plausible effect size and/or the smallest effect size of interest (SESOI), which is the minimum effect that could be considered meaningful based on practical relevance, theoretical importance and/or specific research interests [@riesthuis2024simulation; @mesquida2024effect], which are not always easy to determine.
 The use of critical effect size values can be used in a simple an efficient way to evaluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
@@
 Beyond critical effect size values, researchers should reason before conducting a study what conclusions can realistically be drawn from their experimental design.
-The more complex the study design, the higher the critical effect size value.
-Researchers should assess whether their available sample size allows them to address a simpler yet still relevant question.
+As study design complexity increases, so does the critical effect size, due to greater variability, interaction effects, and the need for statistical corrections, all of which reduce statistical power and require a larger sample size to detect small effects reliably.
 If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies should be explored [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building].
@@
 Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
-This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to significant but very small and meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results.
+Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results [@anvari2023not].
 
@@
 The effect size calculated as in the previous step is known to be inflated especially for small samples.
-For this reason, there is a corrected version of the effect size called Hedges's $g$ [@Hedges1981-za, @Viechtbauer2007-tq].
+For this reason, there is a corrected version of the effect size called Hedges's $g$ by [@Hedges1981-za, @Viechtbauer2007-tq].
 The Hedges's $g$ can be calculated for all the $t$-test scenarios.
@@
 From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances.
-Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests [@Borenstein2009-mo].
+Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests [@Viechtbauer2015-tn; @Rubio-Aparicio2020-gd].
 For this reason, we can apply the same equations used for the linear regression.
@@
 
-\bfilippo
 Table \@ref(tab:eqsummary) summarises the equations presented above.
@@
 
-\efilippo
-
 # Examples in R
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package `criticalESvalue` in R.
-The complete package can be accessed at: https://github.com/psicostat/criticalESvalue.
+In this section, we introduce the aforementioned user-friendly implementation of the mathematical computations as functions of the package `criticalESvalue` in R.
 Here, we demonstrate its application through two examples: one example of a $t$-test on real data and a computation of the critical effect size values for a correlation from sample size.

### 2025-02-26 paper/paper.tex
updates

--- 2025-01-07 paper/paper.tex
@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4761}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4856}
 \usepackage{csquotes}
@@
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
-These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest.
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest (Cumming et al., 2012; Transue, 2019).
 Researchers often publish studies with low power for medium to small effect sizes (Szucs \& Ioannidis, 2017), with the main justification for the sample size being resource limitations (Lakens, 2022).
@@
 Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
-Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant.
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant (Anvari et al., 2023).
+The risk is to heuristically accept all effects as potentially important, for this reason ``researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it'' (Anvari et al., 2023).
 
@@
 
-\hypertarget{small-sample-sizes-and-uncertain-sample-size-setermination}{%
-\subsection{Small Sample Sizes and Uncertain Sample Size Setermination}\label{small-sample-sizes-and-uncertain-sample-size-setermination}}
+\hypertarget{small-sample-sizes-and-uncertain-sample-size-determination}{%
+\subsection{Small Sample Sizes and Uncertain Sample Size Determination}\label{small-sample-sizes-and-uncertain-sample-size-determination}}
 
@@
 An a-priori power analysis would typically lead to a similar conclusion, but reporting critical effect size values will focus the attention more strongly on which effect sizes are reasonable to expect.
-In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis (Altoè et al., 2020), but this would require both knowledge of the plausible effect size and/or the smallest effect size of interest (SESOI), which are not always easily to determine.
+In case of already conducted studies with small sample sizes it could be argued that it would be more informative to use retrospective design analysis (Altoè et al., 2020), but this would require both knowledge of the plausible effect size and/or the smallest effect size of interest (SESOI), which is the minimum effect that could be considered meaningful based on practical relevance, theoretical importance and/or specific research interests (Mesquida \& Lakens, 2024; Riesthuis, 2024), which are not always easy to determine.
 The use of critical effect size values can be used in a simple an efficient way to evaluate which findings could have not be found significant due to sample size limitations and based on the complexity of the test.
@@
 Beyond critical effect size values, researchers should reason before conducting a study what conclusions can realistically be drawn from their experimental design.
-The more complex the study design, the higher the critical effect size value.
-Researchers should assess whether their available sample size allows them to address a simpler yet still relevant question.
+As study design complexity increases, so does the critical effect size, due to greater variability, interaction effects, and the need for statistical corrections, all of which reduce statistical power and require a larger sample size to detect small effects reliably.
 If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies should be explored (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023).
@@
 Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
-This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to small but theoretically meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to significant but very small and meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
 
@@
 While the focus of meta-analysis is generally on estimating effect sizes with uncertainty, statistical significance is routinely reported and interpreted.
-Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results.
+Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results (Anvari et al., 2023).
 
@@
 The effect size calculated as in the previous step is known to be inflated especially for small samples.
-For this reason, there is a corrected version of the effect size called Hedges's \(g\) Viechtbauer (2007).
+For this reason, there is a corrected version of the effect size called Hedges's \(g\) by Viechtbauer (2007).
 The Hedges's \(g\) can be calculated for all the \(t\)-test scenarios.
@@
 From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances.
-Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests (Borenstein et al., 2009).
+Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests (Rubio-Aparicio et al., 2020; Viechtbauer et al., 2015).
 For this reason, we can apply the same equations used for the linear regression.
@@
 
+Table \ref{tab:eqsummary} summarises the equations presented above.
+
+\footnotesize
+
+\begin{table}[H]
+\centering
+\caption{\label{tab:eqsummary}Summary of the critical effect size equations.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+\textbf{Model} & \textbf{Equation}\\
+\midrule
+One-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s}$\\
+\cmidrule{1-2}
+ & $d_c = \frac{t_c \times \mbox{SE}_b}{s_p}$\\
+\cmidrule{2-2}
+\multirow{-2}{*}{\raggedright\arraybackslash Paired t-test} & $d_{z_c} = \frac{t_c \times \mbox{SE}_b}{s_D}$\\
+\cmidrule{1-2}
+Between-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s_p}$\\
+\cmidrule{1-2}
+ & $r_c = \frac{t_c}{\sqrt{n - 2 + t^2_c}}$\\
+\cmidrule{2-2}
+\multirow{-2}{*}{\raggedright\arraybackslash Correlation} & $r_{z_c} = \frac{z_c}{\sqrt{n - 3}}$\\
+\cmidrule{1-2}
+Meta-analysis/Regression & $\beta_c = t_c \times \mbox{SE}_{\beta}$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\normalsize
+
 \hypertarget{examples-in-r}{%
@@
 
-In this section, we introduce a user-friendly implementation of the aforementioned mathematical computations as functions of the package \texttt{criticalESvalue} in R.
-The complete package can be accessed at: \url{https://github.com/psicostat/criticalESvalue}.
+In this section, we introduce the aforementioned user-friendly implementation of the mathematical computations as functions of the package \texttt{criticalESvalue} in R.
 Here, we demonstrate its application through two examples: one example of a \(t\)-test on real data and a computation of the critical effect size values for a correlation from sample size.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Borenstein2009-mo}{}}%
-Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2009). \emph{Introduction to {Meta-Analysis}}. \url{https://doi.org/10.1002/9780470743386}
+\leavevmode\vadjust pre{\hypertarget{ref-anvari2023not}{}}%
+Anvari, F., Kievit, R., Lakens, D., Pennington, C.
+R., Przybylski, A.
+K., Tiokhin, L., Wiernik, B.
+M., \& Orben, A.
+(2023).
+Not all effects are indispensable: Psychological science requires verifiable lines of reasoning for whether an effect matters.
+\emph{Perspectives on Psychological Science}, \emph{18}(2), 503--507.
 
@@
 Cortina, J. M., \& Dunlap, W. P. (1997). On the logic and purpose of significance testing. \emph{Psychological Methods}, \emph{2}(2), 161.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cumming2012statistical}{}}%
+Cumming, G., Fidler, F., Kalinowski, P., \& Lai, J.
+(2012).
+The statistical recommendations of the american psychological association publication manual: Effect sizes, confidence intervals, and meta-analysis.
+\emph{Australian Journal of Psychology}, \emph{64}(3), 138--146.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mesquida2024effect}{}}%
+Mesquida, C., \& Lakens, D.
+(2024).
+Is the effect large enough to matter?
+Why exercise physiologists should interpret effect sizes meaningfully: A reply to williams et al.(2023).
+\emph{Journal of Physiology}, \emph{602}(1).
+
 \leavevmode\vadjust pre{\hypertarget{ref-miller2017hypothesis}{}}%
@@
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
+
+\leavevmode\vadjust pre{\hypertarget{ref-riesthuis2024simulation}{}}%
+Riesthuis, P.
+(2024).
+Simulation-based power analyses for the smallest effect size of interest: A confidence-interval approach for minimum-effect and equivalence testing.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{7}(2), 25152459241240722.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rubio-Aparicio2020-gd}{}}%
+Rubio-Aparicio, M., López-López, J.
+A., Viechtbauer, W., Marín-Martínez, F., Botella, J., \& Sánchez-Meca, J.
+(2020).
+Testing categorical moderators in mixed-effects meta-analysis in the presence of heteroscedasticity.
+\emph{Journal of Experimental Education}, \emph{88}, 288--310.
+\url{https://doi.org/10.1080/00220973.2018.1561404}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2015-tn}{}}%
+Viechtbauer, W., López-López, J.
+A., Sánchez-Meca, J., \& Marín-Martínez, F.
+(2015).
+A comparison of procedures to test for moderators in mixed-effects meta-regression models.
+\emph{Psychological Methods}, \emph{20}, 360--374.
+\url{https://doi.org/10.1037/met0000023}
+
 \leavevmode\vadjust pre{\hypertarget{ref-weiss2014spatial}{}}%

### 2025-03-02 paper/paper.Rmd
corrections

@@
 Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
-Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant [@anvari2023not].
-The risk is to heuristically accept all effects as potentially important, for this reason "researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it" [@anvari2023not].
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant [@kirk1996practical; @anvari2023not].
+The risk is that researchers will heuristically or opportunistically accept all effects as potentially important.
+For this reason "researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it" [@anvari2023not].
 
@@
 Beyond critical effect size values, researchers should reason before conducting a study what conclusions can realistically be drawn from their experimental design.
-As study design complexity increases, so does the critical effect size, due to greater variability, interaction effects, and the need for statistical corrections, all of which reduce statistical power and require a larger sample size to detect small effects reliably.
+For example, as interaction effects are examined in addition to main effects, power may decrease because interactions are generally associated with smaller effect sizes, may require corrections for multiple tests, and interaction coefficients present larger standard errors than main effects coefficients.
+Such a reduction in statistical power may require a larger sample size to detect the effects of interest reliably.
 If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies should be explored [@sirois2023pupil; @moshontz2018psychological; @jarke2022roadmap; @byers2020building].
@@
 Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results [@anvari2023not].
+For this reason, the ongoing call to specify which effects constitute an ‘important difference’ [@bakan1967method; @boring1919mathematical; @hodges1954matching; @kirk1996practical; @lakens2021practical; @nunnally1960place; @serlin1985rationality] has become especially urgent.
 

### 2025-03-02 paper/paper.tex
corrections

@@
 Certain types of research questions can be studied by relying on online data collection, which has made it cheaper and more feasible to collect very large samples.
-Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant (Anvari et al., 2023).
-The risk is to heuristically accept all effects as potentially important, for this reason ``researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it'' (Anvari et al., 2023).
+Studies with very large samples are well-powered to detect even small effect sizes, but they also require researchers to carefully consider the possibility that an effect might be statistically significant, but practically insignificant (Anvari et al., 2023; Kirk, 1996).
+The risk is that researchers will heuristically or opportunistically accept all effects as potentially important.
+For this reason ``researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it'' (Anvari et al., 2023).
 
@@
 Beyond critical effect size values, researchers should reason before conducting a study what conclusions can realistically be drawn from their experimental design.
-As study design complexity increases, so does the critical effect size, due to greater variability, interaction effects, and the need for statistical corrections, all of which reduce statistical power and require a larger sample size to detect small effects reliably.
+For example, as interaction effects are examined in addition to main effects, power may decrease because interactions are generally associated with smaller effect sizes, may require corrections for multiple tests, and interaction coefficients present larger standard errors than main effects coefficients.
+Such a reduction in statistical power may require a larger sample size to detect the effects of interest reliably.
 If this is not feasible, and sample size constraints make it impossible to detect significance for a plausible effect because it is smaller than the critical effect size value, alternative data collection strategies, such as multisite studies should be explored (Byers-Heinlein et al., 2020; Jarke et al., 2022; Moshontz et al., 2018; Sirois et al., 2023).
@@
 Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results (Anvari et al., 2023).
+For this reason, the ongoing call to specify which effects constitute an `important difference' (Bakan, 1967; Boring, 1919; Hodges \& Lehmann, 1954; Kirk, 1996; Lakens, 2021; Nunnally, 1960; Serlin \& Lapsley, 1985) has become especially urgent.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bakan1967method}{}}%
+Bakan, D. (1967). On method: Toward a reconstruction of psychological investigation. \emph{(No Title)}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-boring1919mathematical}{}}%
+Boring, E. G. (1919). Mathematical vs. Scientific significance. \emph{Psychological Bulletin}, \emph{16}(10), 335.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Bowley1928-yt}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hodges1954matching}{}}%
+Hodges, J., \& Lehmann, E. (1954). Matching in paired comparisons. \emph{The Annals of Mathematical Statistics}, \emph{25}(4), 787--791.
+
 \leavevmode\vadjust pre{\hypertarget{ref-jarke2022roadmap}{}}%
@@
 \emph{Collabra: Psychology}, \emph{8}(1), 57538.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kirk1996practical}{}}%
+Kirk, R. E. (1996). Practical significance: A concept whose time has come. \emph{Educational and Psychological Measurement}, \emph{56}(5), 746--759.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-nunnally1960place}{}}%
+Nunnally, J. (1960). The place of statistics in psychology. \emph{Educational and Psychological Measurement}, \emph{20}(4), 641--650.
+
 \leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
@@
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
+
+\leavevmode\vadjust pre{\hypertarget{ref-serlin1985rationality}{}}%
+Serlin, R. C., \& Lapsley, D. K. (1985). \emph{Rationality in psychological research: The good-enough principle.}
 

### 2025-03-04 paper/paper.Rmd
changes

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4856"
+wordcount         : "4922"
 
@@
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
-These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest [@cumming2012statistical; @transue2019apa].
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest [@cumming2012statistical; APA, -@apa2020publication].
 Researchers often publish studies with low power for medium to small effect sizes [@szucs2017empirical], with the main justification for the sample size being resource limitations [@lakens2022sample].
@@
 The risk is that researchers will heuristically or opportunistically accept all effects as potentially important.
-For this reason "researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it" [@anvari2023not].
+For this reason "researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it" [@anvari2023not, p.
+504].
 
@@
 In this section, we provide guidance and formulas for computing standardized critical effect sizes with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's $d$), correlations (Pearson's $r$), and raw and standardized coefficients in linear models.
+Table \@ref(tab:eqsummary) summarises those equations.
 The same equations can be used to calculate unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
@@
 From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances.
-Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests [@Viechtbauer2015-tn; @Rubio-Aparicio2020-gd].
+Similarly to standard linear regression, hypothesis testing is performed using Wald $t$ or $z$ tests [@Viechtbauer2015-tn; @borenstein2021introduction, p.
+338].
 For this reason, we can apply the same equations used for the linear regression.
@@
 
-Table \@ref(tab:eqsummary) summarises the equations presented above.
-
-```{r eqsummary, echo=FALSE, tab.cap="Summary of the critical effect size equations."}
+```{r eqsummary, echo=FALSE, tab.cap="Summary of the critical effect size equations, see text for details."}
 library(kableExtra)
@@
 
-In this section, we introduce the aforementioned user-friendly implementation of the mathematical computations as functions of the package `criticalESvalue` in R.
+In this section, we introduce the aforementioned user-friendly implementation of the mathematical computations as functions of the package `criticalESvalue` in R (currently hosted on GitHub).
 Here, we demonstrate its application through two examples: one example of a $t$-test on real data and a computation of the critical effect size values for a correlation from sample size.

### 2025-03-04 paper/paper.tex
changes

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4856}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4922}
 \usepackage{csquotes}
@@
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
-These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest (Cumming et al., 2012; Transue, 2019).
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest (Cumming et al., 2012; APA, 2020).
 Researchers often publish studies with low power for medium to small effect sizes (Szucs \& Ioannidis, 2017), with the main justification for the sample size being resource limitations (Lakens, 2022).
@@
 The risk is that researchers will heuristically or opportunistically accept all effects as potentially important.
-For this reason ``researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it'' (Anvari et al., 2023).
+For this reason ``researchers need to explicitly state the mechanisms that can amplify the importance of an observed effect size and the mechanisms that counteract it'' (Anvari et al., 2023, p.
+504).
 
@@
 In this section, we provide guidance and formulas for computing standardized critical effect sizes with examples for frequently encountered effect sizes including Standardized Mean Differences (Cohen's \(d\)), correlations (Pearson's \(r\)), and raw and standardized coefficients in linear models.
+Table \ref{tab:eqsummary} summarises those equations.
 The same equations can be used to calculate unstandardized effect sizes that can be preferred in some situations (e.g., linear models).
@@
 From a statistical point of view, a meta-analysis can be considered as a weighted linear regression with heterogeneous variances.
-Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests (Rubio-Aparicio et al., 2020; Viechtbauer et al., 2015).
+Similarly to standard linear regression, hypothesis testing is performed using Wald \(t\) or \(z\) tests (Borenstein et al., 2021, p.
+338; Viechtbauer et al., 2015).
 For this reason, we can apply the same equations used for the linear regression.
@@
 
-Table \ref{tab:eqsummary} summarises the equations presented above.
-
 \footnotesize
@@
 \centering
-\caption{\label{tab:eqsummary}Summary of the critical effect size equations.}
+\caption{\label{tab:eqsummary}Summary of the critical effect size equations, see text for details.}
 \centering
@@
 
-In this section, we introduce the aforementioned user-friendly implementation of the mathematical computations as functions of the package \texttt{criticalESvalue} in R.
+In this section, we introduce the aforementioned user-friendly implementation of the mathematical computations as functions of the package \texttt{criticalESvalue} in R (currently hosted on GitHub).
 Here, we demonstrate its application through two examples: one example of a \(t\)-test on real data and a computation of the critical effect size values for a correlation from sample size.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-apa2020publication}{}}%
+American Psychological Association. (2020). \emph{Publication manual of the american psychological association} (7th ed.). American Psychological Association. \url{https://doi.org/10.1037/0000165-000}
+
 \leavevmode\vadjust pre{\hypertarget{ref-anvari2023not}{}}%
@@
 Bakan, D. (1967). On method: Toward a reconstruction of psychological investigation. \emph{(No Title)}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-borenstein2021introduction}{}}%
+Borenstein, M., Hedges, L. V., Higgins, J. P., \& Rothstein, H. R. (2021). \emph{Introduction to meta-analysis}. John wiley \& sons.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Rubio-Aparicio2020-gd}{}}%
-Rubio-Aparicio, M., López-López, J.
-A., Viechtbauer, W., Marín-Martínez, F., Botella, J., \& Sánchez-Meca, J.
-(2020).
-Testing categorical moderators in mixed-effects meta-analysis in the presence of heteroscedasticity.
-\emph{Journal of Experimental Education}, \emph{88}, 288--310.
-\url{https://doi.org/10.1080/00220973.2018.1561404}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Satterthwaite1946-aq}{}}%

### 2025-03-04 paper/paper.Rmd
adding equations with only sample size in the table

@@
 
-```{r eqsummary, echo=FALSE, tab.cap="Summary of the critical effect size equations, see text for details."}
+```{r eqsummary, echo=FALSE, tab.cap="Summary of the critical effect size equations, see text for details.
+When possible, we also included the equation to calculate the critical effect size without computing the standard error.
+For the paired t-test, this is feasible when using the standard deviation of differences.
+For the two-sample t-test, it is possible only under the assumption of homogeneity of variances."}
 library(kableExtra)
@@
             "Paired t-test","Paired t-test","Between-sample t-test",
-                       "Correlation","Correlation","Meta-analysis/Regression"),
+            "Correlation test","Correlation","Meta-analysis/Regression"),
   Numerator = c("$b_c = t_c \\times \\mbox{SE}_b$","$b_c = t_c \\times \\mbox{SE}_b$",
@@
   Denominator = c("$s$", NA, "$s_D$", "$s_p$", NA, NA, NA),
-          Equation = c("$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s}$",
+  Equation = c("$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s} = t_c\\sqrt{\\frac{1}{n}}$",
                "$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s_p}$",
-                       "$d_{z_c} = \\frac{t_c \\times \\mbox{SE}_b}{s_D}$",
-                       "$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s_p}$",
+               "$d_{z_c} = \\frac{t_c \\times \\mbox{SE}_b}{s_D} = t_c\\sqrt{\\frac{1}{n}}$",
+               "$d_c = \\frac{t_c \\times \\mbox{SE}_b}{s_p} = t_c\\sqrt{\\frac{1}{n_1} + \\frac{1}{n_2}}$",
                "$r_c = \\frac{t_c}{\\sqrt{n - 2 + t^2_c}}$","$r_{z_c} = \\frac{z_c}{\\sqrt{n - 3}}$",

### 2025-03-04 paper/paper.tex
adding equations with only sample size in the table

@@
 
-\hypertarget{small-sample-sizes-and-uncertain-sample-size-determination}{%
-\subsection{Small Sample Sizes and Uncertain Sample Size Determination}\label{small-sample-sizes-and-uncertain-sample-size-determination}}
+\subsection{Small Sample Sizes and Uncertain Sample Size Determination}\label{small-sample-sizes-and-uncertain-sample-size-determination}
 
@@
 \begin{table}[H]
-\centering
-\caption{\label{tab:eqsummary}Summary of the critical effect size equations, see text for details.}
+
+\caption{\label{tab:eqsummary}Summary of the critical effect size equations, see text for details.
+When possible, we also included the equation to calculate the critical effect size without computing the standard error.
+For the paired t-test, this is feasible when using the standard deviation of differences.
+For the two-sample t-test, it is possible only under the assumption of homogeneity of variances.}
 \centering
@@
 \midrule
-One-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s}$\\
+One-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s} = t_c\sqrt{\frac{1}{n}}$\\
 \cmidrule{1-2}
@@
 \cmidrule{2-2}
-\multirow{-2}{*}{\raggedright\arraybackslash Paired t-test} & $d_{z_c} = \frac{t_c \times \mbox{SE}_b}{s_D}$\\
+\multirow{-2}{*}{\raggedright\arraybackslash Paired t-test} & $d_{z_c} = \frac{t_c \times \mbox{SE}_b}{s_D} = t_c\sqrt{\frac{1}{n}}$\\
 \cmidrule{1-2}
-Between-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s_p}$\\
+Between-sample t-test & $d_c = \frac{t_c \times \mbox{SE}_b}{s_p} = t_c\sqrt{\frac{1}{n_1} + \frac{1}{n_2}}$\\
 \cmidrule{1-2}
- & $r_c = \frac{t_c}{\sqrt{n - 2 + t^2_c}}$\\
-\cmidrule{2-2}
-\multirow{-2}{*}{\raggedright\arraybackslash Correlation} & $r_{z_c} = \frac{z_c}{\sqrt{n - 3}}$\\
+Correlation test & $r_c = \frac{t_c}{\sqrt{n - 2 + t^2_c}}$\\
+\cmidrule{1-2}
+Correlation & $r_{z_c} = \frac{z_c}{\sqrt{n - 3}}$\\
 \cmidrule{1-2}
@@
 
-\hypertarget{examples-in-r}{%
-\section{Examples in R}\label{examples-in-r}}
+\section{Examples in R}\label{examples-in-r}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-apa2020publication}{}}%
+\bibitem[\citeproctext]{ref-apa2020publication}
 American Psychological Association. (2020). \emph{Publication manual of the american psychological association} (7th ed.). American Psychological Association. \url{https://doi.org/10.1037/0000165-000}
 
-\leavevmode\vadjust pre{\hypertarget{ref-anvari2023not}{}}%
+\bibitem[\citeproctext]{ref-anvari2023not}
 Anvari, F., Kievit, R., Lakens, D., Pennington, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bakan1967method}{}}%
+\bibitem[\citeproctext]{ref-bakan1967method}
 Bakan, D. (1967). On method: Toward a reconstruction of psychological investigation. \emph{(No Title)}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-borenstein2021introduction}{}}%
+\bibitem[\citeproctext]{ref-borenstein2021introduction}
 Borenstein, M., Hedges, L. V., Higgins, J. P., \& Rothstein, H. R. (2021). \emph{Introduction to meta-analysis}. John wiley \& sons.
 
-\leavevmode\vadjust pre{\hypertarget{ref-boring1919mathematical}{}}%
+\bibitem[\citeproctext]{ref-boring1919mathematical}
 Boring, E. G. (1919). Mathematical vs. Scientific significance. \emph{Psychological Bulletin}, \emph{16}(10), 335.
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bowley1928-yt}{}}%
+\bibitem[\citeproctext]{ref-Bowley1928-yt}
 Bowley, A. L. (1928). The standard deviation of the correlation coefficient. \emph{Journal of the American Statistical Association}, \emph{23}, 31. \url{https://doi.org/10.2307/2277400}
 
-\leavevmode\vadjust pre{\hypertarget{ref-byers2020building}{}}%
+\bibitem[\citeproctext]{ref-byers2020building}
 Byers-Heinlein, K., Bergmann, C., Davies, C., Frank, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chow1988significance}{}}%
+\bibitem[\citeproctext]{ref-chow1988significance}
 Chow, S. L. (1988). Significance test or effect size? \emph{Psychological Bulletin}, \emph{103}(1), 105.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cohen1994}{}}%
+\bibitem[\citeproctext]{ref-cohen1994}
 Cohen, J. (1994). The earth is round (p \textless{} .05). \emph{American Psychologist}, \emph{49}(12), 997--1003.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cortina1997logic}{}}%
+\bibitem[\citeproctext]{ref-cortina1997logic}
 Cortina, J. M., \& Dunlap, W. P. (1997). On the logic and purpose of significance testing. \emph{Psychological Methods}, \emph{2}(2), 161.
 
-\leavevmode\vadjust pre{\hypertarget{ref-cumming2012statistical}{}}%
+\bibitem[\citeproctext]{ref-cumming2012statistical}
 Cumming, G., Fidler, F., Kalinowski, P., \& Lai, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fisher1915-ks}{}}%
+\bibitem[\citeproctext]{ref-Fisher1915-ks}
 Fisher, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hodges1954matching}{}}%
+\bibitem[\citeproctext]{ref-hodges1954matching}
 Hodges, J., \& Lehmann, E. (1954). Matching in paired comparisons. \emph{The Annals of Mathematical Statistics}, \emph{25}(4), 787--791.
 
-\leavevmode\vadjust pre{\hypertarget{ref-jarke2022roadmap}{}}%
+\bibitem[\citeproctext]{ref-jarke2022roadmap}
 Jarke, H., Anand-Vembar, S., Alzahawi, S., Andersen, T.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kirk1996practical}{}}%
+\bibitem[\citeproctext]{ref-kirk1996practical}
 Kirk, R. E. (1996). Practical significance: A concept whose time has come. \emph{Educational and Psychological Measurement}, \emph{56}(5), 746--759.
 
-\leavevmode\vadjust pre{\hypertarget{ref-kramer2014experimental}{}}%
+\bibitem[\citeproctext]{ref-kramer2014experimental}
 Kramer, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-krueger2001null}{}}%
+\bibitem[\citeproctext]{ref-krueger2001null}
 Krueger, J. (2001). Null hypothesis significance testing: On the survival of a flawed method. \emph{American Psychologist}, \emph{56}(1), 16.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2013calculating}{}}%
+\bibitem[\citeproctext]{ref-lakens2013calculating}
 Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: A practical primer for t-tests and ANOVAs. \emph{Frontiers in Psychology}, \emph{4}, 62627.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2021practical}{}}%
+\bibitem[\citeproctext]{ref-lakens2021practical}
 Lakens, D. (2021). The practical alternative to the p value is the correctly used p value. \emph{Perspectives on Psychological Science}, \emph{16}(3), 639--648.
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakens2022sample}{}}%
+\bibitem[\citeproctext]{ref-lakens2022sample}
 Lakens, D. (2022). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.
 
-\leavevmode\vadjust pre{\hypertarget{ref-mesquida2024effect}{}}%
+\bibitem[\citeproctext]{ref-mesquida2024effect}
 Mesquida, C., \& Lakens, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-miller2017hypothesis}{}}%
+\bibitem[\citeproctext]{ref-miller2017hypothesis}
 Miller, J. (2017). Hypothesis testing in the real world. \emph{Educational and Psychological Measurement}, \emph{77}(4), 663--672.
 
-\leavevmode\vadjust pre{\hypertarget{ref-morris2002combining}{}}%
+\bibitem[\citeproctext]{ref-morris2002combining}
 Morris, S. B., \& DeShon, R. P. (2002). Combining effect size estimates in meta-analysis with repeated measures and independent-groups designs. \emph{Psychological Methods}, \emph{7}(1), 105.
 
-\leavevmode\vadjust pre{\hypertarget{ref-moshontz2018psychological}{}}%
+\bibitem[\citeproctext]{ref-moshontz2018psychological}
 Moshontz, H., Campbell, L., Ebersole, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nickerson2000null}{}}%
+\bibitem[\citeproctext]{ref-nickerson2000null}
 Nickerson, R. S. (2000). Null hypothesis significance testing: A review of an old and continuing controversy. \emph{Psychological Methods}, \emph{5}(2), 241.
 
-\leavevmode\vadjust pre{\hypertarget{ref-nunnally1960place}{}}%
+\bibitem[\citeproctext]{ref-nunnally1960place}
 Nunnally, J. (1960). The place of statistics in psychology. \emph{Educational and Psychological Measurement}, \emph{20}(4), 641--650.
 
-\leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
+\bibitem[\citeproctext]{ref-orben_crud_2020}
 Orben, A., \& Lakens, D. (2020). Crud ({Re}){Defined}. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(2), 238--247. \url{https://doi.org/10.1177/2515245920917961}
 
-\leavevmode\vadjust pre{\hypertarget{ref-riesthuis2024simulation}{}}%
+\bibitem[\citeproctext]{ref-riesthuis2024simulation}
 Riesthuis, P.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Satterthwaite1946-aq}{}}%
+\bibitem[\citeproctext]{ref-Satterthwaite1946-aq}
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
 
-\leavevmode\vadjust pre{\hypertarget{ref-serlin1985rationality}{}}%
+\bibitem[\citeproctext]{ref-serlin1985rationality}
 Serlin, R. C., \& Lapsley, D. K. (1985). \emph{Rationality in psychological research: The good-enough principle.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-sirois2023pupil}{}}%
+\bibitem[\citeproctext]{ref-sirois2023pupil}
 Sirois, S., Brisson, J., Blaser, E., Calignano, G., Donenfeld, J., Hepach, R., Hochmann, J.-R., Kaldy, Z., Liszkowski, U., Mayer, M., et al.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2015-tn}{}}%
+\bibitem[\citeproctext]{ref-Viechtbauer2015-tn}
 Viechtbauer, W., López-López, J.

### 2025-03-04 paper/paper.Rmd
minor updates

@@
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
-These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest [@cumming2012statistical; APA, -@apa2020publication].
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest [@cumming2012statistical; American Psychological Association [APA], -@apa2020publication].
 Researchers often publish studies with low power for medium to small effect sizes [@szucs2017empirical], with the main justification for the sample size being resource limitations [@lakens2022sample].
@@
 This is not intended to replace other strategies aimed at enhancing the NHST approach to inference.
-Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [@transue2019apa] or the \textit{a priori} planning for statistical power are valuable in their own right.
+Such strategies, such as the emphasis on estimating effect sizes with confidence intervals [APA, -@apa2020publication] or the \textit{a priori} planning for statistical power are valuable in their own right.
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).

### 2025-03-04 paper/paper.tex
minor updates

--- 2025-03-04 paper/paper.tex
@@
 Over the past decades, numerous proposals have emerged to improve the use of hypothesis testing.
-These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest (Cumming et al., 2012; APA, 2020).
+These include complementing hypothesis tests with effect sizes and their confidence intervals, preregistering hypotheses before the data is collected, and conducting power analyses to determine the sample size \textit{a priori} based on a smallest effect size of interest (Cumming et al., 2012; American Psychological Association {[}APA{]}, 2020).
 Researchers often publish studies with low power for medium to small effect sizes (Szucs \& Ioannidis, 2017), with the main justification for the sample size being resource limitations (Lakens, 2022).
@@
 \centering
-\caption{\label{tab:eqsummary}Summary of the critical effect size equations, see text for details.}
+\caption{\label{tab:eqsummary}Summary of the critical effect size equations, see text for details.
+When possible, we also included the equation to calculate the critical effect size without computing the standard error.
+For the paired t-test, this is feasible when using the standard deviation of differences.
+For the two-sample t-test, it is possible only under the assumption of homogeneity of variances.}
 \centering
@@
 This is not intended to replace other strategies aimed at enhancing the NHST approach to inference.
-Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (Transue, 2019) or the \textit{a priori} planning for statistical power are valuable in their own right.
+Such strategies, such as the emphasis on estimating effect sizes with confidence intervals (APA, 2020) or the \textit{a priori} planning for statistical power are valuable in their own right.
 Instead, our proposal serves as a complementary tool, especially beneficial for facilitating the interpretation of results when statistical power deviates from an optimal level (typically falling below, but occasionally exceeding it).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-transue2019apa}{}}%
-Transue, B. (2019). \emph{APA style 7th edition}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-Viechtbauer2007-tq}{}}%

### 2025-03-04 paper/paper.Rmd
changes

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4922"
+wordcount         : "4934"
 
@@
 Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
-This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to significant but very small and meaningless non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to significant but very small and not meaningfully interpretable non-zero effects, a phenomenon referred to as the 'crud factor' [@orben_crud_2020].
 

### 2025-03-04 paper/paper.tex
changes

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4922}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4934}
 \usepackage{csquotes}
@@
 Reporting critical effect size values will focus the attention of researchers on the difference between statistical significance and practical significance, and raises awareness of the importance to interpret the size of the effect.
-This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to significant but very small and meaningless non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
+This is especially important in correlational studies with large sample sizes in psychology, where systematic but uncontrolled sources of variability may lead to significant but very small and not meaningfully interpretable non-zero effects, a phenomenon referred to as the `crud factor' (Orben \& Lakens, 2020).
 

### 2025-03-05 paper/paper.Rmd
final

@@
 keywords          : critical effect size values, statistical significance, hypothesis testing, effect sizes 
-wordcount         : "4934"
+wordcount         : "4927"
 
@@
 Subsequently, they compute the critical effect size values, revealing Cohen's $d = -0.75$ and $d = 0.75$.
-This indicates that any observed Cohen's $d$ between -0.75 and 0.75 will certainly fail to reach statistical significance.
+This indicates that any observed Cohen's $d$ between $-0.75$ and $0.75$ will certainly fail to reach statistical significance.
 By reporting these critical effect size values the researchers transparently convey that estimated effects much larger than most effects in the psychological literature will always fall short of significance.
@@
 The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected [@weiss2014spatial].
-The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
-Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+The two groups were compared on a variety of tasks, related to verbal working memory and tasks related to auditory skills.
+Regarding the verbal working memory, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
 A two-tailed $t$-test was conducted to compare the groups.
-The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's $d$ calculated on summary statistics is of $-.74$, critical $d = \pm .62$.
-But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $\pm.62$.
+The resulting effect was statistically significant for the maximal span, $t(55) = -2.5$, $p = .017$, the Cohen's $d$ calculated on summary statistics is of $-.74$, critical $d = - .62$ and $d = .62$.
+But for the total number of sequences the test did not reach statistical significance ($t(55) = -1.8$, $p = 0.076$), actual $d = -0.31$ and critical $d$ again $-.62$ and $.62$.
 Thus, given the limited sample size, a result must be of rather large magnitude to be associated with $p < 0.05$, while medium to small effect sizes will simply fail to reach significance.
@@
 Imagine researchers who gain access to a very large archival dataset ($n = 5,000$) and decide to explore bivariate correlations between variables.
-With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are $r = \pm 0.03$.
+With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are $r = -0.03$ and $r = 0.03$.
 It is worth noting that such small effects may potentially reflect just minor artifacts [@wilson2020science], such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
@@
 Based on this and other results from the study, the authors conclude that "the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks" [@kramer2014experimental].
-The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $+0.006$.
+The critical values for Cohen's $d$ in this experimental design are $-0.006$ and $0.006$.
 It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
@@
 Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results [@anvari2023not].
-For this reason, the ongoing call to specify which effects constitute an ‘important difference’ [@bakan1967method; @boring1919mathematical; @hodges1954matching; @kirk1996practical; @lakens2021practical; @nunnally1960place; @serlin1985rationality] has become especially urgent.
+For this reason, the ongoing call to specify which effects constitute an ‘important difference’ [@boring1919mathematical; @hodges1954matching; @kirk1996practical] has become especially urgent.
 

### 2025-03-05 paper/paper.tex
final

@@
 {}{\PatchFailed}
-\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4934}
+\keywords{critical effect size values, statistical significance, hypothesis testing, effect sizes\newline\indent Word count: 4927}
 \usepackage{csquotes}
@@
 Subsequently, they compute the critical effect size values, revealing Cohen's \(d = -0.75\) and \(d = 0.75\).
-This indicates that any observed Cohen's \(d\) between -0.75 and 0.75 will certainly fail to reach statistical significance.
+This indicates that any observed Cohen's \(d\) between \(-0.75\) and \(0.75\) will certainly fail to reach statistical significance.
 By reporting these critical effect size values the researchers transparently convey that estimated effects much larger than most effects in the psychological literature will always fall short of significance.
@@
 The study of interest was conducted on musicians and non musicians to detect differences in working memory, a modest sample size of 57 participants (42 musicians and 15 controls) was collected (Weiss et al., 2014).
-The two groups were compared on a variety of tasks, related to verbal working memory (VWM) and tasks related to auditory skills.
-Regarding the VWM, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
+The two groups were compared on a variety of tasks, related to verbal working memory and tasks related to auditory skills.
+Regarding the verbal working memory, the syllable-span task was administered and scores were computed for the maximal span and for the total number of sequences they correctly repeated.
 A two-tailed \(t\)-test was conducted to compare the groups.
-The resulting effect was statistically significant for the maximal span, \(t(55) = -2.5\), \(p = .017\), the Cohen's \(d\) calculated on summary statistics is of \(-.74\), critical \(d = \pm .62\).
-But for the total number of sequences the test did not reach statistical significance (\(t(55) = -1.8\), \(p = 0.076\)), actual \(d = -0.31\) and critical \(d\) again \(\pm.62\).
+The resulting effect was statistically significant for the maximal span, \(t(55) = -2.5\), \(p = .017\), the Cohen's \(d\) calculated on summary statistics is of \(-.74\), critical \(d = - .62\) and \(d = .62\).
+But for the total number of sequences the test did not reach statistical significance (\(t(55) = -1.8\), \(p = 0.076\)), actual \(d = -0.31\) and critical \(d\) again \(-.62\) and \(.62\).
 Thus, given the limited sample size, a result must be of rather large magnitude to be associated with \(p < 0.05\), while medium to small effect sizes will simply fail to reach significance.
@@
 Imagine researchers who gain access to a very large archival dataset (\(n = 5,000\)) and decide to explore bivariate correlations between variables.
-With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are \(r = \pm 0.03\).
+With such a large sample size, very small correlations will reach statistical significance, and the critical effect size values for a two-sided test are \(r = -0.03\) and \(r = 0.03\).
 It is worth noting that such small effects may potentially reflect just minor artifacts (Wilson et al., 2020), such as subtle experimenter effects or slight non-independence among observations, even in meticulously designed studies.
@@
 Based on this and other results from the study, the authors conclude that ``the emotions expressed by friends, via online social networks, influence our own moods, constituting, to our knowledge, the first experimental evidence for massive-scale emotional contagion via social networks'' (Kramer et al., 2014).
-The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(+0.006\).
+The critical values for Cohen's \(d\) in this experimental design are \(-0.006\) and \(0.006\).
 It should have been clear from the outset that the researchers should have considered the effect size beyond merely testing for the statistical significance.
@@
 Signaling the critical effect size values beforehand can serve as a clear warning that statistically significant results should not automatically be interpreted as practically significant, thus urging caution when interpreting the results (Anvari et al., 2023).
-For this reason, the ongoing call to specify which effects constitute an `important difference' (Bakan, 1967; Boring, 1919; Hodges \& Lehmann, 1954; Kirk, 1996; Lakens, 2021; Nunnally, 1960; Serlin \& Lapsley, 1985) has become especially urgent.
+For this reason, the ongoing call to specify which effects constitute an `important difference' (Boring, 1919; Hodges \& Lehmann, 1954; Kirk, 1996) has become especially urgent.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bakan1967method}{}}%
-Bakan, D. (1967). On method: Toward a reconstruction of psychological investigation. \emph{(No Title)}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-borenstein2021introduction}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nunnally1960place}{}}%
-Nunnally, J. (1960). The place of statistics in psychology. \emph{Educational and Psychological Measurement}, \emph{20}(4), 641--650.
-
 \leavevmode\vadjust pre{\hypertarget{ref-orben_crud_2020}{}}%
@@
 Satterthwaite, F. E. (1946). An approximate distribution of estimates of variance components. \emph{Biometrics Bulletin}, \emph{2}, 110. \url{https://doi.org/10.2307/3002019}
-
-\leavevmode\vadjust pre{\hypertarget{ref-serlin1985rationality}{}}%
-Serlin, R. C., \& Lapsley, D. K. (1985). \emph{Rationality in psychological research: The good-enough principle.}
 
